### 2017-05-09 paper/introduction.tex
case, abstract

The Vienna Development Method (VDM) is a formal method that supports modeling and analysis of software systems at various levels of abstraction. By the VDM specification language (VDM-SL), one can specify software systems with rich types,

### 2017-05-09 paper/studies.tex
case, abstract

This section gives studies that apply our encoding to selected VDM-SL models and discusses the results. All the selected VDM-SL models are from the Overture repository\footnote{a url link here}.

\subsection{Abstract Pacemaker}

This model specifies an abstract pacemaker with some core functionalities of pulse trace selection\footnote{This model was developed by Sten Agerholm et al. in 1999 in connection with FM'99}. The data types used in the model are defined as follows:
% The model is described in VDM-SL as a short, flat specification. This enables abstraction from design considerations and ensures maximum focus on high-level, precise and systematic analysis

\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\small]
  Trace = seq of [Event];
  Event = <A> | <V>;

  state Pacemaker of
    aperiod : nat 
    vdelay  : nat
  init p == p = mk_Pacemaker(15,10)
  end
\end{Verbatim}
\end{mdframed}

{\tt Trace} is of sequence type of {\tt Event}, where {\tt Event} is of quote type with two values {\tt <A>} and {\tt <V>}. The state is defined as two natural numbers with initial value of {\tt (15,10)}. In Z3, natural numbers is encoded as positive integers (integers with constraint of greater and equal to zero); quote type can be encoded as enumeration type {\tt EnumSort}; sequence type can be encoded as {\tt ArraySort}. The definitions of {\tt Event} and {\tt Trace} are as follows:

\begin{mdframed}[roundcorner=5pt,shadow=true]
\begin{Verbatim}[fontsize=\small]
  Event_lift, (A, V, nil, NDF) = 
    EnumSort('Event_lift', 
            ['A', 'V', 'nil', 'NDF'])

  Trace = ArraySort(IntSort(),Event_lift)
\end{Verbatim}
\end{mdframed}

Because VDM uses partial functions, to encode as SMT formulas in Z3, we need to lift {\tt Event} to include the undefined case {\tt NDF}. Also, square brackets of {\tt [Evnet]} means the value can be {\tt nil}. Furthermore, since {\tt Trace} is defined as an array in Z3, we added constraints that limit an array as a sequence. Note that in Z3, the constraints are applied on each instance of type {\tt Trace} ({\tt tr} in the following formulas), not on the type itself.

\begin{mdframed}[roundcorner=5pt,shadow=true]
\begin{Verbatim}[fontsize=\small]
  tr = Const('tr',Trace)
  [i,j] = Ints('i j')

  ForAll(i, Implies( i<=0, tr[i]==NDF ) )

  ForAll(i,
    Implies(
      And(i>=1,tr[i]!=NDF),
      ForAll(j,
        Implies(And(j>=1,j<=i),tr[j]!=NDF)
      )
    )
  )

  ForAll(i,
    Implies(
      And(i>=1,tr[i]==NDF),
      ForAll(j,
        Implies(j>=i,tr[j]==NDF)
      )
    )
  )
\end{Verbatim}
\end{mdframed}

The first constraint says that the all indexes of {\tt tr} below 1 should be undefined because in VDM-SL, a sequence's index starts from 1. The second constraint says that if an index is defined, i.e., not undefined, all indexes lower than it should be all defined. The third constraint says tat if an index is undefined, all indexes higher than it should be all undefined. From the above three constraints, we limited an array as a sequence that starts from index 1 to its last element.

% \begin{itemize}
% \item
% A sequence's index is from 1, so indexes below 1 ($<=0$) should be all undefined.
% $\forall i \in \mathcal{Z}, i<=0. ~tr[i] = NDF $
% \item
% If an index is defined in the sequence, all indexes lower then it should be all defined.
% $\forall i \in \mathcal{Z}. ~tr[i] \neq NDF \rightarrow (\forall j \in \mathcal{Z}, 1<=j \land j <= i. ~tr[j] \neq NDF)$
% \item
% If an index is undefined, all indexes higher then it should be all undefined.
% $\forall i \in \mathcal{Z}. ~(i >= 1 \land tr[i] = NDF) \rightarrow (\forall j \in \mathcal{Z}, j >= i. ~tr[j] = NDF)$
% \end{itemize}

% where $NDF$ stands for undefined. The above formulas are added as constraints whenever the instance {\tt tr} of type {\tt Trace} is used.

Since in the VDM-SL model, the length function of a sequence is used, we also need to define the length function in Z3. Here we defined an uninterpreted function {\tt len\_tr} for {\tt tr} the instance of {\tt Trace}.
% $(len\_tr(tr) = 0 \land tr[1] = 0) \lor (len\_tr(tr) > 0 \land tr[len\_tr(tr)] \neq NDF \land tr[len\_tr(tr)+1] = NDF$

\begin{mdframed}[roundcorner=5pt,shadow=true]
\begin{Verbatim}[fontsize=\small]
len_tr=Function('len_tr',Trace,IntSort())

Or(
  And(
    len_tr(tr)==0,
    tr[1]==NDF
  ),
  And(
    len_tr(tr)>0,
    tr[len_tr(tr)]!=NDF,
    tr[len_tr(tr)+1]==NDF
  )
)
\end{Verbatim}
\end{mdframed}

{\tt len\_tr} is defined as an uninterpreted function that takes {\tt tr} as input and returns an integer indicates the length of {\tt tr}. Since the length of a sequence can only be equal or greater than 0, the constraint of {\tt len\_tr} is divided into two cases:

\begin{enumerate}
\item
If {\tt len\_tr} returns 0, then {\tt tr} is an sequence with no element. That is, {\tt tr[1]} has to be undefined so that all indexes of {\tt tr} are then undefined based on previous constraints.
\item
If {\tt len\_tr} returns an integer greater than 0, then {\tt tr} has defined elements till its length index {\tt len\_tr(tr)} such that the element at and after index {\tt len\_tr(tr)+1} should be undefined.
\end{enumerate}

We have defined the context information related to type {\tt Trace}. Now we can procceed to encode and prove the POs of the model. For this VDM-SL model, Overture generated nine POs. Most of the POs are similar so we only demonstrate PO8 and PO3.

$\bullet$ PO8 is a legal sequence application obligation of operation {\tt Pace} which is explicitly specified. This PO requires that the formulas specified with trace {\tt tr} be legal, i.e., computable to get its result.

\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\small]
(forall tr:Trace, aperi:nat1, vdel:nat1,
  oldstate & 
    (forall i in set (inds (tl tr)) &
      (((i mod aperi) = (vdel + 1)) =>
        (i in set (inds tr)))
))
\end{Verbatim}
\end{mdframed}

In PO8 showed above, {\tt tl tr} is the tail of {\tt tr}, and {\tt inds} is the operator of gathering indexes of defined elements of {\tt tr} as the set of natural numbers. To prove PO8, firstly we took negation of PO8 which results a quantifier eliminated formula.

\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\small]
Exists tr:Trace, aperi:nat1, vdel:nat1 &
  (Exists i in set (inds (tl tr)) &
     Not(
      ((i mod aperi) = (vdel + 1)) =>
       (i in set (inds tr))
     )
  )
\end{Verbatim}
\end{mdframed}

Note that the {\tt oldstate} is not used in the operation so that we remove it in the above formula. To encode the PO in Z3, the PO can be recognized as the formula $\exists~[tr:Trace,aperi:nat1,vdel:nat1,i:nat] P \land Q \land \neg R$ where $P$, $Q$, and $\neg R$ can be encoded separately.

\begin{mdframed}[roundcorner=5pt,shadow=true]
\begin{Verbatim}[fontsize=\small]
  And(aperi>=1,vdel>=1)        #nat1
  And(i>=1, i<=len_tr(tr)-1)   #P
  i%aperi==vdel+1              #Q
  Not(And(i>=1,i<=len_tr(tr))) #not R
\end{Verbatim}
\end{mdframed}

In the encoding of $P$, the tail related formula can be encoded as in the range of $[1,(len~tr)-1]$ so that it is not required to define {\tt tl} and {\tt inds} for sequences. Finally, we checked the satisfiability of the encoded Z3 code and got {\tt unsat}, which means that the PO is proved since we have negated it.

% {\tt Periodic}: legal sequence application obligation\\
% This PO is meant to represent the applicability of statements used in the function {\tt Periodic}, that is, the expressions in {\tt Periodic} are computable to get a result. The formula of the PO is in the form of 
% $\forall tr,e,p. ~\forall t \in inds ~tr. ~(~P \rightarrow ~(~Q \rightarrow ~(~R \rightarrow ~(~S \rightarrow ~(~\forall i. ~T \rightarrow U~)~)~)~)$. where 
% \begin{itemize}
% \item
% $tr: Trace$;~$e:Event$;~$p:nat1$;~$t:nat$
% \item
% $P = t \in set~(inds~tr)$
% \item
% $Q = (tr(t)=e)$
% \item
% $R = (t+p) ~<=~ (len~tr)$
% \item
% $S = (tr((t+p))=e) ~\land~ \forall i \in \{(t+1, \ldots, (t+p)-1\}. ~tr(i) \neq e )$
% \item
% $T = ((t+p) > (len~tr)~)$
% \item
% $U = (\forall i \in \{(t+1), \ldots, len~tr \}. ~i \in set (inds~tr)$
% \end{itemize}

% After taken negation of the formula, we got 
% $\exists tr,e,p,t,i. ~(P \land Q \land Q \land S \land T \land \neg U~)$. In the negated case, the expected result of the negated PO is unsatisfiable, i.e., unsat. The

$\bullet$ PO3 is the postcondition satisfiable obligation of the implicitly specified operation {\tt FaultHeart} that generates a trace of heart pulse signals. PO3 says that the postcondition of {\tt FaultHeart} has to be satisfiable so that the specification of {\tt FaultHeart} is implementable.

% \medskip
\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\small]
exists tr:Trace &
  post_FaultyHeart(oldstate, tr, newstate)

FaultyHeart() tr : Trace
post len tr = 100 and
  Periodic(tr,<A>,aperiod) and 
  not Periodic(tr,<V>,aperiod);
\end{Verbatim}
\end{mdframed}
% \medskip

The above VDM-SL code also shows that definition of {\tt FaultyHeart}. Recall that the state is defined as a pair of natural numbers {\tt aperiod} and {\tt vdelay}. The postcondition of {\tt FaultyHeart} is not specified with {\tt oldstate} which has suffixed tilde, i.e, {\tt aperiod\textasciitilde} and {\tt vdelay\textasciitilde}. Therefore, PO3 is a simplified postcondition satisfiable obligation saying that there exists a {\tt Trace} {\tt tr} which satisfies the postcondition specified with only {\tt newstate}\footnote{How the {\tt newstate} is calculated from {\tt oldstate} is not specified in {\tt FaltyHeart}.}. The postcondition uses the function {\tt Periodic} which is explicitly specified and returns a boolean value. We need to encode {\tt Periodic} as well to prove PO3.

\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\small]
Periodic: Trace * Event * nat1 -> bool
Periodic(tr,e,p) ==
  forall t in set inds tr &
   (tr(t) = e) =>
   (t + p <= len tr =>
   ((tr(t+p) = e and
     forall i in set {t+1, ..., t+p-1} &
       tr(i) <> e)) and
    (t + p > len tr =>
     forall i in set {t+1, ..., len tr} &
       tr(i) <> e));
\end{Verbatim}
\end{mdframed}

While the specification of {\tt Periodic} is relatively long. The encoding of {\tt Periodic} is similar to the encoding of PO8. We may observe some similar computations on the trace {\tt tr} such as {\tt in set inds tr} and {\tt len tr}, and we can encode them in the same way as handling PO8. Note that we did not apply negation to the formula since existence quantifier in the prefix of a formula is preferred in Z3. As a result, we got a result of {\tt sat} so that PO3 is proved.

The results of all nine POs of the Abstracted Pacemaker VDM-SL model is showed in Table~\ref{tbl:result1} with information of whether negation is applied, and the time used.

\begin{table}[htb]
\begin{center}
\begin{tabular}{|c|c|r|r|}
\hline
PO\#	&	negated?	&	result	&	time (sec.) \\ \hline
1		&	Y			&	sat		&	0.031 \\ \hline
2		&	Y			&	unsat	&	0.031 \\ \hline
3		&	N			&	sat		&	15.109 \\ \hline
4		&	Y			&	unsat	&	0.032 \\ \hline
5		&	Y			&	unsat	&	0.046 \\ \hline
6		&	Y			&	unsat	&	0.048 \\ \hline
7		&	Y			&	unsat	&	0.062 \\ \hline
8		&	Y			&	unsat	&	0.031 \\ \hline
9		&	Y			&	sat		&	0.047 \\ \hline
\end{tabular}
\end{center}
\caption{Absract Pacemaker Result}
\label{tbl:result1}
\end{table}

\subsection{Telephone Exchange}

This model specifies an abstracted telephone exchange system. In this model, the operations specify the events which can be initiated either by the system or by a subscriber (user) with implicit style. The system state monitors the calling status of users and the connecting status among users. Type in the model are based on quote types used to indicate the discrete states of users, then maps are specified to relate the calling and connecting status to users as the system state. An invariant is specified for the state {\tt Exchange} to declare the constraints among users and their status in the system.

\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\small]
Subscriber = token;
Initiator =  <AI> | <WI> | <SI>;
Recipient = <WR> | <SR>;
Status = <fr> | <un> |
         Initiator | Recipient;

state Exchange of
  status: map Subscriber to Status
  calls:  inmap Subscriber to Subscriber
inv mk_Exchange(status, calls) == 
  forall i in set dom calls & 
    (status(i) = <WI> and
     status(calls(i)) = <WR>) or
    (status(i) = <SI> and
     status(calls(i)) = <SR>)
init s == s = mk_Exchange({|->},{|->})
end
\end{Verbatim}
\end{mdframed}

The strategy of encoding the above types is (1) applying user-defined types in Z3 to encode the quote types; (2) applying uninterpreted functions to encode map type. Here we only demonstrate a few selected types since the others are similar.

\begin{mdframed}[roundcorner=5pt,shadow=true]
\begin{Verbatim}[fontsize=\small]
Recipient = Datatype('Recipient')
Recipient.declare('WR')
Recipient.declare('SR')
Recipient = Recipient.create()

Status = Datatype('Status')
Status.declare('fr')
Status.declare('un')
Status.declare('INITIATOR',
               ('get_initiator', Initiator))
Status.declare('RECIPIENT',
               ('get_recipient', Recipient))
Status = Status.create()

Status_lift = Datatype('Status_lift')
Status_lift.declare('STATUS',
                    ('get_status', Status))
Status_lift.declare('NDF')
Status_lift = Status_lift.create()

status = Function('status',
                  Subscriber,Status_lift)
calls = Function('calls',
                 Subscriber,Subscriber_lift)
\end{Verbatim}
\end{mdframed}

From above code, the type {\tt Recipient} is defined as an user-defined type with values {\tt WR} and {\tt SR}. Type {\tt Status} is further defined as a user-defined type including {\tt Recipient} and {\tt Initiator}. Note that since VDM uses partial functions, we also defined a lifted type of {\tt Status}. Though {\tt Subscriber} is of token type which does not have specific values when declared, we treated token types as quote types with predefined values such as {\tt S1}, {\tt S2}, and so on. Finally the two maps {\tt status} and {\tt calls} are defined as uninterpreted functions. Overture generated 27 POs for this model. Here we present the encoding and proving of PO1 and PO14.

$\bullet$ PO1 is a legal map application obligation that checks whether maps in the system state are applicable.

\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\small]
forall
  mk_Exchange(status, calls):EXCH`Exchange &
    (forall i in set (dom calls) &
      (i in set (dom status))
    )
)
\end{Verbatim}
\end{mdframed}

Since PO1 explicitly uses the system state, we need to consider the invariant of the system state as the context information of PO1. 
Instead of demonstrate the full encoding of the invariant, we only demonstrate the key encoding for {\tt i in set (dom calls)} and {\tt i in set (dom status)}. Though there is set inclusion used, instead of defining a set, we can encode the set inclusion as whether {\tt calls(i)} is defined or not.

\begin{mdframed}[roundcorner=5pt,shadow=true]
\begin{Verbatim}[fontsize=\small]
  calls(i) != Subscriber_lift.NDF
  status(i)!= Status_lift.NDF
\end{Verbatim}
\end{mdframed}

For PO1, we applied negation to the formula so that the {\tt forall} quantifier for maps {\tt calls} and {\tt status} is removed. The result was {\tt unsat} which means PO1 was proved.

\begin{mdframed}[roundcorner=5pt,shadow=true]
\begin{Verbatim}[fontsize=\small]
Not(
  ForAll(i,
    Implies(
      calls(i)!=Subscriber_lift.NDF,
      status(i)!=Status_lift.NDF
    )
  )
)
\end{Verbatim}
\end{mdframed}

$\bullet$ PO14 is an enumeration map injectivity obligatoin of the operation {\tt Answer} of the system. PO14 has free variables, {\tt r} and {\tt status}, in its formula. These free variables are treated as the quantifier {\tt forall} is prefixed. There are also intermediate variables in the formula such as {\tt m1}, {\tt m2}, {\tt d3}, and {\tt d4}. Note that the inverse function of a map, and the range restriction function {\tt :>} are {\tt applied} on calls and {\tt status}. Also note that {\tt calls} is a one-to-one map {\tt inmap}.

\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\small]
(r in set (dom (status :> {<WR>}))) =>
  (forall m1, m2 in 
    set {{r |-> <SR>}, 
         {(inverse calls)(r) |-> <SI>}} &
    (forall d3 in set (dom m1),
            d4 in set (dom m2) &
      ((d3 = d4) => (m1(d3) = m2(d4)))
))
\end{Verbatim}
\end{mdframed}

We negated PO14 so that the {\tt forall} quantifiers are removed. To encode the negated formula of PO14, similar technique for  encoding {\tt set (dom status)} for PO1 can be applied. For the range restriction function, since the range is a single set with value {\tt <WR>}, there is no need of applying set notations but limiting the map value {\tt status(r)} to {\tt <WR>}. Thus, {\tt r in set (dom (status :> \{<WR>\}))} is encoded as follows.

\begin{mdframed}[roundcorner=5pt,shadow=true]
\begin{Verbatim}[fontsize=\small]
status(r)==
  Status_lift.STATUS(
    Status.RECIPIENT(Recipient.WR) )
\end{Verbatim}
\end{mdframed}

For the map inverse function, we encoded {\tt m1 = \{(inverse calls)(r) |-> <SI>\}} as multiple conditions: (1) the part of map, where key {\tt i} is used and inverse is applied, should be defined; (2) the inversed domain, i.e., the range of original map, is {\tt r}; (3) then {\tt m1(i)} is {\tt <SI>}; (4) for all other indexes, {\tt m1} should be undefined.

\begin{mdframed}[roundcorner=5pt,shadow=true]
\begin{Verbatim}[fontsize=\small]
# inverse map is not empty
calls(i)!=Subscriber_lift.NDF 
# inverse_calls(r) == i
Subscriber_lift.get_subscriber(calls(i))==r

m1(i)==Status_lift.STATUS(
         Status.INITIATOR(Initiator.SI)),
ForAll(k, 
  Implies(k!=i, m1(k)==Status_lift.NDF) )
\end{Verbatim}
\end{mdframed}

The result of checking PO14 ware expected to be {\tt unsat} but we got {\tt sat} and a model that satisfies the formula and constraints.

\subsection{Discussion}

Overture generated 27 POs for this model. These POs can be divided into several groups as showed below. In this case study we encoded and checked the first three groups using the above encoding of types. The reason of not handling the postcondition obligations is that the maps defined using uninterpreted functions cannot be used together with quantifiers. Also, map inverse obligations have {\tt is\_(name, type)}, the function of type check provided by VDM, in their formulas. There is no appropriate encoding for such type check functions in Z3.

\begin{enumerate}
\item
Legal map application obligations, invariant satisfiable obligations: 1 to 7.
\item
Enumeration map injectivity obligations: 9, 14, 18, 21, 25.
\item
Legal map application obligations with inverse map: 12, 17, 20, 23.
\item
Operation postcondition satisfiable obligations: 8, 10, 11, 15, 16, 19, 22, 26, 27
\item
Map inverse obligations: 13, 24.
\end{enumerate}

### 2017-05-09 AbstractPacemaker_po.py

@@ -1,0 +1,483 @@
+from z3 import *
+import time
+
+s = Solver()
+
+#types
+Event_lift, (A, V, nil, NDF) = EnumSort('Event_lift', ['A', 'V', 'nil', 'NDF'])
+Trace = ArraySort(IntSort(),Event_lift)
+
+#state veriable mk_Pacemaker(aperiod,vdelay)
+aperiod = Int('aperiod')
+s.add(aperiod>=0) # nat
+vdelay = Int('vdelay')
+s.add(vdelay>=0)  # nat
+
+#variable tr: Trace
+tr = Const('tr',Trace)
+
+#constraints of sequence
+[i,j] = Ints('i j')
+
+#Index i must from 0, trace is undefined when i below 0
+s.add(
+  ForAll(i, Implies( i<=0, tr[i]==NDF ) )
+)
+
+#trace should be within a index range from 0 to its length 
+s.add(
+  ForAll(i, #if an index is defined, all indices lower then it are defined.
+    Implies(
+      And(i>=1,tr[i]!=NDF),
+      ForAll(j,
+        Implies(And(j>=1,j<=i),tr[j]!=NDF)
+      )
+    )
+  ),
+  ForAll(i, #if an index is undefined, all indices higher then it are undefined.
+    Implies(
+      And(i>=1,tr[i]==NDF),
+      ForAll(j,
+        Implies(j>=i,tr[j]==NDF)
+      )
+    )
+  )
+)
+
+# defined len as a function
+len_tr = Function('len_tr',Trace,IntSort())
+#s.add(
+#  ForAll(tr, len_tr(tr)>=0)
+#)
+s.add(
+  Or(
+    And(
+      len_tr(tr)==0,
+      tr[1]==NDF
+    ),
+    And(
+      len_tr(tr)>0,
+      tr[len_tr(tr)]!=NDF,
+      tr[len_tr(tr)+1]==NDF
+    )
+  )
+)
+#s.push()
+#s.add(len_tr(tr)==1)
+#s.check()
+#print s.model()
+#s.pop()
+
+
+#Proof Obligation 1
+#FaultyHeart: type compatibility obligation @ in 'DEFAULT' (AbstractPacemaker.vdmsl) at line 30:22
+#(((len tr) = 100) => (aperiod > 0))
+"""
+The proof obligation is in the form:
+  ForAll[tr,aperiod] (P -> Q)
+After negated, the obligation becomes the form:
[…]

### 2017-05-19 paper/introduction.tex
intro, vdm

@@
-The Vienna Development Method (VDM) is a formal method that supports modeling and analysis of software systems at various levels of abstraction.
-By the VDM specification language (VDM-SL), one can specify software systems with rich types,
+% General information about VDM
+The Vienna Development Method (VDM)~\cite{Jones:1990:SSD:94062,Fitzgerald:2005:VDO:1044891,Fitz:PGLarson:2009book} is a formal method which supports modeling and analysis of software systems at various levels of abstraction.
+A VDM specification, i.e.
+software specifications described in a VDM model, uses a combination of implicit and/or explicit definitions of functionalities to describe software specifications acquired from software requirements.
+VDM has a strong record of applications for design and specification of software systems in industry~\cite{Larsen:2007:RIA:2227886.2227894,DBLP:conf/fm/KuritaCN08,DBLP:journals/ijsi/KuritaN09}.
+The well-known basic benefits of using VDM to describe software specifications are from the accuracy and unambiguity of VDM, which is common for other formal methods like B.
+
+% Validation methods of VDM
+To validate a VDM model, existing tools such as Overture~\cite{Larsen:2010:OII:1668862.1668864} and VDMTools provide static consistency check regarding the syntax and type constraints of VDM specifications.
+The semantic validation of VDM specifications are rely on proof obligations (POs)~\cite{AL:97:POGV} generated by the above tools, and theorem provers are applied to discharge the generated POs.
+Testing (specification animation) by running a VDM model with an interpreter~\cite{Prehn:1991:LNCS551} is the alternative way of validation.
+This requires that the VDM model to be specified explicitly so that an interpreter can produce specific values.
+
+% More about proof obligations and the motivation: solve as many as possible POs
+
+In this paper, we propose an alternative approach to discharging the POs of VDM specifications using SMT solvers.
+Instead of translating the whole VDM model, we encode each POs whith its context information such type constraints involved, then release the encoded formulas to prove the PO.
+The advantages of proposed approach are (1) the encoding involves only a few segments of a VDM specification; (2) the proof process is automated with SMT solvers; (3) If a PO's proof fails, a counterexample model is returned for further examination.
+
+More specifically, in this approach we encode and prove POs of VDM-SL models as solving SMT formulas with Python API of Z3~\cite{MB:08:ZSS}.
+Z3 is one of the popular SMT solvers widely used in software verification and its Python API provides flexibility in building SMT formulas than SMT-LIB.
+We have conducted some case studies the results showed that our approach can discharge the majority of POs of a VDM-SL model with efficiency.
+
+The structure of this paper is as follows: blah blah blah.
+
+Main story: To validate VDM models, discharging POs is the major way.
+However, the POs of a VDM model may rise to hundreds [give an example LOC of model vs.
+number of POs].
+Because of undecidability of VDM POs, discharging all POs is very difficult.
+Our approach aims to release POs as many as possible by introducing SMT solvers.
+We chose Z3 because it is one of the most popular SMT solver.
+The Python API is handy for flexible encoding.
+The results are good.

### 2017-05-19 paper/obligations.tex
intro, vdm

Objective: explain the semantics of Overture generated POs.

Explain what is proof obligation in VDM. There are kinds of POs: domain, subtype, satisfiability, and termination. Any example?

Explain POs generated by Overture tool. Any example?

### 2017-05-19 paper/related.tex
intro, vdm

(1) work about discharging POs using theorem provers. (2) other work about validation/verfication of VDM models rather than discharging POs.

For example, \cite{Verm:2007:master,Vermolen:2010:PCV:1774088.1774608} translate VDM models to HOL and then prove the POs within the environment of HOL. However, this approach has several issues: (1) the translation is applied to the whole VDM model so that it may be tedious and error-prone; (2) The proof process is usually complicated and requires an expert; (3) It is not clear where the problem is when the proof fails.

\cite{AL:97:POGV,CFP:14:TVCSAP,CT:15:EOCGTIS}

### 2017-05-19 paper/studies.tex
intro, vdm

@@
 To encode the negated formula of PO14, similar technique for  encoding {\tt set (dom status)} for PO1 can be applied.
-For the range restriction function, since the range is a single set with value {\tt <WR>}, there is no need of applying set notations but limiting the map value {\tt status(r)} to {\tt <WR>}.
+For the range restriction function, since the range is a singleton set with element value of {\tt <WR>}, we can skip encoding set notations while limiting the map value {\tt status(r)} to {\tt <WR>}.
 Thus, {\tt r in set (dom (status :> \{<WR>\}))} is encoded as follows.
@@
 
-For the map inverse function, we encoded {\tt m1 = \{(inverse calls)(r) |-> <SI>\}} as multiple conditions: (1) the part of map, where key {\tt i} is used and inverse is applied, should be defined; (2) the inversed domain, i.e., the range of original map, is {\tt r}; (3) then {\tt m1(i)} is {\tt <SI>}; (4) for all other indexes, {\tt m1} should be undefined.
+For the map inverse function, we chose not to define a new map as the inverse of {\tt calls}, but encoded {\tt m1 = \{(inverse calls)(r) |-> <SI>\}} as a case of {\tt m1 in set \{\{r |-> <SR>\}, \{(inverse calls)(r) |-> <SI>\}\}}.
+The case was encoded as the collection of several conditions.
 
@@
          Status.INITIATOR(Initiator.SI)),
+
 ForAll(k, 
@@
 
-The result of checking PO14 ware expected to be {\tt unsat} but we got {\tt sat} and a model that satisfies the formula and constraints.
+In above code, the first two conditions promise that (1) the part of map, where key {\tt i} is used and inverse is applied, has to be defined; (2) the domain of the inversed map, i.e., the range of the original map, is {\tt r}.
+Then {\tt m1(i)} where key {\tt i} is referred should be {\tt <SI>}.
+Finally the last condition guarantees that {\tt m1} should be undefined for all indexes except {\tt i}.
+
+Recall that {\tt calls} is defined as {\tt inmap}, a one-to-one map.
+Thus, we had to add  the conditions about {\tt inmap} as context information: for any two different users, either both users calls no one (undefined) or they are calling different users.
+
+\begin{mdframed}[roundcorner=5pt,shadow=true]
+\begin{Verbatim}[fontsize=\small]
+k  = Const('k',Subscriber)
+l  = Const('l',Subscriber)
+ForAll([k,l],
+  If(
+    k==l,
+    calls(k)==calls(l),
+    Or(
+      And(
+        calls(k)==Subscriber_lift.NDF,
+        calls(l)==Subscriber_lift.NDF
+      ),
+      calls(k)!=calls(l)
+    )
+  )
+)
+\end{Verbatim}
+\end{mdframed}
+
+The result of checking PO14 (negated) ware expected to be {\tt unsat}, but we got {\tt sat}.
+Z3 returned a model that satisfies the negated formula and constraints so that we could examine where is the problem.
+We found that {\tt calls} in the returned model is a map where users calls themselves.
+This does not violate the type of {\tt inmap}, but an user cannot call him/her-self in a telephone exchange system.
+Therefore, the checking of PO14 suggested that the VDM-SL model should add a constraint on {\tt calls}.
+As an example of correction, we may add the constraint below as the type invariant of {\tt calls} or the additional state invariant\footnote{We have added the constraint and checked again, and the PO was proved successfully.}.
+
+\begin{mdframed}[roundcorner=5pt]
+\begin{Verbatim}[fontsize=\small]
+forall i in set dom calls & calls(i) != i
+\end{Verbatim}
+\end{mdframed}
+
+In this case study, we did not check all the 27 POs since the encoding could not cover all POs.
+The main issue is that the maps are encoded as uninterpreted functions so that {\tt forall} quantifier cannot be applied.
+This results that we could not check the postcondition satisfiable obligations of this model\footnote{Besides postcondition satisfiable obligations, there are two POs with type check function {\tt is\_(name, type)}, for which Z3 encoding is not available and then skipped.}.
+Finally, we checked 16 out of all 27 POs.
 
@@
 
-Overture generated 27 POs for this model.
+% Overture generated 27 POs for this model.
 These POs can be divided into several groups as showed below.
@@
 
-\begin{enumerate}
-\item
-Legal map application obligations, invariant satisfiable obligations: 1 to 7.
-\item
-Enumeration map injectivity obligations: 9, 14, 18, 21, 25.
-\item
-Legal map application obligations with inverse map: 12, 17, 20, 23.
-\item
-Operation postcondition satisfiable obligations: 8, 10, 11, 15, 16, 19, 22, 26, 27
-\item
-Map inverse obligations: 13, 24.
-\end{enumerate}
+% \begin{enumerate}
+% \item
+% Legal map application obligations, invariant satisfiable obligations: 1 to 7.
+% \item
+% Enumeration map injectivity obligations: 9, 14, 18, 21, 25.
+% \item
+% Legal map application obligations with inverse map: 12, 17, 20, 23.
+% \item
+% Operation postcondition satisfiable obligations: 8, 10, 11, 15, 16, 19, 22, 26, 27
+% \item
+% Map inverse obligations: 13, 24.
+% \end{enumerate}
+
+Points of discussion:
+(1) Automation: Is the encoding can be automated? partly or fully? 
+(2) Guidance and patterns: What patterns are repeated used in encoding? Any guide of encoding?

### 2017-05-19 paper/vdm.tex
intro, vdm

% Give a simple example with the syntax used in case studies and explain implicit and explicit specification in VDM-SL.

% General information of VDM
The Vienna Development Method (VDM) was originally developed in the 1970's at the IBM laboratories in Vienna~\cite{DBLP:conf/fm/1978}. The VDM Specification Language (VDM-SL) is a higher-order language with formally defined syntax and semantics~\cite{Prehn:1991:LNCS551,Larsen1995585}. VDM provides various abstract data types: basic types such as booleans, natural numbers, and tokens; advanced types such as record, product, set, and map. Typed variables (state) may be restricted by invariants and operations/functions may be specified with preconditions and postconditions.

% Explaining the sample VDM-SL model: part1
As a brief explanation of VDM-SL, Fig.~\ref{fig:module_sample} shows a sample VDM-SL model {\tt CMDS}. The specification is divided into several blocks: types (line 1-4), state (line 6-12), operations (line 14-18), and functions (line 20-27). {\tt CMD} is a union type of quotes which is usually used for enumeration. {\tt CMD} represents commands {\tt <R>} and {\tt <L>} which may be recognized as right and left. {\tt CMD\_series} is defined as a sequence type which is a list of commands. Note that the type of elements {\tt [CMD]} with square brackets means the union type {\tt POS|nil}. We also defined a map type for counting the occurrences of commands. The state of the model is defined as a sequence of commands {\tt CMD\_series} and its initial value is defined as empty. Its invariant specifies that adjacent elements of {\tt commands} have to be different. The invariant is specified with {\tt forall} quantifier and {\tt len} operation that returns the length of a sequence.

% An example of VDM-SL model (or some segments of VDM-SL code) should have
% types: quote and union, sequence, token, map
% operations of sequence: len, hd, tl
% operations of map: range restriction
% operation of set: in set
% others: state def., forall statement
\begin{figure}[t]
\begin{center}
\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\small,numbers=left]
types
CMD = <R> | <L>;
CMD_series = seq of [CMD];
CMD_times = map CMD to nat;

state S of
  commands : CMD_series
  inv s == forall k in set 
    {1,...,len s.commands - 1} &
    s.commands(k) <> s.commands(k+1)
  init p == p = mk_S([])
end

operations
  push_cmd(a:[CMD])
  pre commands = [] or hd commands <> a
  post hd commands = a and
       tl commands = commands~;

functions
  times_count : CMD_series -> CMD_times
  times_count(a) == {
    <R> |->
    len [ i | i in set inds a & a(i)=<R> ],
    <L> |->
    len [ i | i in set inds a & a(i)=<L> ]
  };
\end{Verbatim}
\end{mdframed}
\caption{A sample VDM-SL specification: CMDS}
\label{fig:module_sample}
\vspace{-25pt}
\end{center}
\end{figure}

% Explaining the sample VDM-SL model: part2
The operation {\tt push\_cmd} is specified to accept a command {\tt CMD} or {\tt nil} and pushes the command as the head of {\tt commands}. The postcondition is specified using the old state {\tt commands\textasciitilde} with postfixed tilde and says that the old state is the tail of the new state {\tt commands}. Note that {\tt push\_cmd} is an implicit operation that only specifies the signature with pre- and post-conditions. The function that counts the occurrences of commands is specified explicitly. Instead of using loop, the function uses sequence comprehension to construct the map of type {\tt CMD\_times}. This also shows the expression power of VDM in specification.

% tools and dialects of VDM (optional)
Existing tools such as Overture Tool~\cite{Larsen:2010:OII:1668862.1668864} and VDMTools~\cite{2008:VAS:1361213.1361214} provide graphical user interfaces for easy editing/building of VDM models. These tools also provide functionalities such as type/syntax checking and testing/animation (execution by the interpreter~\cite{Prehn:1991:LNCS551}) for validation and verification of VDM models. There are other dialects of VDM: VDM++ and VDM Real-Time (VDM-RT)~\cite{10.1007/11813040_11}. VDM++ is the extension of VDM-SL with object-orient concepts; VDM-RT further extends VDM++ with scheduling controls of threads or processes.

### 2017-05-23 paper/introduction.tex
intro, vdm, obligation, strategy(new)

@@
 Testing (specification animation) by running a VDM model with an interpreter~\cite{Prehn:1991:LNCS551} is the alternative way of validation.
-This requires that the VDM model to be specified explicitly so that an interpreter can produce specific values.
+This requires that the VDM model to be specified explicitly so that an interpreter can produce specific values for functions and operations.
 
 % More about proof obligations and the motivation: solve as many as possible POs
-
-In this paper, we propose an alternative approach to discharging the POs of VDM specifications using SMT solvers.
-Instead of translating the whole VDM model, we encode each POs whith its context information such type constraints involved, then release the encoded formulas to prove the PO.
-The advantages of proposed approach are (1) the encoding involves only a few segments of a VDM specification; (2) the proof process is automated with SMT solvers; (3) If a PO's proof fails, a counterexample model is returned for further examination.
+Discharging POs of VDM models is essential to guarantee the consistency of the formal specifications.
+For a VDM model, the number of POs is usually large compare to the size of the model.
+For example, the sample VDM-SL model shown in Fig.~\ref{fig:module_sample} has less than thirty lines, and Overture generates nine POs for the model.
+In this paper, we propose an approach to discharging the POs of VDM specifications using SMT solvers.
+Here we focus on VDM-SL models and the objective is to discharge POs of an VDM-SL model as many as possible through SMT solvers.
+In this approach, we encode each POs whith its context information such type and state constraints, then release the encoded formulas to prove the PO.
+The advantages of our approach are (1) the encoding involves only a few segments of a VDM specification; (2) the proof process is automated with SMT solvers; (3) If a PO's proof fails, a counterexample model is returned for further examination.
 
 More specifically, in this approach we encode and prove POs of VDM-SL models as solving SMT formulas with Python API of Z3~\cite{MB:08:ZSS}.
-Z3 is one of the popular SMT solvers widely used in software verification and its Python API provides flexibility in building SMT formulas than SMT-LIB.
-We have conducted some case studies the results showed that our approach can discharge the majority of POs of a VDM-SL model with efficiency.
+Z3 is one of the popular SMT solvers widely used in software verification and its Python API provides flexibility in building SMT formulas than using pure SMT-LIB~\cite{BarFT-SMTLIB}.
+We have conducted some case studies and the results showed that our approach can efficiently discharge the significant part of POs of a VDM-SL model.
+We chose the Overture tool as the PO generator in this paper\footnote{POs generated by Overture and VDMTools are slightly different in the numbers and styles of POs}.
 
-The structure of this paper is as follows: blah blah blah.
+The structure of this paper is as follows: Section~\ref{section:vdm} briefly introduce VDM with a sample VDM-SL example; Section~\ref{section:proof-obligations} describes the POs of VDM and explains the POs generated by the Overture tool; section~\ref{section:encode-strategy} describes our strategy of encoding VDM expressions with Python API of Z3; section~\ref{section:case-studies} demonstrates case studies with discussions; section~\ref{section:related-work} compares our approach to related work; section~\ref{section:conclusions} concludes this paper and points out the future work.
 
-Main story: To validate VDM models, discharging POs is the major way.
+% Main story: To validate VDM models, discharging POs is the major way.
 However, the POs of a VDM model may rise to hundreds [give an example LOC of model vs.

### 2017-05-23 paper/obligations.tex
intro, vdm, obligation, strategy(new)

@@
-Objective: explain the semantics of Overture generated POs.
+% Objective: explain the semantics of Overture generated POs.
+% Explain what is proof obligation in VDM. There are kinds of POs: domain, subtype, satisfiability, and termination. Any example?
+% Explain POs generated by Overture tool. Any example?
 
-Explain what is proof obligation in VDM. There are kinds of POs: domain, subtype, satisfiability, and termination. Any example?
+% Generation information of POs
+For a VDM model, a proof obligation is a statement that must be proved in order to ensure the consistency of the model.
+A PO contains a predicate with its context information as shown below.
+The context information are code segments of the VDM model that relate to the predicate to be proved.
 
-Explain POs generated by Overture tool. Any example?
+\begin{mdframed}[roundcorner=5pt]
+\begin{Verbatim}[fontsize=\small]
+PO: context information ==> predicate
+\end{Verbatim}
+\end{mdframed}
+
+% Classifications of POs and sample POs from sample VDM-SL model
+POs of VDM-SL can be classified type compatibility (subtype checking), domain checking, and satisfiability~\cite{AL:97:POGV,Vermolen:2010:PCV:1774088.1774608}.
+Type compatibility relates to types with invariant or subtypes; domain checking relates to the use of partial functions and partial operators; satisfiability relates postconditions of functions and operations.
+Both Overture and VDMTools generate POs with the syntax of VDM.
+
+For the sample VDM-SL model shown in Fig.~\ref{fig:module_sample}, Overture generated nine POs.
+Here selected POs are demonstrated and explained.
+PO1 is called ``legal sequence application obligation'' that relates to the sequence involved expression (line 9-10) of the VDM-SL model {\tt CMDS}).
+PO1 says that for all k in the set from 1 to {\tt len s.commands -1}, k should be always within the bound of defined indexes of the sequence {\tt s.commands}.
+Also, it can be recognized that if we use {\tt len s.commands} instead of {\tt len s.commands -1} the proof will fail.
+Note that in VDM, a sequence are indexed from 1 to its length.
+Also note that in PO1, the state {\tt CMDS'S} is mentioned, and we have to include the state invariant as well though it is not explicitly described in PO1.
+
+\begin{mdframed}[roundcorner=5pt]
+\begin{Verbatim}[fontsize=\small]
+(forall s:CMDS`S & 
+  (forall k in set 
+    {1, ... ,((len (s.commands)) - 1)} & 
+    (k in set (inds (s.commands)))))
+\end{Verbatim}
+\end{mdframed}
+
+PO6 is an operation postcondition satisfiable obligation for {\tt push\_cmd} (line 15-18).
+PO6 says that for every input {\tt a:CMD} and {\tt oldstate}, the postcondition which takes inputs of {\tt a}, {\tt oldstate}, and {\tt newstate} should be satisfied.
+Note that in PO6, instead of the postcondition and state in the VDM-SL model, special names {\tt oldstate}, {\tt newstate}, and {\tt post\_push\_cmd} are used.
+In VDM-SL, the {\tt post\_} prefix is an evaluable function if given appropriate inputs.
+In a satisfiable obligation, the name of state (in this case {\tt commands}) represents {\tt newstate} while the same name postfixed with ``~\textasciitilde~'' (in this case {\tt commands\textasciitilde}) represents {\tt oldstate}.
+
+\begin{mdframed}[roundcorner=5pt]
+\begin{Verbatim}[fontsize=\small]
+(forall a:[CMD], oldstate:CMDS`S &
+  (exists newstate:CMDS`S &
+    post_push_cmd(a, oldstate, newstate)))
+\end{Verbatim}
+\end{mdframed}

### 2017-05-23 paper/strategy.tex
intro, vdm, obligation, strategy(new)

% Explain the encode strategy: (1) explain Z3 and python API; (2) strategy on types; (3) take negation or not

% Z3 introduction, API introduction
As mentioned in Section~\ref{section:introduction}, our approach encodes and prove VDM POs with the Z3 SMT solver~\cite{MB:08:ZSS}. Z3 is one of the popular SMT solver wildly used in verification of programs and software systems. Z3 has APIs for major programming languages such as C, Java, and Python. The APIs provide a better way of constructing SMT formulas than SMT-LIB~\cite{BarFT-SMTLIB} because one can adopt the flow control characteristics like loop and if-then-else of programming languages to build and solve formulas in a flexible and smart way than using pure SMT-LIB format. In our approach, we chose Z3's Python API (Z3py) as the encoding environment.

% Issues of encoding
VDM has rich types such as natural number, quote, sequence, and mapping and there are no direct matching to types in Z3. For example, to encode a natural number, we need to declare an integer and then add a constraint of greater and equal to zero on it. The enumeration in Z3 can be used to encode simple quote types. For complex types, we may use uninterpreted functions or arrays to capture the characteristic and operators of the type we want to encode. There is not yet a systematic way of encoding VDM types to Z3 so we have to choose the encoding carefully based on the types and expressions used in a VDM-SL model. However, it is expected that we may find some patterns of encoding for frequent seen styles in VDM-SL models.

% Demonstrate Z3Py encoding
For more details, the Z3py encoding firstly needs to include the Z3 module and declare a solver object. We can add constraints of a PO to the solver object and solve the formulas for proof.

\begin{mdframed}[roundcorner=5pt,shadow=true]
\begin{Verbatim}[fontsize=\small]
  from z3 import *
  s = Solver()
  i = Int('i')
  s.add(i>=0)
\end{Verbatim}
\end{mdframed}

The above Python code imports the z3 module and create a solver object, then declare an integer {\tt i} with a constraint that {\tt i} is greater and equal to zero. In this case, {\tt i} is the encoding an instance of type {\tt nat} in VDM. Note that we cannot define a user-defined type in Z3 but only can declare every instance as Z3 type and add necessary constraints to achieve the encoding of variables of VDM types. This means that if there are several variables declared as VDM type that Z3 does not have, instances of the same number of VDM variables have to be declared in Z3 with constraints for each instance added to the solver object. In Section~\ref{section:case-studies}, we will show more details of the encoding through the explanation of case studies.

### 2017-05-23 paper/studies.tex
intro, vdm, obligation, strategy(new)

@@
 \hline
-PO\#	&	negated?	&	result	&	time (sec.) \\ \hline
+PO\#	&	negated	&	result	&	time (sec.) \\ \hline
 1		&	Y		&	sat		&	0.031 \\ \hline
@@
 \end{mdframed}
+\medskip
 

### 2017-05-24 paper/introduction.tex
polish introduction

@@
 The Vienna Development Method (VDM)~\cite{Jones:1990:SSD:94062,Fitzgerald:2005:VDO:1044891,Fitz:PGLarson:2009book} is a formal method which supports modeling and analysis of software systems at various levels of abstraction.
-A VDM specification, i.e.
-software specifications described in a VDM model, uses a combination of implicit and/or explicit definitions of functionalities to describe software specifications acquired from software requirements.
+A VDM specification can be explicit, implicit, or both.
+An explicit specification is operational; it describes \emph{how} system components work.
+An implicit specificiation on the other hand is declarative; it describes \emph{what} system components should work.
+Software specifications in VDM models define functionalities explicitly or implicitly according to software requirements.
 VDM has a strong record of applications for design and specification of software systems in industry~\cite{Larsen:2007:RIA:2227886.2227894,DBLP:conf/fm/KuritaCN08,DBLP:journals/ijsi/KuritaN09}.
@@
 To validate a VDM model, existing tools such as Overture~\cite{Larsen:2010:OII:1668862.1668864} and VDMTools provide static consistency check regarding the syntax and type constraints of VDM specifications.
-The semantic validation of VDM specifications are rely on proof obligations (POs)~\cite{AL:97:POGV} generated by the above tools, and theorem provers are applied to discharge the generated POs.
+The semantic validation of VDM specifications are established by discharging proof obligations (POs)~\cite{AL:97:POGV} generated by the above tools.
+POs are logic formulas characterizing consistency of VDM specifications.
+Theorem provers have been used to discharge the generated POs.
 Testing (specification animation) by running a VDM model with an interpreter~\cite{Prehn:1991:LNCS551} is the alternative way of validation.
@@
 For example, the sample VDM-SL model shown in Fig.~\ref{fig:module_sample} has less than thirty lines, and Overture generates nine POs for the model.
-In this paper, we propose an approach to discharging the POs of VDM specifications using SMT solvers.
+Consequently, it is not uncommon to have hundreds of POs for large VDM models.
+All of them have to be discharged to guarantee consistency of software specifications.
+Manually discharging so many POs with theorem provers however can be a daunting task in practice.
+Furthermore, both errorneous VDM specifications and bad proof strategies induce un-dischargeable POs.
+Theorem provers provide little information when they fail to discharge POs.
+It is unclear whether a VDM specification is flawed or a better proof strategy is possible when POs cannot be discharged by theorem provers.
+
+% SMT
+Satisfiability Modulo Theories (SMT) solvers extend SAT solvers by incorporating decision procedures for first-order theories such as linear arithmetic.
+Given a quantifier-free first-order logic formula in mixed theories, SMT solvers will try to find a model satisfying the given formula or report that the formula is unsatisfiable.
+Thanks to their efficiency, SMT solvers have found a wide range of applications from software verification to software testing.
+Recent SMT solvers can even check satisfiability of general first-order formulas.
+It is therefore interesting to see if one can take advantages of recent developments of SMT solvers to discharge first-order POs in VDM models.
+
+%Of course, the complexity of deciding satisfiability of general first-order formulas increases drastically even for decidable theories.
+Modern SMT solvers have to employ sophisticated heuristics to improve efficiency in practice.
+
+In this paper, we apply SMT solvers to discharge POs of VDM specifications.
 Here we focus on VDM-SL models and the objective is to discharge POs of an VDM-SL model as many as possible through SMT solvers.
-In this approach, we encode each POs whith its context information such type and state constraints, then release the encoded formulas to prove the PO.
+In our approach, we encode each POs with its context information such as type and state constraints as a first-order formula, and then prove the PO by sending the formula to SMT solvers.
 The advantages of our approach are (1) the encoding involves only a few segments of a VDM specification; (2) the proof process is automated with SMT solvers; (3) If a PO's proof fails, a counterexample model is returned for further examination.
 
-More specifically, in this approach we encode and prove POs of VDM-SL models as solving SMT formulas with Python API of Z3~\cite{MB:08:ZSS}.
-Z3 is one of the popular SMT solvers widely used in software verification and its Python API provides flexibility in building SMT formulas than using pure SMT-LIB~\cite{BarFT-SMTLIB}.
-We have conducted some case studies and the results showed that our approach can efficiently discharge the significant part of POs of a VDM-SL model.
+More specifically, in this approach we encode and prove POs of VDM-SL models by solving SMT formulas with Python API of Z3~\cite{MB:08:ZSS}.
+Z3 is a popular SMT solver widely used in software verification and its Python API provides flexibility in constructing SMT formulas~\cite{BarFT-SMTLIB}.
+In order to encode POs, built-in VDM types such as sequences and maps must be characterized by theories in Z3.
+We conducted case studies and the results showed that our approach can efficiently discharge the significant part of POs of a VDM-SL model.
 We chose the Overture tool as the PO generator in this paper\footnote{POs generated by Overture and VDMTools are slightly different in the numbers and styles of POs}.

### 2017-05-24 paper/vdm.tex
intro, vdm, obligation, strategy(new)
polish vdm

@@
 VDM provides various abstract data types: basic types such as booleans, natural numbers, and tokens; advanced types such as record, product, set, and map.
-Typed variables (state) may be restricted by invariants and operations/functions may be specified with preconditions and postconditions.
+A state consists of typed variables.
+Typed variables in a state may be restricted by invariants and operations/functions may be specified with preconditions and postconditions.
 
@@
 As a brief explanation of VDM-SL, Fig.~\ref{fig:module_sample} shows a sample VDM-SL model {\tt CMDS}.
-The specification is divided into several blocks: types (line 1-4), state (line 6-12), operations (line 14-18), and functions (line 20-27).
+The specification is divided into several blocks: types (lines 1-4), state (lines 6-12), operations (lines 14-18), and functions (lines 20-27).
 {\tt CMD} is a union type of quotes which is usually used for enumeration.
-{\tt CMD} represents commands {\tt <R>} and {\tt <L>} which may be recognized as right and left.
+{\tt CMD} represents commands {\tt <R>} and {\tt <L>} denoting right and left.
 {\tt CMD\_series} is defined as a sequence type which is a list of commands.
-Note that the type of elements {\tt [CMD]} with square brackets means the union type {\tt POS|nil}.
+Note that the type of elements {\tt [CMD]} with square brackets means the union type {\tt CMD|nil}.
 We also defined a map type for counting the occurrences of commands.
-The state of the model is defined as a sequence of commands {\tt CMD\_series} and its initial value is defined as empty.
-Its invariant specifies that adjacent elements of {\tt commands} have to be different.
-The invariant is specified with {\tt forall} quantifier and {\tt len} operation that returns the length of a sequence.
+The state of the model is defined as a sequence of commands {\tt CMD\_series}.
+The state invariant specifies that adjacent elements of {\tt commands} have to be different.
+It is specified with {\tt forall} quantifier and {\tt len} operation that returns the length of a sequence.
+Finally, the initial value of \texttt{commands} is defined to be the empty sequence.
 
@@
 \end{mdframed}
+\vspace{-10pt}
 \caption{A sample VDM-SL specification: CMDS}
 \label{fig:module_sample}
-\vspace{-25pt}
+% \vspace{-25pt}
 \end{center}
@@
 The operation {\tt push\_cmd} is specified to accept a command {\tt CMD} or {\tt nil} and pushes the command as the head of {\tt commands}.
+The precondition demands that \texttt{command} is the empty sequence or its head is different from the argument.
 The postcondition is specified using the old state {\tt commands\textasciitilde} with postfixed tilde and says that the old state is the tail of the new state {\tt commands}.
-Note that {\tt push\_cmd} is an implicit operation that only specifies the signature with pre- and post-conditions.
-The function that counts the occurrences of commands is specified explicitly.
+Observe that {\tt push\_cmd} is an implicit operation that only specifies the signature with pre- and post-conditions without any operation.
+The function \texttt{times\_count} that counts the occurrences of commands is specified explicitly.
 Instead of using loop, the function uses sequence comprehension to construct the map of type {\tt CMD\_times}.
+Specifically, the resultant map has the domain $\{ \texttt{<R>}, \texttt{<L>} \}$.
+It maps $\texttt{<R>}$ to the length of the sequence which contains the indices of $\texttt{a}$ having the value $\texttt{<R>}$.
+That is, the resultant map gives the number of $\texttt{<R>}$ in the sequence $\texttt{a}$ on $\texttt{<R>}$.
+Simiarly, it gives the number of $\texttt{<L>}$ in $\texttt{a}$ on $\texttt{<L>}$.
 This also shows the expression power of VDM in specification.

### 2017-05-25 paper/obligations.tex
corrected postcondition obligation

@@
 
-For the sample VDM-SL model shown in Fig.~\ref{fig:module_sample}, Overture generated nine POs.
+For the sample VDM-SL model shown in Fig.~\ref{fig:module_sample}, Overture generated ten POs.
 Here selected POs are demonstrated and explained.
@@
 
-PO6 is an operation postcondition satisfiable obligation for {\tt push\_cmd} (line 15-18).
-PO6 says that for every input {\tt a:CMD} and {\tt oldstate}, the postcondition which takes inputs of {\tt a}, {\tt oldstate}, and {\tt newstate} should be satisfied.
-Note that in PO6, instead of the postcondition and state in the VDM-SL model, special names {\tt oldstate}, {\tt newstate}, and {\tt post\_push\_cmd} are used.
-In VDM-SL, the {\tt post\_} prefix is an evaluable function if given appropriate inputs.
-In a satisfiable obligation, the name of state (in this case {\tt commands}) represents {\tt newstate} while the same name postfixed with ``~\textasciitilde~'' (in this case {\tt commands\textasciitilde}) represents {\tt oldstate}.
+PO7 is an operation postcondition satisfiable obligation for {\tt push\_cmd} (line 15-18).
+PO7 says that for every input {\tt a:CMD} and {\tt oldstate}, if the precondition whch takes inputs {\tt a} and {\tt oldstate} is satisfied, the postcondition which takes inputs {\tt a}, {\tt oldstate}, and {\tt newstate} should be satisfied.
+Note that instead of the specification of precondition, postcondition, and state in {\tt CMDS} module, special names {\tt oldstate}, {\tt newstate}, and {\tt post\_push\_cmd} are used.
+In VDM-SL, the {\tt pre\_} and {\tt post\_} prefix represent the precondition and postcondition of a function/operator separately.
+They are evaluable functions if given appropriate inputs, and return a boolean value.
+In a satisfiable obligation of an operation in which the state change is usually expected, the name of state (in this case {\tt commands}) represents {\tt newstate} while the same name postfixed with ``~\textasciitilde~'' (in this case {\tt commands\textasciitilde}) represents {\tt oldstate}.
 
@@
 (forall a:[CMD], oldstate:CMDS`S & 
+  (pre_push_cmd(a, oldstate) => 
     (exists newstate:CMDS`S & 
-    post_push_cmd(a, oldstate, newstate)))
+      post_push_cmd(a, oldstate, newstate))
+))
 \end{Verbatim}

### 2017-05-25 paper/strategy.tex
complete encode strategy

@@
 Z3 has APIs for major programming languages such as C, Java, and Python.
-The APIs provide a better way of constructing SMT formulas than SMT-LIB~\cite{BarFT-SMTLIB} because one can adopt the flow control characteristics like loop and if-then-else of programming languages to build and solve formulas in a flexible and smart way than using pure SMT-LIB format.
+The APIs provide a better way of constructing SMT formulas than SMT-LIB~\cite{BarFT-SMTLIB} because one can adopt the flow control characteristics like loop and if-then-else of programming languages to build and solve formulas in a flexible and smart way.
 In our approach, we chose Z3's Python API (Z3py) as the encoding environment.
 
-% Issues of encoding
-VDM has rich types such as natural number, quote, sequence, and mapping and there are no direct matching to types in Z3.
-For example, to encode a natural number, we need to declare an integer and then add a constraint of greater and equal to zero on it.
-The enumeration in Z3 can be used to encode simple quote types.
-For complex types, we may use uninterpreted functions or arrays to capture the characteristic and operators of the type we want to encode.
-There is not yet a systematic way of encoding VDM types to Z3 so we have to choose the encoding carefully based on the types and expressions used in a VDM-SL model.
-However, it is expected that we may find some patterns of encoding for frequent seen styles in VDM-SL models.
+% steps of encoding
+% 1. check context information: types, functions, and operations involved in the obligation
+% 2. encode context information: types and type invariants, functions, and operations (pre- and post-conditions if needed)
+% 3. negation and forall quantifier elimination
+Currently, the encoding is manually performed based on the obligation to be discharged and below are the steps of encoding and solving an obligation with SMT solvers.
+
+\begin{enumerate}
+\item
+Determine the context information of the PO: The context information is the definitions specified in the VDM model that are involved in the obligation.
+The context information may include types, functions, and operations.
+Note that if a type is involved, its type invariant should be included, too.
+\item
+Encode context information: this step encodes the context information determined by the first step, and can be divided into several minor steps.
+\begin{enumerate}
+\item
+For a VDM type in the context information, both the type itself and its invariant have to be encoded.
+\item
+For pre- and post-conditions of a function/operation specified in the VDM model, the encoding is needed if {\tt pre\_<name>} or {\tt post\_<name>} are used in the PO.
+For example, the postcondition satisfiable obligation (PO7) of the {\tt CMDS} module described in Section~\ref{section:proof-obligations} used both precondition and postcondition.
+\item
+If a PO only considers a part of the specification of a type, function, or an operations, it is only needed to encode the expressions that are directly used in the PO.
+In other words, if a type, a function, or an operation is not directly used such as variables of the type, evaluation using the function or the operation, there is no need to encode the full specification of the type, the function, or the operation.
+\end{enumerate}
+\item
+Quantifier elimination and obligation negation
+\begin{enumerate}
+\item
+Obligation negation: In model checking with SMT solvers, the assertion to be checked is usually negated, and the result of unsat is expected and means the proof of the assertion succeeds.
+For the result of sat which means the assertion can be violated, we can get a model as the counterexample that shows the information of how the assertion is violated.
+In this work, we also take negation on the PO to be discharged.
+However, the elimination of forall quantifier should be considered together when applying the negation to the PO.
+\item
+Elimination of forall quantifiers: POs of VDM usually have forall quantifiers.
+To solve the formulas with forall quantifiers in Z3, though Z3 accepts formulas with forall quantifiers, there are limitations and we may need to eliminate forall quantifiers that may cause problem before encoding.
+In this work, forall quantifier usually causes difficulties in encoding and solving when complex types are involved.
+Obligation negation described above may achieve the elimination of troublesome foall quantifiers.
+However, obligation negation should be applied carefully not to introduce troublesome forall quantifiers from exists quantifiers.
+\end{enumerate}
+\end{enumerate}
+
+% More details of encoding types, operators, and expressions
+VDM has rich types such as natural number, quote, sequence, and mapping, and there are no direct matching for all VDM types to the types in Z3.
+For example, to encode a natural number, we need to declare an integer and then add a constraint of greater and equal to zero because there is no natural number type in Z3.
+Also, the enumeration in Z3 can be used to encode simple quote types.
+For complex types like sequences, maps, or records, we may use uninterpreted functions or arrays to capture the characteristic and operators of the types we want to encode.
+At this point, there is not yet a complete and systematic way of encoding VDM types to Z3, what we can do is to choose the encoding carefully based on the types and expressions in a VDM model.
+However, based on our case studies, there are some patterns of encoding that are frequently used for corresponding expression styles in VDM-SL models.
 
@@
 This means that if there are several variables declared as VDM type that Z3 does not have, instances of the same number of VDM variables have to be declared in Z3 with constraints for each instance added to the solver object.
-In Section~\ref{section:case-studies}, we will show more details of the encoding through the explanation of case studies.
+The strategies discussed above will be demonstrated with details in Section~\ref{section:case-studies} through the report of case studies.
+
+% Z3py encoding: taking negation and forall quantifier elimination
+% Besides the encoding of VDM types, 
+%  For example, if a VDM type is encoded as an uninterpreted function, the type cannot be used within forall quantifier.
+A simple way of quantifier elimination is negation the formula so the forall quantifier turns into exists quantifier which can be omitted in SMT formulas usually.

### 2017-05-25 paper/studies.tex
studies: case1 completed

@@
-This section gives studies that apply our encoding to selected VDM-SL models and discusses the results. All the selected VDM-SL models are from the Overture repository\footnote{a url link here}.
+This section gives studies that apply our encoding to selected VDM-SL models and discusses the results.
+All the selected VDM-SL models can be found in Overture repository\footnote{http://overturetool.org/download/examples/VDMSL/}.
 
@@
 
-This model specifies an abstract pacemaker with some core functionalities of pulse trace selection\footnote{This model was developed by Sten Agerholm et al.
-in 1999 in connection with FM'99}.
-The data types used in the model are defined as follows:
-% The model is described in VDM-SL as a short, flat specification. This enables abstraction from design considerations and ensures maximum focus on high-level, precise and systematic analysis
+The first case is a small VDM-SL model specifies an abstract pacemaker.
+The model is not a full specification and focuses on part of definitions and operations related to heartbeat signals detected and recorded in a pacemaker.
+For this model of 58 lines, Overture generated nine POs.
+Here we report two the encoding and solving of obligations PO8 and PO3 for this case.
+
+\subsubsection{Proof Obligation 8}
+
+PO8 shown in Fig.~\ref{fig:po8_case1} is a legal sequence application obligation of operation {\tt Pace}.
+PO8 requires that the formulas specified with trace {\tt tr} of type {\tt Trace} be legal, i.e., computable to get its result.
+For more understanding of PO8, we may take a look on the definition of the operation {\tt Pace} as shown in Fig.~\ref{fig:pace_case1}.
+{\tt Pace} is an explicitly specified operation and takes three inputs of types {\tt Trace}, {\tt nat1}, and {\tt nat1}.
+It is easy to recognize that PO8 is checking about the sequence used in line 4 in Fig.~\ref{fig:pace_case1}.
+Note that to discharge PO8, we do not need the full definition of {\tt Pace}.
+
+\begin{figure}[t]
+\begin{center}
+\begin{mdframed}[roundcorner=5pt]
+\begin{Verbatim}[fontsize=\small]
+(forall tr:Trace, aperi:nat1, vdel:nat1,
+  oldstate & 
+    (forall i in set (inds (tl tr)) &
+      (((i mod aperi) = (vdel + 1)) =>
+        (i in set (inds tr)))
+))
+\end{Verbatim}
+\end{mdframed}
+\vspace{-10pt}
+\caption{Proof Obligation 8 of Abstract Pacemaker}
+\label{fig:po8_case1}
+\end{center}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\begin{mdframed}[roundcorner=5pt]
+\begin{Verbatim}[fontsize=\small]
+Pace: Trace * nat1 * nat1 ==> Trace
+Pace(tr,aperi,vdel) ==
+  return [nil] ^
+     [ if (i mod aperi = vdel + 1) and
+          tr(i) <> <V> 
+       then <V> else nil
+     | i in set inds tl tr];
+\end{Verbatim}
+\end{mdframed}
+\vspace{-10pt}
+\caption{Pace of Abstract Pacemaker}
+\label{fig:pace_case1}
+\end{center}
+\end{figure}
+
+As mentioned in Section~\ref{section:proof-obligations}, PO8 is the formula to be proved and we need to include the context information related to the obligation as the first step of encoding described in Section~\ref{section:encode-strategy}.
+To decide what to include as the context information, we firstly examine the types used in PO8 and we have {\tt nat1} and {\tt Pace}.
+{\tt nat1} is a type of natural numbers that excludes zero and thus starts from one.
+{\tt nat1} is a primitive type of VDM so we do not need more information about it.
+{\tt Pace} is a user-defined type so we have check its definition in the VDM-SL model.
+The type definition of {\tt Trace} in the model is as follows.
+Note that {\tt oldstate} bounded by the forall quantifier is used in the expression of PO8 and therefore is ignored in the encoding.
 
@@
   Event = <A> | <V>;
-
-  state Pacemaker of
-    aperiod : nat 
-    vdelay  : nat
-  init p == p = mk_Pacemaker(15,10)
-  end
 \end{Verbatim}
@@
 
-{\tt Trace} is of sequence type of {\tt Event}, where {\tt Event} is of quote type with two values {\tt <A>} and {\tt <V>}.
-The state is defined as two natural numbers with initial value of {\tt (15,10)}.
-In Z3, natural numbers is encoded as positive integers (integers with constraint of greater and equal to zero); quote type can be encoded as enumeration type {\tt EnumSort}; sequence type can be encoded as {\tt ArraySort}.
-The definitions of {\tt Event} and {\tt Trace} are as follows:
+Now we know that {\tt Trace} is of sequence type of {\tt Event}, where {\tt Event} is of quote type with two values {\tt <A>} and {\tt <V>} that are stand for the A-pace and V-pace of heartbeat signals.
+At this point, we have necessary context information we need to discharge PO8 and we may proceed to encoding step.
+For discharging PO3, we need to encode types used in PO8: {\tt nat1} and {\tt Trace}.
+{\tt nat1} is the natural number that excludes zero, so the encoding of {\tt nat1} is as follows.
+
+\begin{mdframed}[roundcorner=5pt,shadow=true]
+\begin{Verbatim}[fontsize=\small]
+  aperi = Int('aperi')
+  s.add(aperi>=1)
+  vdel = Int('vdel')
+  s.add(vdel>=1)
+\end{Verbatim}
+\end{mdframed}
+
+As the above code shows, we defined two instances of integer in Z3, {\tt aperi} and {\tt vdel}, then the constraint of greater and equal to one is added into the solver object for both instances.
+Note that in Z3, we cannot add constraints to types but instances of a Z3 type.
+This has been mentioned in Section~\ref{section:encode-strategy}.
+The remaining type not encoded yet is {\tt Trace}, the sequence type.
+We defined {\tt Event} firstly since {\tt Trace} is a trace of {\tt Event}.
+The code below shows the encoding of {\tt Event} and {\tt Trace}.
+Since {\tt Trace} is not exactly the sequence of {\tt Event} but {\tt [Event]} (a union with {\tt nil}, we did not define {\tt Event} but define {\tt [Event]}.
+Also, because VDM uses partial functions, a type can be undefined.
+This means that besides ordinary values as specification shows, we have to consider the case that the value is not defined.
+In our work, we used {\tt NDF} to represent the undefined value of {\tt Event}, and define {\tt Event\_lift} in Z3.
+The {\tt \_lift} indicates that we have considered the undefined value.
 
@@
 
-Because VDM uses partial functions, to encode as SMT formulas in Z3, we need to lift {\tt Event} to include the undefined case {\tt NDF}.
-Also, square brackets of {\tt [Evnet]} means the value can be {\tt nil}.
-Furthermore, since {\tt Trace} is defined as an array in Z3, we added constraints that limit an array as a sequence.
-Note that in Z3, the constraints are applied on each instance of type {\tt Trace} ({\tt tr} in the following formulas), not on the type itself.
+In the above encoding, {\tt Trace} is encoded as an array type in Z3 and this array type is by default indexed through the integer domain.
+An array type can not be a sequence type unless we add constraints regarding to sequences to limit an array to a sequence.
+The idea of adding constraints is the same as the encoding of {\tt nat1} but more complicate.
+Fig.~\ref{fig:tr_constraint_case1} shows the encoding of the constraint to {\tt tr}, the instance of {\tt Trace} as an array.
+Note again that the constraints are applied on {\tt tr}, the instance of {\tt Trace}, not the type {\tt Trace}.
+In Fig.~\ref{fig:tr_constraint_case1}, we added three constraints:
 
+\begin{figure}[t]
+\begin{center}
 \begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
+\begin{Verbatim}[fontsize=\small,numbers=left]
   tr = Const('tr',Trace)
@@
 \end{mdframed}
+\vspace{-10pt}
+\caption{Encode constraints for {\tt tr:Trace} in Z3py}
+\label{fig:tr_constraint_case1}
+\end{center}
+\end{figure}
 
-The first constraint says that the all indexes of {\tt tr} below 1 should be undefined because in VDM-SL, a sequence's index starts from 1.
-The second constraint says that if an index is defined, i.e., not undefined, all indexes lower than it should be all defined.
-The third constraint says tat if an index is undefined, all indexes higher than it should be all undefined.
-From the above three constraints, we limited an array as a sequence that starts from index 1 to its last element.
+\begin{itemize}
+\item
+The first constraint (line 4) says that the all indexes of {\tt tr} below 1 should be undefined because in VDM-SL, a sequence's index starts from 1.
+\item
+The second constraint (lines 6-13) says that if an index is defined, i.e., not undefined, all indexes lower than it should be all defined.
+\item
+The third constraint (line 15-22) says tat if an index is undefined, all indexes higher than it should be all undefined.
+\end{itemize}
 
-% \begin{itemize}
-% \item
-% A sequence's index is from 1, so indexes below 1 ($<=0$) should be all undefined.
-% $\forall i \in \mathcal{Z}, i<=0. ~tr[i] = NDF $
-% \item
-% If an index is defined in the sequence, all indexes lower then it should be all defined.
-% $\forall i \in \mathcal{Z}. ~tr[i] \neq NDF \rightarrow (\forall j \in \mathcal{Z}, 1<=j \land j <= i. ~tr[j] \neq NDF)$
-% \item
-% If an index is undefined, all indexes higher then it should be all undefined.
-% $\forall i \in \mathcal{Z}. ~(i >= 1 \land tr[i] = NDF) \rightarrow (\forall j \in \mathcal{Z}, j >= i. ~tr[j] = NDF)$
-% \end{itemize}
+From the above three constraints, we limited an array ({\tt tr}) to a sequence that starts from index 1 to its last element.
+Now we have encoded both types {\tt nat1} and {\tt Trace} used in PO8 of the VDM-SL model.
+However, before proceed to encode the obligation, it is necessary to encode the function {\tt len} that computes the length of a sequence.
+As shown in Fig.~\ref{fig:tr_len_case1}, we defined an uninterpreted function {\tt len\_tr} for {\tt tr}, the instance of {\tt Trace}.
+Line 1 is the type declaration of {\tt len\_tr} which is an uninterpreted function accepting the input of a {\tt Trace} type and returning an integer type.
+Similar to defining constraints for {\tt tr:Trace}, we gave {\tt len\_tr} the constraint that consists of two cases: the length of a sequence can only be equal or greater than 0.
 
-% where $NDF$ stands for undefined. The above formulas are added as constraints whenever the instance {\tt tr} of type {\tt Trace} is used.
-
-Since in the VDM-SL model, the length function of a sequence is used, we also need to define the length function in Z3.
-Here we defined an uninterpreted function {\tt len\_tr} for {\tt tr} the instance of {\tt Trace}.
-% $(len\_tr(tr) = 0 \land tr[1] = 0) \lor (len\_tr(tr) > 0 \land tr[len\_tr(tr)] \neq NDF \land tr[len\_tr(tr)+1] = NDF$
-
+\begin{figure}[t]
+\begin{center}
 \begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
+\begin{Verbatim}[fontsize=\small,numbers=left]
 len_tr=Function('len_tr',Trace,IntSort())
@@
 \end{mdframed}
-
-{\tt len\_tr} is defined as an uninterpreted function that takes {\tt tr} as input and returns an integer indicates the length of {\tt tr}.
-Since the length of a sequence can only be equal or greater than 0, the constraint of {\tt len\_tr} is divided into two cases:
+\vspace{-10pt}
+\caption{Encode {\tt len\_tr} for {\tt tr:Trace} in Z3py}
+\label{fig:tr_len_case1}
+\end{center}
+\end{figure}
 
@@
 \item
-If {\tt len\_tr} returns 0, then {\tt tr} is an sequence with no element. That is, {\tt tr[1]} has to be undefined so that all indexes of {\tt tr} are then undefined based on previous constraints.
+If {\tt len\_tr} returns 0, then {\tt tr} is an sequence with no element.
+That is, {\tt tr[1]} has to be undefined so that all indexes of {\tt tr} are then undefined based on previous constraints of {\tt tr}.
+(lines 4-7)
 \item
 If {\tt len\_tr} returns an integer greater than 0, then {\tt tr} has defined elements till its length index {\tt len\_tr(tr)} such that the element at and after index {\tt len\_tr(tr)+1} should be undefined.
+(line 8-12)
 \end{enumerate}
 
-We have defined the context information related to type {\tt Trace}.
-Now we can procceed to encode and prove the POs of the model.
-For this VDM-SL model, Overture generated nine POs.
-Most of the POs are similar so we only demonstrate PO8 and PO3.
+At this point, we have defined types and constraints needed for encoding PO8.
+According to the steps described in Section~\ref{section:encode-strategy}, we now proceed to the next step: obligation negation.
+The negated PO8 is shown in Fig.~\ref{fig:po8_negation_case1}.
+Compare to the original PO8 in Fig.~\ref{fig:po8_case1}, the negated PO8 also eliminated the forall quantifier so that we do not need to use forall in Z3py encoding.
+The PO can be recognized as the formula $\exists~[tr:Trace,aperi:nat1,vdel:nat1,i:nat] P \land Q \land \neg R$ where $P$, $Q$, and $\neg R$ can be encoded separately.
 
-$\bullet$ PO8 is a legal sequence application obligation of operation {\tt Pace} which is explicitly specified.
-This PO requires that the formulas specified with trace {\tt tr} be legal, i.e., computable to get its result.
-
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small]
-(forall tr:Trace, aperi:nat1, vdel:nat1,
-  oldstate & 
-    (forall i in set (inds (tl tr)) &
-      (((i mod aperi) = (vdel + 1)) =>
-        (i in set (inds tr)))
-))
-\end{Verbatim}
-\end{mdframed}
-
-In PO8 showed above, {\tt tl tr} is the tail of {\tt tr}, and {\tt inds} is the operator of gathering indexes of defined elements of {\tt tr} as the set of natural numbers.
-To prove PO8, firstly we took negation of PO8 which results a quantifier eliminated formula.
-
+\begin{figure}[t]
+\begin{center}
 \begin{mdframed}[roundcorner=5pt]
@@
 \end{mdframed}
+\vspace{-10pt}
+\caption{Obligation negation of PO8}
+\label{fig:po8_negation_case1}
+\end{center}
+\end{figure}
 
-Note that the {\tt oldstate} is not used in the operation so that we remove it in the above formula.
-To encode the PO in Z3, the PO can be recognized as the formula $\exists~[tr:Trace,aperi:nat1,vdel:nat1,i:nat] P \land Q \land \neg R$ where $P$, $Q$, and $\neg R$ can be encoded separately.
-
+\begin{figure}[t]
+\begin{center}
 \begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
+\begin{Verbatim}[fontsize=\small,numbers=left]
   And(aperi>=1,vdel>=1)        #nat1
@@
 \end{mdframed}
+\vspace{-10pt}
+\caption{Encoded formulas of PO8 in Z3py}
+\label{fig:po8_encoded_case1}
+\end{center}
+\end{figure}
 
-In the encoding of $P$, the tail related formula can be encoded as in the range of $[1,(len~tr)-1]$ so that it is not required to define {\tt tl} and {\tt inds} for sequences.
+Finally, we encoded the negated PO8 as the Z3py code shown in Fig.~\ref{fig:po8_encoded_case1}.
+The encoding has four constraints: two {\tt nat1} variables (line 1); encoding of $P$ ({\tt i in set (inds (tl tr)}) that is actually indexes from 1 to {\tt len tr -1} (line 2); encoding of $Q$ ({\tt (i mod aperi) = (vdel + 1)}) (line 3); encoding of $\neg R$ ({\tt not(i in set (inds tr))}) (line 4).
+Note that {\tt hd} and {\tt tl tr} are operators of the head and tail of a sequence, and {\tt inds} is the operator of gathering indexes of defined elements of a sequence and returns a set of natural numbers.
+Also note that in the encoding in Fig.~\ref{fig:po8_encoded_case1}, we did not explicitly encode {\tt hd}, {\tt tl}, and {\tt inds} and simplified the results of these operators as constraints related to ranges (upper and lower bound) of {\tt tr}.
+The encoded PO8 (negated) was checked by the solver object in Z3py for the satisfiability.
+The result was {\tt unsat}, which means that PO8 is proved.
+
+% With the above two cases of {\tt len\_tr} that defines the length function of traces, together with constraints of {\tt tr}, and instances of types {\tt nat1}, we encoded PO8 as shown in
+
+% The data types used in the model are defined as follows:
+
+% \begin{mdframed}[roundcorner=5pt]
+% \begin{Verbatim}[fontsize=\small]
+%   Trace = seq of [Event];
+%   Event = <A> | <V>;
+
+%   state Pacemaker of
+%     aperiod : nat 
+%     vdelay  : nat
+%   init p == p = mk_Pacemaker(15,10)
+%   end
+% \end{Verbatim}
+% \end{mdframed}
+
+% {\tt Trace} is of sequence type of {\tt Event}, where {\tt Event} is of quote type with two values {\tt <A>} and {\tt <V>}.
+The state is defined as two natural numbers with initial value of {\tt (15,10)}.
+In Z3, natural numbers is encoded as positive integers (integers with constraint of greater and equal to zero); quote type can be encoded as enumeration type {\tt EnumSort}; sequence type can be encoded as {\tt ArraySort}.
+The definitions of {\tt Event} and {\tt Trace} are as follows:
+
+% Furthermore, since {\tt Trace} is defined as an array in Z3, we added constraints that limit an array as a sequence.
+Note that in Z3, the constraints are applied on each instance of type {\tt Trace} ({\tt tr} in the following formulas), not on the type itself.
+
+% $(len\_tr(tr) = 0 \land tr[1] = 0) \lor (len\_tr(tr) > 0 \land tr[len\_tr(tr)] \neq NDF \land tr[len\_tr(tr)+1] = NDF$
+
+% {\tt len\_tr} is defined as an uninterpreted function that takes {\tt tr} as input and returns an integer indicates the length of {\tt tr}.
+Since the length of a sequence can only be equal or greater than 0, the constraint of {\tt len\_tr} is divided into two cases:
+
+% We have defined the context information related to type {\tt Trace}.
+Now we can procceed to encode and prove the POs of the model.
+For this VDM-SL model, Overture generated nine POs.
+Most of the POs are similar so we only demonstrate PO8 and PO3.
+
+% $\bullet$ PO8 is a legal sequence application obligation of operation {\tt Pace} which is explicitly specified.
+This PO requires that the formulas specified with trace {\tt tr} be legal, i.e., computable to get its result.
+
+% In PO8 showed above, {\tt tl tr} is the tail of {\tt tr}, and {\tt inds} is the operator of gathering indexes of defined elements of {\tt tr} as the set of natural numbers.
+To prove PO8, firstly we took negation of PO8 which results a quantifier eliminated formula.
+
+% Note that the {\tt oldstate} is not used in the operation so that we remove it in the above formula. 
+% To encode the PO in Z3, the PO can be recognized as the formula $\exists~[tr:Trace,aperi:nat1,vdel:nat1,i:nat] P \land Q \land \neg R$ where $P$, $Q$, and $\neg R$ can be encoded separately.
+
+% {\tt tl tr} is the tail of {\tt tr}, and {\tt inds} is the operator of gathering indexes of defined elements of {\tt tr} as the set of natural numbers.
+% In the encoding of $P$, the tail related formula can be encoded as in the range of $[1,(len~tr)-1]$ so that it is not required to define {\tt tl} and {\tt inds} for sequences.
 Finally, we checked the satisfiability of the encoded Z3 code and got {\tt unsat}, which means that the PO is proved since we have negated it.
@@
 
-$\bullet$ PO3 is the postcondition satisfiable obligation of the implicitly specified operation {\tt FaultHeart} that generates a trace of heart pulse signals.
+\subsubsection{Proof Obligation 3}
+
+PO3 of the abstracted pacemaker model is the postcondition satisfiable obligation of the implicitly specified operation {\tt FaultHeart} that generates a trace of heart pulse signals.
 PO3 says that the postcondition of {\tt FaultHeart} has to be satisfiable so that the specification of {\tt FaultHeart} is implementable.
+Fig.~\ref{fig:po3_case1} shows PO3 along with the specification of state and operation {\tt FaultyHeart}.
 
 % \medskip
+\begin{figure}[t]
+\begin{center}
 \begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small]
+\begin{Verbatim}[fontsize=\small,numbers=left]
 exists tr:Trace &
   post_FaultyHeart(oldstate, tr, newstate)
+
+state Pacemaker of
+    aperiod : nat 
+    vdelay  : nat
+  init p == p = mk_Pacemaker(15,10)
+  end
 
@@
 \end{mdframed}
+\vspace{-10pt}
+\caption{PO3, State {\tt Pacemaker}, and operation {\tt FaultyHeart} of Abstraced Pacemaker}
+\label{fig:po3_case1}
+\end{center}
+\end{figure}
 % \medskip
 
-The above VDM-SL code also shows that definition of {\tt FaultyHeart}.
-Recall that the state is defined as a pair of natural numbers {\tt aperiod} and {\tt vdelay}.
+The state {\tt Pacemaker} is specified as a pair of natural numbers {\tt aperiod} and {\tt vdelay}.
 The postcondition of {\tt FaultyHeart} is not specified with {\tt oldstate} which has suffixed tilde, i.e, {\tt aperiod\textasciitilde} and {\tt vdelay\textasciitilde}.
 Therefore, PO3 is a simplified postcondition satisfiable obligation saying that there exists a {\tt Trace} {\tt tr} which satisfies the postcondition specified with only {\tt newstate}\footnote{How the {\tt newstate} is calculated from {\tt oldstate} is not specified in {\tt FaltyHeart}.}.
-The postcondition uses the function {\tt Periodic} which is explicitly specified and returns a boolean value.
-We need to encode {\tt Periodic} as well to prove PO3.
+The postcondition of {\tt FaultyHeart} calls the function {\tt Periodic} with three inputs.
+{\tt Periodic} shown below is an explicitly specified function and returns a boolean value.
+Therefore, state {\tt Pacemaker}, postcondition of {\tt FaultyHeart} ({\tt post\_FaultyHeart}, and function {\tt Periodic} are the context information of PO3.
 
@@
 
-While the specification of {\tt Periodic} is relatively long.
-The encoding of {\tt Periodic} is similar to the encoding of PO8.
-We may observe some similar computations on the trace {\tt tr} such as {\tt in set inds tr} and {\tt len tr}, and we can encode them in the same way as handling PO8.
-Note that we did not apply negation to the formula since existence quantifier in the prefix of a formula is preferred in Z3.
-As a result, we got a result of {\tt sat} so that PO3 is proved.
+Though the specification of {\tt Periodic} is relatively long, the encoding of {\tt Periodic} is similar to the encoding of PO8.
+It is easy to observe similar computations on the trace {\tt tr} such as {\tt in set inds tr} and {\tt len tr}, and we can encode them in the same way as handling PO8.
+Note that the type {\tt Trace} is again needed in proving PO3 so that the encoding in PO8 above can be used as constraints here.
+In the case of PO3, we did not apply negation to the formula since existence quantifier in the prefix of a formula is preferred in Z3.
+As a result, we got a result of {\tt sat} along with a model that satisfies PO3.
+Therefore, PO3 was proved.
 
@@
 (1) Automation: Is the encoding can be automated? partly or fully? 
-(2) Guidance and patterns: What patterns are repeated used in encoding? Any guide of encoding?
+(2) Patterns: What patterns are repeated used in encoding?

### 2017-05-26 paper/studies.tex
studies: both cases complete

@@
 The system state monitors the calling status of users and the connecting status among users.
-Type in the model are based on quote types used to indicate the discrete states of users, then maps are specified to relate the calling and connecting status to users as the system state.
-An invariant is specified for the state {\tt Exchange} to declare the constraints among users and their status in the system.
+Overture tool generated twenty-seven POs, and we selected PO1 and PO14 for report the case study.
 
+\subsubsection{Proof Obligation 1}
+
+PO1 shown in Fig.~\ref{fig:po1_case2} is a legal map application obligation that checks whether maps in the system state are applicable.
+In PO1, types of the state {\tt Exchange} which is a record type of {\tt status} and {\tt calls} is used.
+Therefore, the context information of PO1 is the state including its invariant.
+Fig.~\ref{fig:types_case2} shows the definitions of state {\tt Exchange} with corresponding quote types.
+The state {\tt Exchange} contains two maps: {\tt status} indicates the discrete calling status of users, and is a map from {\tt Subscriber} to {\tt Status}; {\tt calls} models the connection relation of users, i.e., which users are connected, and is a map from {\tt Subscriber} to {\tt Subscriber}.
+Note that {\tt calls} is specified as {\tt inmap} type which means an one-to-one map.
+This means if a user is calling another user, both of the users can not have other connections to third party users.
+Also, {\tt Exchange} has state invariant that is specified for the constraints about connections of users and their calling status.
+At this point, the context information of PO1 is determined and we may proceed to encoding.
+
+\begin{figure}[t]
+\begin{center}
 \begin{mdframed}[roundcorner=5pt]
 \begin{Verbatim}[fontsize=\small]
+forall
+  mk_Exchange(status, calls):EXCH`Exchange &
+    (forall i in set (dom calls) &
+      (i in set (dom status))
+    )
+)
+\end{Verbatim}
+\end{mdframed}
+\vspace{-10pt}
+\caption{PO1 of Telephone Exchange}
+\label{fig:po1_case2}
+\end{center}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\begin{mdframed}[roundcorner=5pt]
+\begin{Verbatim}[fontsize=\small,numbers=left]
 Subscriber = token;
@@
 \end{mdframed}
+\vspace{-10pt}
+\caption{Types and state of Telephone Exchange}
+\label{fig:types_case2}
+\end{center}
+\end{figure}
 
-The strategy of encoding the above types is (1) applying user-defined types in Z3 to encode the quote types; (2) applying uninterpreted functions to encode map type.
-Here we only demonstrate a few selected types since the others are similar.
+The strategy of encoding the types shown in Fit.~\ref{fig:types_case2} is (1) applying user-defined types in Z3 to encode the quote types; (2) applying uninterpreted functions to encode map type.
+In Fig.~\ref{fig:encode_types_case2}, we demonstrate the encoding of selected types.
 
+\begin{figure}[t]
+\begin{center}
 \begin{mdframed}[roundcorner=5pt,shadow=true]
@@
 \end{mdframed}
+\vspace{-10pt}
+\caption{Encoding of quote and map types in Fig.\ref{fig:types_case2} (partial)}
+\label{fig:encode_types_case2}
+\end{center}
+\end{figure}
 
-From above code, the type {\tt Recipient} is defined as an user-defined type with values {\tt WR} and {\tt SR}.
+In Fig.~\ref{fig:encode_types_case2}, the type {\tt Recipient} is defined as an user-defined type with values {\tt WR} and {\tt SR}.
 Type {\tt Status} is further defined as a user-defined type including {\tt Recipient} and {\tt Initiator}.
-Note that since VDM uses partial functions, we also defined a lifted type of {\tt Status}.
-Though {\tt Subscriber} is of token type which does not have specific values when declared, we treated token types as quote types with predefined values such as {\tt S1}, {\tt S2}, and so on.
-Finally the two maps {\tt status} and {\tt calls} are defined as uninterpreted functions.
-Overture generated 27 POs for this model.
-Here we present the encoding and proving of PO1 and PO14.
+Though {\tt Status} is actually an union type combines {\tt Initiator} and {\tt Recipient}, and can be considered an enumeration, we encoded {\tt Status} using user-defined type style in Z3py.
+The reason is that we prefer the subtypes {\tt Initiator} and {\tt Status} can be distinguishable in Z3.
 
-$\bullet$ PO1 is a legal map application obligation that checks whether maps in the system state are applicable.
+Note that since VDM uses partial functions, for the type to be declared as the range of map, i.e., {\tt Subscriber} and {\tt Status}, we defined their lifted type in Z3py to take into consideration the case that the value to a key of a map is undefined.
+Also, though {\tt Subscriber} is of token type which does not have specific values when declared, we treated token types as quote types with predefined domain of values such as {\tt S1}, {\tt S2}, and so on.
+The two maps {\tt status} and {\tt calls} in the state {\tt Exchange} are encoded as uninterpreted functions and the state invariant is encoded by the encoded (and lifted) types mentioned above.
 
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small]
-forall
-  mk_Exchange(status, calls):EXCH`Exchange &
-    (forall i in set (dom calls) &
-      (i in set (dom status))
-    )
-)
-\end{Verbatim}
-\end{mdframed}
+Fig.~\ref{fig:encode_state_inv_case2} shows the encoding of the state invariant.
+Note that the encoding of {\tt i in set (dom calls)} (line 11 of Fig.~\ref{fig:types_case2}) is encoded as {\tt calls(i)!=Subscriber\_lift.NDF} (line 3 of Fig.~\ref{fig:encode_state_inv_case2}) because the set inclusion only means that the domain (key) of {\tt calls} has defined value in the range of {\tt calls}.
+Therefore, we do not have to encode the set of domain neither the set inclusion operator.
+The condition that the range value of a map is defined or not is essential here.
+For example, to encode {\tt status(calls(i)) = <WR>} (line 13 of Fig.~\ref{fig:types_case2}), we have ensure the condition that {\tt calls(i)} has to be defined so that additional constraint {\tt calls(i)!=Subscriber\_lift.NDF} (line 9 of Fig.~\ref{fig:encode_state_inv_case2}).
 
-Since PO1 explicitly uses the system state, we need to consider the invariant of the system state as the context information of PO1. 
-Instead of demonstrate the full encoding of the invariant, we only demonstrate the key encoding for {\tt i in set (dom calls)} and {\tt i in set (dom status)}.
+% Instead of demonstrate the full encoding of the invariant, we only demonstrate the key encoding on {\tt i in set (dom calls)} and {\tt i in set (dom status)}.
 Though there is set inclusion used, instead of defining a set, we can encode the set inclusion as whether {\tt calls(i)} is defined or not.
 
+\begin{figure}[t]
+\begin{center}
 \begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
-  calls(i) != Subscriber_lift.NDF
-  status(i)!= Status_lift.NDF
-\end{Verbatim}
-\end{mdframed}
-
-For PO1, we applied negation to the formula so that the {\tt forall} quantifier for maps {\tt calls} and {\tt status} is removed. The result was {\tt unsat} which means PO1 was proved.
-
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
-Not(
+\begin{Verbatim}[fontsize=\small,numbers=left]
 ForAll(i,
@@
   calls(i)!=Subscriber_lift.NDF,
-      status(i)!=Status_lift.NDF
-    )
-  )
-)
+  Or(
+    And(
+      status(i)==
+        Status_lift.STATUS(
+        Status.INITIATOR(Initiator.WI)),
+	  calls(i)!=Subscriber_lift.NDF,
+      status(Subscriber_lift.
+        get_subscriber(calls(i)))==
+        Status_lift.STATUS(
+          Status.RECIPIENT(Recipient.WR))),
+    And(
+      status(i)==
+        Status_lift.STATUS(
+        Status.INITIATOR(Initiator.SI)),
+      calls(i)!=Subscriber_lift.NDF,
+      status(Subscriber_lift.
+        get_subscriber(calls(i)))==
+        Status_lift.STATUS(
+          Status.RECIPIENT(Recipient.SR)))
+)))
 \end{Verbatim}
 \end{mdframed}
-
-$\bullet$ PO14 is an enumeration map injectivity obligatoin of the operation {\tt Answer} of the system.
-PO14 has free variables, {\tt r} and {\tt status}, in its formula.
-These free variables are treated as the quantifier {\tt forall} is prefixed.
-There are also intermediate variables in the formula such as {\tt m1}, {\tt m2}, {\tt d3}, and {\tt d4}.
-Note that the inverse function of a map, and the range restriction function {\tt :>} are {\tt applied} on calls and {\tt status}.
-Also note that {\tt calls} is a one-to-one map {\tt inmap}.
-
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small]
-(r in set (dom (status :> {<WR>}))) =>
-  (forall m1, m2 in 
-    set {{r |-> <SR>}, 
-         {(inverse calls)(r) |-> <SI>}} &
-    (forall d3 in set (dom m1),
-            d4 in set (dom m2) &
-      ((d3 = d4) => (m1(d3) = m2(d4)))
-))
-\end{Verbatim}
-\end{mdframed}
-
-We negated PO14 so that the {\tt forall} quantifiers are removed.
-To encode the negated formula of PO14, similar technique for  encoding {\tt set (dom status)} for PO1 can be applied.
-For the range restriction function, since the range is a singleton set with element value of {\tt <WR>}, we can skip encoding set notations while limiting the map value {\tt status(r)} to {\tt <WR>}.
-Thus, {\tt r in set (dom (status :> \{<WR>\}))} is encoded as follows.
-
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
-status(r)==
-  Status_lift.STATUS(
-    Status.RECIPIENT(Recipient.WR) )
-\end{Verbatim}
-\end{mdframed}
-
-For the map inverse function, we chose not to define a new map as the inverse of {\tt calls}, but encoded {\tt m1 = \{(inverse calls)(r) |-> <SI>\}} as a case of {\tt m1 in set \{\{r |-> <SR>\}, \{(inverse calls)(r) |-> <SI>\}\}}.
-The case was encoded as the collection of several conditions.
-
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
-# inverse map is not empty
-calls(i)!=Subscriber_lift.NDF 
-# inverse_calls(r) == i
-Subscriber_lift.get_subscriber(calls(i))==r
-
-m1(i)==Status_lift.STATUS(
-         Status.INITIATOR(Initiator.SI)),
-
-ForAll(k, 
-  Implies(k!=i, m1(k)==Status_lift.NDF) )
-\end{Verbatim}
-\end{mdframed}
-
-In above code, the first two conditions promise that (1) the part of map, where key {\tt i} is used and inverse is applied, has to be defined; (2) the domain of the inversed map, i.e., the range of the original map, is {\tt r}.
-Then {\tt m1(i)} where key {\tt i} is referred should be {\tt <SI>}.
-Finally the last condition guarantees that {\tt m1} should be undefined for all indexes except {\tt i}.
+\vspace{-10pt}
+\caption{Encoding of state invariant of Telephone Exchange}
+\label{fig:encode_state_inv_case2}
+\end{center}
+\end{figure}
 
@@
 \end{mdframed}
-\medskip
 
-The result of checking PO14 (negated) ware expected to be {\tt unsat}, but we got {\tt sat}.
-Z3 returned a model that satisfies the negated formula and constraints so that we could examine where is the problem.
-We found that {\tt calls} in the returned model is a map where users calls themselves.
-This does not violate the type of {\tt inmap}, but an user cannot call him/her-self in a telephone exchange system.
+The encoding of PO1 is done by encoding the negated PO1.
+As the following code shows, outside forall of {\tt Exchange} of PO1 is eliminated by the negation.
+The result was {\tt unsat} which means PO1 was proved.
+
+% \begin{mdframed}[roundcorner=5pt,shadow=true]
+% \begin{Verbatim}[fontsize=\small]
+%   calls(i) != Subscriber_lift.NDF
+%   status(i)!= Status_lift.NDF
+% \end{Verbatim}
+% \end{mdframed}
+
+% For PO1, we applied negation to the formula so that the {\tt forall} quantifier for maps {\tt calls} and {\tt status} is removed. The result was {\tt unsat} which means PO1 was proved.
+
+\begin{mdframed}[roundcorner=5pt,shadow=true]
+\begin{Verbatim}[fontsize=\small]
+Not(
+  ForAll(i,
+    Implies(
+      calls(i)!=Subscriber_lift.NDF,
+      status(i)!=Status_lift.NDF
+    )
+  )
+)
+\end{Verbatim}
+\end{mdframed}
+
+\subsubsection{Proof Obligation 14}
+
+PO14 is an enumeration map injectivity obligation of the operation {\tt Answer} of the system.
+PO14 has free variables, {\tt r} and {\tt status}, in its formula.
+These free variables are treated as the bounded by forall quantifier.
+There are also intermediate variables in the formula such as {\tt m1}, {\tt m2}, {\tt d3}, and {\tt d4}.
+Note that the inverse function of a map, and the range restriction function {\tt :>} are on {\tt calls} and {\tt status} separately.
+(Recall that {\tt calls} is a one-to-one map {\tt inmap}.) The context information of PO14 includes the two maps {\tt status} and {\tt calls} but not the state {\tt Exchange}.
+Therefore, only the two maps have to be encoded.
+The state invariant is excluded in the encoding.
+Note that the constraints for {\tt calls} of type {\tt inmap} is included in the context information of PO14, and the encoding in PO1 can be reused here.
+
+\begin{mdframed}[roundcorner=5pt]
+\begin{Verbatim}[fontsize=\small]
+(r in set (dom (status :> {<WR>}))) =>
+  (forall m1, m2 in 
+    set {{r |-> <SR>}, 
+         {(inverse calls)(r) |-> <SI>}} &
+    (forall d3 in set (dom m1),
+            d4 in set (dom m2) &
+      ((d3 = d4) => (m1(d3) = m2(d4)))
+))
+\end{Verbatim}
+\end{mdframed}
+
+When encoding the formula of PO14, firstly, we negated PO14 and eliminated the forall quantifiers.
+To encode the negated formula of PO14, similar technique for  encoding {\tt set (dom status)} in PO1 can be applied.
+For the range restriction function, since the range is a singleton set with element value of {\tt <WR>}, we do not have to encode the set notations and only limit the map's range {\tt status(r)} to {\tt <WR>}.
+Thus, {\tt r in set (dom (status :> \{<WR>\}))} is encoded as follows.
+
+\begin{mdframed}[roundcorner=5pt,shadow=true]
+\begin{Verbatim}[fontsize=\small]
+status(r)==
+  Status_lift.STATUS(
+    Status.RECIPIENT(Recipient.WR) )
+\end{Verbatim}
+\end{mdframed}
+
+For the map inverse function, we chose not to define a new map as the inverse of {\tt calls}, but encoded {\tt m1 = \{(inverse calls)(r) |-> <SI>\}} as a case of {\tt m1 in set \{\{r |-> <SR>\}, \{(inverse calls)(r) |-> <SI>\}\}}.
+The case was encoded as the collection of several conditions.
+
+\begin{mdframed}[roundcorner=5pt,shadow=true]
+\begin{Verbatim}[fontsize=\small,numbers=left]
+# ensure the inverse map is not empty
+calls(i)!=Subscriber_lift.NDF 
+# inverse_calls(r) == i
+Subscriber_lift.get_subscriber(calls(i))==r
+
+m1(i)==Status_lift.STATUS(
+         Status.INITIATOR(Initiator.SI)),
+
+ForAll(k, 
+  Implies(k!=i, m1(k)==Status_lift.NDF) )
+\end{Verbatim}
+\end{mdframed}
+
+In above code, the first two conditions (lines 1-4) promise that (1) the part of map, where key {\tt i} is used and inverse is applied, has to be defined; (2) the domain key of the inversed map, i.e., the range value of the original map, is {\tt r}.
+Then {\tt m1(i)} where key {\tt i} is referred should be {\tt <SI>} (lines 6-7).
+(3) the last condition guarantees that {\tt m1} should be undefined for all indexes except {\tt i} (lines 9-10) because {\tt m1} is a single map that has only one key and one value.
+
+The result of checking PO14 (negated) ware expected to be {\tt unsat}, but we got {\tt sat} instead.
+Z3 returned a model that satisfies the negated formula of PO14.
+After examined the counterexample returned by Z3, we found that {\tt calls} in the returned model is a map where users calls themselves.
+This does not violate the type constraint of {\tt inmap} encoded above, but an user is not allowed to call him/her-self in a telephone exchange system.
 Therefore, the checking of PO14 suggested that the VDM-SL model should add a constraint on {\tt calls}.
-As an example of correction, we may add the constraint below as the type invariant of {\tt calls} or the additional state invariant\footnote{We have added the constraint and checked again, and the PO was proved successfully.}.
+As an example of correction, we may add the constraint below as the type invariant of {\tt calls} or add it to the state invariant\footnote{We have added the constraint and checked again, and the PO was proved successfully.}.
 
@@
 
-In this case study, we did not check all the 27 POs since the encoding could not cover all POs.
+In this case study, we did not check all the twenty-seven POs since the encoding demonstrated in this paper could not cover all POs.
 The main issue is that the maps are encoded as uninterpreted functions so that {\tt forall} quantifier cannot be applied.
 This results that we could not check the postcondition satisfiable obligations of this model\footnote{Besides postcondition satisfiable obligations, there are two POs with type check function {\tt is\_(name, type)}, for which Z3 encoding is not available and then skipped.}.
-Finally, we checked 16 out of all 27 POs.
+After all, we successfully checked 16 out of all 27 POs for the telephone exchange VDM-SL model.
 

### 2017-05-26 paper/studies.tex
studies complete, refs.bib update

@@
 \subsection{Abstract Pacemaker}
+\label{section:case1}
 
@@
 \subsection{Telephone Exchange}
+\label{section:case2}
 
@@
 This results that we could not check the postcondition satisfiable obligations of this model\footnote{Besides postcondition satisfiable obligations, there are two POs with type check function {\tt is\_(name, type)}, for which Z3 encoding is not available and then skipped.}.
-After all, we successfully checked 16 out of all 27 POs for the telephone exchange VDM-SL model.
-
-\subsection{Discussion}
+After all, we successfully checked sixteen out of all twenty-seven POs for the telephone exchange VDM-SL model.
 
@@
 
-Points of discussion:
-(1) Automation: Is the encoding can be automated? partly or fully? 
-(2) Patterns: What patterns are repeated used in encoding?
+\subsection{Discussion}
+
+As an overall result of our case studies, the objective of our approach that to discharge POs of a VDM-SL model as many as possible is achieved.
+In the first case study (Abstracted Pacemaker), we discharged all nine POs; in the second case study, we discharged sixteen POs our of total twenty-seven POs (about 60\%).
+To increase the percentage of discharging POs, we may further used different encoding, for example, encoding maps using arrays in Z3, to solve remaining POs.
+
+% Usability and Applicability of the Encode Style
+The two case studies demonstrated the application of the encode strategy described in Section~\ref{section:encode-strategy}.
+From our experiences, the cost and complexity of encoding varies from specifications in a VDM model.
+Usually, the complexity of encoded formulas usually depends on the complexity of types used in a VDM model.
+For example, natural number based types such as {\tt seq of nat} or {\tt map from nat to nat} have less complexity, and it shall be easier to encode and solve this kind of types with SMT solvers.
+On the contrary, for types such as a map of sequences or a set of maps, the encoding may be difficult and the complexity of the encoded formulas may grow a lot.
+
+The encoding demonstrated in the two case studies can be referred as encoding patterns.
+For example, the encoding of union type and quote type with user-defined types or enumeration type, and the encoding of maps with uninterpreted functions, are useful for later encoding of POs of other VDM-SL models.
+Also, the patterns of encoding like the constraints for defining a sequence from arrays, and set inclusion of indexes of a sequence, or set inclusion of domain/range of a map, are also reusable.
+In our case studies, these encoding patterns are actually applied multiple times.
+
+% Automation of the encoding style and strategy
+Though the encoding patterns are useful, applying the encoding patterns still depends on specification of VDM models because that the encoding patterns are not systemic or syntax-based.
+Considering automated encoding of POs and related context information, the types are easier to encode with automation, but encoding formulas of POs remains manual and rely on encode patterns.
+
+% Scalability and Efficiency}
+From our case studies, it can be recognized that the efficiency of SMT solvers is one the advantages of our approach.
+Also, the counterexample (model) generation of SMT solvers provides hints for revision and correction of VDM models.
+As reported in Section~\ref{section:case2}, encoding and solving PO14 of the telephone exchange VDM-SL model found an inconsistency and the counterexample provides a guidance of adding a constraint to fix the inconsistency\footnote{A previous work~\cite{kenneth:ifm2013} that translates VDM-SL to Alloy~\cite{kenneth:ifm2013} checked the VDM-SL model and found a different problem.
+The model (from the Overture repository) we used for case study is the fixed one which means that the inconsistency we found was not found in~\cite{kenneth:ifm2013}.}.
+Finally, the scalability of our approach is not serious because we only encode the formula of a PO and its context information.
+The size of VDM-SL code to be encoded does not escalate with the size of a VDM-SL model.
+
+% Points of discussion:
+% (1) Automation: Is the encoding can be automated? partly or fully? 
+% (2) Patterns: What patterns are repeated used in encoding?

### 2017-05-26 paper/conclusions.tex
conslusions complete

In this paper, we have proposed an approach to validation of VDM models by discharging proof obligations (POs) with SMT solvers. In our approach, Overture tool is used as the proof obligation generator, and the Python API of Z3 SMT sover is used for encoding and solving formulas of generated POs. Encode strategies including steps of encoding as well as patterns of encoding are demonstrated and two case studies are reported. From our experiences, it is recognized that discharging POs using SMT solvers is efficient, and the counterexample generation of SMT solvers when the discharging fails can provide useful guidance for correcting inconsistency of VDM models. We also discussed other issues such as applicability of encoding strategies considering the automation of our approach. For the next steps of our approach, we are considering more patterns of encoding and how to organize the patterns of encoding for challenging the automation issue.

### 2017-05-26 paper/related.tex
related.tex complete, refs.bib update

@@
-(1) work about discharging POs using theorem provers. (2) other work about validation/verfication of VDM models rather than discharging POs.
+% Development history of VDM POG.
+Proof is an essential to guarantee the consistency of VDM models~\cite{978-3-540-19813-0}, and the proof obligation generation of VDM is firstly proposed by B.
+K.
+Aichernig and P.
+G.
+Larsen~\cite{AL:97:POGV}.
+Later, A.
+Ribeiro and P.
+G.
+Larsen worked on proof obligation generation and discharging related to termination of recursive functions~\cite{Ribeiro2010}.
 
-For example, \cite{Verm:2007:master,Vermolen:2010:PCV:1774088.1774608} translate VDM models to HOL and then prove the POs within the environment of HOL.
+% Work related to discharging POs using theorem provers. 
+There are several researches focusing on translation from VDM to theorem provers and discharging POs of VDM with theorem provers.
+S.
+Agerholm~\cite{Agerholm1996} proposed translation and from VDM-SL to PVS.
+This work proposed a translation from a subset of VDM-SL to PVS for type checking.
+S.
+Maharaj and J.
+Bicarregui~\cite{632849} used Agerholm's translation to assist the verification of VDM-SL models and refinements.
+Later, S.
+D.
+Vermolen et al.~\cite{Verm:2007:master,Vermolen:2010:PCV:1774088.1774608} utilized the parsing mechanism of Overture tools and developed automated translation from a subset of VDM-SL to HOL theorem prover.
+Vermolen's work also provided some tactics to support automated discharging of POs.
+L.
+D.
+Cuto et al.~\cite{CFP:14:TVCSAP} used the proof obligation generation mechanism for translating CML, a modeling language combing VDM-SL and CSP, to Isabella/HOL for automated proof.
+Recently, work on improvement of translating VDM-SL to HOL are also proposed~\cite{CT:15:EOCGTIS}.
+
+% Comparison to our approach
+The theorem prover based approaches has common issues: (1) the translation from VDM-SL to theorem provers may be erroneous since the translation applies on the whole VDM-SL model; (2) The proof process is usually complicated and requires an expert; (3) if theorem provers fail to prove a PO, there is little information to indicate that the problem is due to incorrect VDM-SL model, incorrect translation, or incorrect tactics.
+Compare to our approach that discharges VDM-SL proof obligations with SMT solvers, we adopt the efficiency of SMT solvers and our encoding does not require to be applied on the whole VDM-SL model.
+Furthermore, the functionality of counterexample generation is helpful to determine where the problem is.
+However, encoding in our approach is manual at this point and still need improvements.
+
+% Work related to validation/verfication of VDM models with model checking.
+On the other hand, verification and validation of VDM-SL models with model checking techniques are recently proposed.
+K.
+Lausdahl~\cite{kenneth:ifm2013} proposed a semantics-preserving translation that constructs an Alloy model from a subset of VDM-SL.
+This work aims to support the validation of implicitly specified VDM-SL models with Alloy.
+H-H.
+Lin et al.~\cite{DBLP:conf/ftscs/LinOKA15} proposed an approach that combines SPIN model checker and the VDMJ interpreter for verifying explicitly specified VDM-SL models.
+Lausdahl's work is similar to our approach since Alloy is based on SAT solvers, but SMT solvers in our approach can solve formulas more efficiently.
+Both the two works require types involved in a VDM-SL model being bounded, while bounds are not necessary for the encoding in our approach.
+Also, there are issues for Alloy and SPIN to represent the rich and complex VDM-SL types.
+
+% For example, \cite{Verm:2007:master,Vermolen:2010:PCV:1774088.1774608} translate VDM models to HOL and then prove the POs within the environment of HOL.
 However, this approach has several issues: (1) the translation is applied to the whole VDM model so that it may be tedious and error-prone; (2) The proof process is usually complicated and requires an expert; (3) It is not clear where the problem is when the proof fails.
 
-\cite{AL:97:POGV,CFP:14:TVCSAP,CT:15:EOCGTIS}
+% \cite{CFP:14:TVCSAP,CT:15:EOCGTIS}

### 2017-05-27 paper/conclusions.tex
polished (all)

@@
 In our approach, Overture tool is used as the proof obligation generator, and the Python API of Z3 SMT sover is used for encoding and solving formulas of generated POs.
-Encode strategies including steps of encoding as well as patterns of encoding are demonstrated and two case studies are reported.
+Encoding strategies including steps of encoding as well as encoding patterns are demonstrated, and two case studies are reported.
 From our experiences, it is recognized that discharging POs using SMT solvers is efficient, and the counterexample generation of SMT solvers when the discharging fails can provide useful guidance for correcting inconsistency of VDM models.
 We also discussed other issues such as applicability of encoding strategies considering the automation of our approach.
-For the next steps of our approach, we are considering more patterns of encoding and how to organize the patterns of encoding for challenging the automation issue.
+As the future work, we are gathering and organizing more encoding patterns for challenging the automation issue.

### 2017-05-27 paper/introduction.tex
polished (all)

@@
 
-More specifically, in this approach we encode and prove POs of VDM-SL models by solving SMT formulas with Python API of Z3~\cite{MB:08:ZSS}.
+More specifically, in this approach we chose the Overture tool as the PO generator.
+For the POs generated by the Overture tool, we encode and prove POs of VDM-SL models by solving SMT formulas with Python API of Z3~\cite{MB:08:ZSS}.
 Z3 is a popular SMT solver widely used in software verification and its Python API provides flexibility in constructing SMT formulas~\cite{BarFT-SMTLIB}.
@@
 We conducted case studies and the results showed that our approach can efficiently discharge the significant part of POs of a VDM-SL model.
-We chose the Overture tool as the PO generator in this paper\footnote{POs generated by Overture and VDMTools are slightly different in the numbers and styles of POs}.
+%  \footnote{POs generated by Overture and VDMTools are slightly different in the numbers and styles of POs}.
 

### 2017-05-27 paper/obligations.tex
polished (all)

@@
 % Generation information of POs
-For a VDM model, a proof obligation is a statement that must be proved in order to ensure the consistency of the model.
+For a VDM model, a proof obligation (PO) is a statement that must be proved in order to ensure the consistency of the model.
 A PO contains a predicate with its context information as shown below.
@@
 PO1 is called ``legal sequence application obligation'' that relates to the sequence involved expression (line 9-10) of the VDM-SL model {\tt CMDS}).
-PO1 says that for all k in the set from 1 to {\tt len s.commands -1}, k should be always within the bound of defined indexes of the sequence {\tt s.commands}.
+PO1 says that for all k in the set from 1 to {\tt len s.commands -1}, k should be always within the bound of defined indices of the sequence {\tt s.commands}.
 Also, it can be recognized that if we use {\tt len s.commands} instead of {\tt len s.commands -1} the proof will fail.
@@
 In VDM-SL, the {\tt pre\_} and {\tt post\_} prefix represent the precondition and postcondition of a function/operator separately.
-They are evaluable functions if given appropriate inputs, and return a boolean value.
+They are evaluable functions if given appropriate arguments, and return a boolean value.
 In a satisfiable obligation of an operation in which the state change is usually expected, the name of state (in this case {\tt commands}) represents {\tt newstate} while the same name postfixed with ``~\textasciitilde~'' (in this case {\tt commands\textasciitilde}) represents {\tt oldstate}.

### 2017-05-27 paper/related.tex
polished (all)

@@
 % Development history of VDM POG.
-Proof is an essential to guarantee the consistency of VDM models~\cite{978-3-540-19813-0}, and the proof obligation generation of VDM is firstly proposed by B.
+Proof is essential to guarantee the consistency of VDM models~\cite{978-3-540-19813-0}, and the proof obligation generation of VDM is firstly proposed by B.
 K.
@@
 S.
-Agerholm~\cite{Agerholm1996} proposed translation and from VDM-SL to PVS.
-This work proposed a translation from a subset of VDM-SL to PVS for type checking.
+Agerholm~\cite{Agerholm1996} proposed a translation from a subset of VDM-SL to PVS for type checking.
 S.
 Maharaj and J.
-Bicarregui~\cite{632849} used Agerholm's translation to assist the verification of VDM-SL models and refinements.
+Bicarregui~\cite{632849} used Agerholm's translation to assist the verification of VDM-SL models and their refinements.
 Later, S.
@@
 D.
-Cuto et al.~\cite{CFP:14:TVCSAP} used the proof obligation generation mechanism for translating CML, a modeling language combing VDM-SL and CSP, to Isabella/HOL for automated proof.
-Recently, work on improvement of translating VDM-SL to HOL are also proposed~\cite{CT:15:EOCGTIS}.
+Cuto et al.~\cite{CFP:14:TVCSAP} used the proof obligation generation mechanism of Overture tools for translating CML, a modeling language combing VDM-SL and CSP, to Isabella/HOL for automated proof.
+Recently, work on improvement of translating VDM-SL to Isabella/HOL are proposed~\cite{CT:15:EOCGTIS}.
 
 % Comparison to our approach
-The theorem prover based approaches has common issues: (1) the translation from VDM-SL to theorem provers may be erroneous since the translation applies on the whole VDM-SL model; (2) The proof process is usually complicated and requires an expert; (3) if theorem provers fail to prove a PO, there is little information to indicate that the problem is due to incorrect VDM-SL model, incorrect translation, or incorrect tactics.
-Compare to our approach that discharges VDM-SL proof obligations with SMT solvers, we adopt the efficiency of SMT solvers and our encoding does not require to be applied on the whole VDM-SL model.
-Furthermore, the functionality of counterexample generation is helpful to determine where the problem is.
+The theorem prover based approaches has common issues: (1) the translation from VDM-SL to theorem provers may be errorneous since the translation applies on the whole VDM-SL model; (2) The proof process is usually complicated and requires an expert even with the help of proof tactics; (3) Theorem provers provide little information when they fail to discharge POs.
+Compare to our approach that discharges POs of VDM-SL with SMT solvers, we adopt the efficiency of SMT solvers and the encoding does not require to be applied on the whole VDM-SL model.
+Furthermore, the counterexample generation of SMT solvers is helpful for locating the problem.
 However, encoding in our approach is manual at this point and still need improvements.
@@
 Lin et al.~\cite{DBLP:conf/ftscs/LinOKA15} proposed an approach that combines SPIN model checker and the VDMJ interpreter for verifying explicitly specified VDM-SL models.
-Lausdahl's work is similar to our approach since Alloy is based on SAT solvers, but SMT solvers in our approach can solve formulas more efficiently.
-Both the two works require types involved in a VDM-SL model being bounded, while bounds are not necessary for the encoding in our approach.
-Also, there are issues for Alloy and SPIN to represent the rich and complex VDM-SL types.
+Lausdahl's work is similar to our approach since Alloy uses SAT solvers, but SMT solvers in our approach can solve wider formulas with more efficiency.
+Both Lausdahl's and Lin's work require types involved in a VDM-SL model being bounded, and there are issues for Alloy and SPIN to represent the rich and complex VDM-SL types.
+Compare to them, bounds are not necessary for SMT solvers in our approach and our encoding patterns can applied to complex types thanks for build-in theories in SMT sovers.
 

### 2017-05-27 paper/strategy.tex
minor fix encode strategy
polished (all)

@@
 % Z3 introduction, API introduction
-As mentioned in Section~\ref{section:introduction}, our approach encodes and prove VDM POs with the Z3 SMT solver~\cite{MB:08:ZSS}.
+As mentioned in Section~\ref{section:introduction}, our approach encodes and proves VDM POs with the Z3 SMT solver.
 Z3 is one of the popular SMT solver wildly used in verification of programs and software systems.
@@
 For pre- and post-conditions of a function/operation specified in the VDM model, the encoding is needed if {\tt pre\_<name>} or {\tt post\_<name>} are used in the PO.
-For example, the postcondition satisfiable obligation (PO7) of the {\tt CMDS} module described in Section~\ref{section:proof-obligations} used both precondition and postcondition.
+For example, PO7 of the {\tt CMDS} module described in Section~\ref{section:proof-obligations} mentioned both precondition and postcondition, so both conditions shall be encoded.
 \item
 If a PO only considers a part of the specification of a type, function, or an operations, it is only needed to encode the expressions that are directly used in the PO.
-In other words, if a type, a function, or an operation is not directly used such as variables of the type, evaluation using the function or the operation, there is no need to encode the full specification of the type, the function, or the operation.
+In other words, if a type, a function, or an operation is not directly mentioned, for example, a variable mentioned with {\tt name:type} style, or a function mentioned with {\tt fun\_name(arguments)} style, there is no need to encode the full specification of the type, the function, or the operation.
 \end{enumerate}
 \item
-Quantifier elimination and obligation negation
+Obligation negation and quantifier elimination
 \begin{enumerate}
 \item
-Obligation negation: In model checking with SMT solvers, the assertion to be checked is usually negated, and the result of unsat is expected and means the proof of the assertion succeeds.
-For the result of sat which means the assertion can be violated, we can get a model as the counterexample that shows the information of how the assertion is violated.
+Obligation negation: In model checking with SMT solvers, the assertion to be checked is usually negated, and the result of unsatmeans the proof of the assertion succeeds.
+For the result of sat which means the assertion can be violated, SMT solvers return a model as the counterexample that shows an example of how the assertion is violated.
 In this work, we also take negation on the PO to be discharged.
@@
 Elimination of forall quantifiers: POs of VDM usually have forall quantifiers.
-To solve the formulas with forall quantifiers in Z3, though Z3 accepts formulas with forall quantifiers, there are limitations and we may need to eliminate forall quantifiers that may cause problem before encoding.
+To solve the formulas with forall quantifiers in Z3, though Z3 accepts formulas with forall quantifiers, there are limitations applying forall quantifiers in Z3 so that by default we have to eliminate forall quantifiers before encoding.
 In this work, forall quantifier usually causes difficulties in encoding and solving when complex types are involved.
 Obligation negation described above may achieve the elimination of troublesome foall quantifiers.
-However, obligation negation should be applied carefully not to introduce troublesome forall quantifiers from exists quantifiers.
+On the other hand, obligation negation should be applied carefully not to introduce troublesome forall quantifiers from exists quantifiers.
 \end{enumerate}
+\item
+Encode the predicate of obligation and check for satisfiability.
 \end{enumerate}
@@
 % More details of encoding types, operators, and expressions
-VDM has rich types such as natural number, quote, sequence, and mapping, and there are no direct matching for all VDM types to the types in Z3.
+VDM has rich types from natural number and quote to sequence and maps.
+There are no direct matching for all VDM types to the types in Z3.
 For example, to encode a natural number, we need to declare an integer and then add a constraint of greater and equal to zero because there is no natural number type in Z3.
-Also, the enumeration in Z3 can be used to encode simple quote types.
-For complex types like sequences, maps, or records, we may use uninterpreted functions or arrays to capture the characteristic and operators of the types we want to encode.
-At this point, there is not yet a complete and systematic way of encoding VDM types to Z3, what we can do is to choose the encoding carefully based on the types and expressions in a VDM model.
-However, based on our case studies, there are some patterns of encoding that are frequently used for corresponding expression styles in VDM-SL models.
+The enumeration type in Z3 can be used to encode simple quote types.
+For complex types like sequences, maps, or records, we may use uninterpreted functions or arrays as the base type and capture the characteristic and operators of the VDM types by adding corresponding constraints.
+At this point, there is not yet a complete and systematic way of encoding VDM types in Z3.
+What we can do is to choose the encoding carefully based on the types and expressions involved in a PO.
+Based on our case studies, there are patterns of encoding that are frequently used for corresponding expression styles in VDM-SL models.
 
@@
 In this case, {\tt i} is the encoding an instance of type {\tt nat} in VDM.
-Note that we cannot define a user-defined type in Z3 but only can declare every instance as Z3 type and add necessary constraints to achieve the encoding of variables of VDM types.
-This means that if there are several variables declared as VDM type that Z3 does not have, instances of the same number of VDM variables have to be declared in Z3 with constraints for each instance added to the solver object.
-The strategies discussed above will be demonstrated with details in Section~\ref{section:case-studies} through the report of case studies.
+Note that we cannot add constraints on types but on instance of types in Z3.
+This means that if there are several variables of the same VDM type that Z3 does not have, instances of the same number of variables have to be declared in Z3 with constraints for every instance added to the solver object.
+The steps and strategies of encoding discussed above will be demonstrated with details in Section~\ref{section:case-studies} through the report of case studies.
 

### 2017-05-27 paper/studies.tex
polished (all)

--- 2017-05-26 paper/studies.tex
@@
-This section gives studies that apply our encoding to selected VDM-SL models and discusses the results.
-All the selected VDM-SL models can be found in Overture repository\footnote{http://overturetool.org/download/examples/VDMSL/}.
+This section gives studies that apply our approach to selected VDM-SL models and discusses the results\footnote{All the selected VDM-SL models can be found in Overture repository: {\tt http://overturetool.org/download/examples/VDMSL/}, and the encoding and results with Z3py can be founded in our repository: {\tt https://github.com/fmlab-iis/VDM-SMT}}.
 
 \subsection{Abstract Pacemaker}
+\label{section:case1}
 
@@
 For this model of 58 lines, Overture generated nine POs.
-Here we report two the encoding and solving of obligations PO8 and PO3 for this case.
+Here we report the encoding and solving of obligations PO8 and PO3 below.
 
@@
 PO8 shown in Fig.~\ref{fig:po8_case1} is a legal sequence application obligation of operation {\tt Pace}.
-PO8 requires that the formulas specified with trace {\tt tr} of type {\tt Trace} be legal, i.e., computable to get its result.
+PO8 requires that the formula specified with trace {\tt tr} of type {\tt Trace} be legal, i.e., computable to get its result.
 For more understanding of PO8, we may take a look on the definition of the operation {\tt Pace} as shown in Fig.~\ref{fig:pace_case1}.
-{\tt Pace} is an explicitly specified operation and takes three inputs of types {\tt Trace}, {\tt nat1}, and {\tt nat1}.
-It is easy to recognize that PO8 is checking about the sequence used in line 4 in Fig.~\ref{fig:pace_case1}.
+{\tt Pace} is an explicitly specified operation and takes three arguments of types {\tt Trace}, {\tt nat1}, and {\tt nat1}.
+It is easy to recognize that PO8 is checking about the sequence used in line 4 of PO8 in Fig.~\ref{fig:pace_case1}.
 Note that to discharge PO8, we do not need the full definition of {\tt Pace}.
@@
 \end{center}
+% \vspace{-20pt}
 \end{figure}
@@
 \end{center}
+\vspace{-20pt}
 \end{figure}
 
-As mentioned in Section~\ref{section:proof-obligations}, PO8 is the formula to be proved and we need to include the context information related to the obligation as the first step of encoding described in Section~\ref{section:encode-strategy}.
-To decide what to include as the context information, we firstly examine the types used in PO8 and we have {\tt nat1} and {\tt Pace}.
-{\tt nat1} is a type of natural numbers that excludes zero and thus starts from one.
+As described in Section~\ref{section:encode-strategy}, the first step of encoding is to determine the context information of PO8.
+To decide what to include as the context information, we firstly examine the types used in PO8: {\tt nat1} and {\tt Pace}.
+{\tt nat1} is a type of natural numbers that excludes zero and starts from one.
 {\tt nat1} is a primitive type of VDM so we do not need more information about it.
-{\tt Pace} is a user-defined type so we have check its definition in the VDM-SL model.
-The type definition of {\tt Trace} in the model is as follows.
-Note that {\tt oldstate} bounded by the forall quantifier is used in the expression of PO8 and therefore is ignored in the encoding.
+{\tt Pace} is a user-defined VDM type so we have check its specification in the VDM-SL model.
+Below shows the type definition of {\tt Trace}.
+Note that {\tt oldstate} bounded by the forall quantifier is never used in the predicate of PO8 and therefore will be ignored in the encoding.
 
@@
 
-Now we know that {\tt Trace} is of sequence type of {\tt Event}, where {\tt Event} is of quote type with two values {\tt <A>} and {\tt <V>} that are stand for the A-pace and V-pace of heartbeat signals.
-At this point, we have necessary context information we need to discharge PO8 and we may proceed to encoding step.
-For discharging PO3, we need to encode types used in PO8: {\tt nat1} and {\tt Trace}.
+{\tt Trace} is of type sequence of {\tt Event}, where {\tt Event} is of quote type with two values {\tt <A>} and {\tt <V>} that stand for the A-pace and V-pace of heartbeat signals.
+At this point, we have necessary context information needed for discharging PO8 and we may proceed to the next step: encoding context information.
+For PO3, we need to encode types used in PO8: {\tt nat1} and {\tt Trace}.
 {\tt nat1} is the natural number that excludes zero, so the encoding of {\tt nat1} is as follows.
@@
 
-As the above code shows, we defined two instances of integer in Z3, {\tt aperi} and {\tt vdel}, then the constraint of greater and equal to one is added into the solver object for both instances.
-Note that in Z3, we cannot add constraints to types but instances of a Z3 type.
-This has been mentioned in Section~\ref{section:encode-strategy}.
+As the above code shows, we defined two instances of integer in Z3, {\tt aperi} and {\tt vdel}.
+Recall that in Z3, we cannot add constraints to types but instances, so the constraint of greater and equal to one is added twice to the solver object for both {\tt aperi} and {\tt vdel}.
 The remaining type not encoded yet is {\tt Trace}, the sequence type.
@@
 The code below shows the encoding of {\tt Event} and {\tt Trace}.
-Since {\tt Trace} is not exactly the sequence of {\tt Event} but {\tt [Event]} (a union with {\tt nil}, we did not define {\tt Event} but define {\tt [Event]}.
-Also, because VDM uses partial functions, a type can be undefined.
-This means that besides ordinary values as specification shows, we have to consider the case that the value is not defined.
-In our work, we used {\tt NDF} to represent the undefined value of {\tt Event}, and define {\tt Event\_lift} in Z3.
-The {\tt \_lift} indicates that we have considered the undefined value.
+Since {\tt Trace} is not exactly the sequence of {\tt Event} but {\tt [Event]} (a union with {\tt nil}), we did not define {\tt Event} but define {\tt [Event]} directly.
+Furthermore, a type in VDM can be undefined because of VDM's logic of partial functions.
+This means that besides predefined values as in the specification, we have to consider the case that the value is undefined.
+In our work, we used {\tt NDF} to represent the undefined value of {\tt Event}, and defined {\tt Event\_lift} in Z3.
+The postfixed {\tt \_lift} indicates that we have considered the undefined value.
 
@@
 Note again that the constraints are applied on {\tt tr}, the instance of {\tt Trace}, not the type {\tt Trace}.
-In Fig.~\ref{fig:tr_constraint_case1}, we added three constraints:
+As shown in Fig.~\ref{fig:tr_constraint_case1}, we added three constraints:
 
@@
 \end{center}
+\vspace{-20pt}
 \end{figure}
@@
 \item
-The first constraint (line 4) says that the all indexes of {\tt tr} below 1 should be undefined because in VDM-SL, a sequence's index starts from 1.
+The first constraint (line 4) says that all indices of {\tt tr} below 1 should be undefined because in VDM-SL, a sequence's index starts from 1.
 \item
-The second constraint (lines 6-13) says that if an index is defined, i.e., not undefined, all indexes lower than it should be all defined.
+The second constraint (lines 6-13) says that if an index is defined, i.e., not undefined, all indices lower than it should be all defined.
 \item
-The third constraint (line 15-22) says tat if an index is undefined, all indexes higher than it should be all undefined.
+The third constraint (line 15-22) says tat if an index is undefined, all indices greater than it should be all undefined.
 \end{itemize}
@@
 From the above three constraints, we limited an array ({\tt tr}) to a sequence that starts from index 1 to its last element.
-Now we have encoded both types {\tt nat1} and {\tt Trace} used in PO8 of the VDM-SL model.
-However, before proceed to encode the obligation, it is necessary to encode the function {\tt len} that computes the length of a sequence.
+Now we have encoded both types {\tt nat1} and {\tt Trace} used in PO8.
+Before proceed to encode the obligation, it is necessary to encode the function {\tt len} that computes the length of a sequence.
 As shown in Fig.~\ref{fig:tr_len_case1}, we defined an uninterpreted function {\tt len\_tr} for {\tt tr}, the instance of {\tt Trace}.
-Line 1 is the type declaration of {\tt len\_tr} which is an uninterpreted function accepting the input of a {\tt Trace} type and returning an integer type.
-Similar to defining constraints for {\tt tr:Trace}, we gave {\tt len\_tr} the constraint that consists of two cases: the length of a sequence can only be equal or greater than 0.
+Line 1 is the type declaration of {\tt len\_tr} which is an uninterpreted function taking an argument of a {\tt Trace} type and returning an integer type.
+Similar to defining constraints for {\tt tr:Trace}, we gave {\tt len\_tr} the constraint considering two cases: the length of a sequence can be equal or greater than 0.
 
@@
 \item
-If {\tt len\_tr} returns 0, then {\tt tr} is an sequence with no element.
-That is, {\tt tr[1]} has to be undefined so that all indexes of {\tt tr} are then undefined based on previous constraints of {\tt tr}.
-(lines 4-7)
+(lines 4-7) If {\tt len\_tr} returns 0, then {\tt tr} is an sequence with no element.
+That is, {\tt tr[1]} has to be undefined so that all indices of {\tt tr} are then undefined based on previous constraints of {\tt tr}.
 \item
-If {\tt len\_tr} returns an integer greater than 0, then {\tt tr} has defined elements till its length index {\tt len\_tr(tr)} such that the element at and after index {\tt len\_tr(tr)+1} should be undefined.
-(line 8-12)
+(line 8-12) If {\tt len\_tr} returns an integer greater than 0, then {\tt tr} has defined elements till its length index {\tt len\_tr(tr)} such that the element at and after index {\tt len\_tr(tr)+1} should be undefined.
 \end{enumerate}
@@
 The negated PO8 is shown in Fig.~\ref{fig:po8_negation_case1}.
-Compare to the original PO8 in Fig.~\ref{fig:po8_case1}, the negated PO8 also eliminated the forall quantifier so that we do not need to use forall in Z3py encoding.
-The PO can be recognized as the formula $\exists~[tr:Trace,aperi:nat1,vdel:nat1,i:nat] P \land Q \land \neg R$ where $P$, $Q$, and $\neg R$ can be encoded separately.
+Compare to the original PO8 in Fig.~\ref{fig:po8_case1}, the obligation negation also eliminated the forall quantifier.
+The negated PO8 can be recognized as the formula $\exists~[tr:Trace,aperi:nat1,vdel:nat1,i:nat] P \land Q \land \neg R$ where $P$, $Q$, and $\neg R$ can be encoded separately.
 
@@
 Finally, we encoded the negated PO8 as the Z3py code shown in Fig.~\ref{fig:po8_encoded_case1}.
-The encoding has four constraints: two {\tt nat1} variables (line 1); encoding of $P$ ({\tt i in set (inds (tl tr)}) that is actually indexes from 1 to {\tt len tr -1} (line 2); encoding of $Q$ ({\tt (i mod aperi) = (vdel + 1)}) (line 3); encoding of $\neg R$ ({\tt not(i in set (inds tr))}) (line 4).
-Note that {\tt hd} and {\tt tl tr} are operators of the head and tail of a sequence, and {\tt inds} is the operator of gathering indexes of defined elements of a sequence and returns a set of natural numbers.
+The encoding has four constraints: two {\tt nat1} variables (line 1); encoding of $P$ ({\tt i in set (inds (tl tr)}) that is actually indices from 1 to {\tt len tr -1} (line 2); encoding of $Q$ ({\tt (i mod aperi) = (vdel + 1)}) (line 3); encoding of $\neg R$ ({\tt not(i in set (inds tr))}) (line 4).
+Note that {\tt hd} and {\tt tl tr} are operators of the head and tail of a sequence, and {\tt inds} is the operator of gathering indices of defined elements of a sequence and returns a set of natural numbers.
 Also note that in the encoding in Fig.~\ref{fig:po8_encoded_case1}, we did not explicitly encode {\tt hd}, {\tt tl}, and {\tt inds} and simplified the results of these operators as constraints related to ranges (upper and lower bound) of {\tt tr}.
-The encoded PO8 (negated) was checked by the solver object in Z3py for the satisfiability.
-The result was {\tt unsat}, which means that PO8 is proved.
-
-% With the above two cases of {\tt len\_tr} that defines the length function of traces, together with constraints of {\tt tr}, and instances of types {\tt nat1}, we encoded PO8 as shown in
-
-% The data types used in the model are defined as follows:
-
-% \begin{mdframed}[roundcorner=5pt]
-% \begin{Verbatim}[fontsize=\small]
-%   Trace = seq of [Event];
-%   Event = <A> | <V>;
-
-%   state Pacemaker of
-%     aperiod : nat 
-%     vdelay  : nat
-%   init p == p = mk_Pacemaker(15,10)
-%   end
-% \end{Verbatim}
-% \end{mdframed}
-
-% {\tt Trace} is of sequence type of {\tt Event}, where {\tt Event} is of quote type with two values {\tt <A>} and {\tt <V>}.
-The state is defined as two natural numbers with initial value of {\tt (15,10)}.
-In Z3, natural numbers is encoded as positive integers (integers with constraint of greater and equal to zero); quote type can be encoded as enumeration type {\tt EnumSort}; sequence type can be encoded as {\tt ArraySort}.
-The definitions of {\tt Event} and {\tt Trace} are as follows:
-
-% Furthermore, since {\tt Trace} is defined as an array in Z3, we added constraints that limit an array as a sequence.
-Note that in Z3, the constraints are applied on each instance of type {\tt Trace} ({\tt tr} in the following formulas), not on the type itself.
-
-% $(len\_tr(tr) = 0 \land tr[1] = 0) \lor (len\_tr(tr) > 0 \land tr[len\_tr(tr)] \neq NDF \land tr[len\_tr(tr)+1] = NDF$
-
-% {\tt len\_tr} is defined as an uninterpreted function that takes {\tt tr} as input and returns an integer indicates the length of {\tt tr}.
-Since the length of a sequence can only be equal or greater than 0, the constraint of {\tt len\_tr} is divided into two cases:
-
-% We have defined the context information related to type {\tt Trace}.
-Now we can procceed to encode and prove the POs of the model.
-For this VDM-SL model, Overture generated nine POs.
-Most of the POs are similar so we only demonstrate PO8 and PO3.
-
-% $\bullet$ PO8 is a legal sequence application obligation of operation {\tt Pace} which is explicitly specified.
-This PO requires that the formulas specified with trace {\tt tr} be legal, i.e., computable to get its result.
-
-% In PO8 showed above, {\tt tl tr} is the tail of {\tt tr}, and {\tt inds} is the operator of gathering indexes of defined elements of {\tt tr} as the set of natural numbers.
-To prove PO8, firstly we took negation of PO8 which results a quantifier eliminated formula.
-
-% Note that the {\tt oldstate} is not used in the operation so that we remove it in the above formula. 
-% To encode the PO in Z3, the PO can be recognized as the formula $\exists~[tr:Trace,aperi:nat1,vdel:nat1,i:nat] P \land Q \land \neg R$ where $P$, $Q$, and $\neg R$ can be encoded separately.
-
-% {\tt tl tr} is the tail of {\tt tr}, and {\tt inds} is the operator of gathering indexes of defined elements of {\tt tr} as the set of natural numbers.
-% In the encoding of $P$, the tail related formula can be encoded as in the range of $[1,(len~tr)-1]$ so that it is not required to define {\tt tl} and {\tt inds} for sequences.
-Finally, we checked the satisfiability of the encoded Z3 code and got {\tt unsat}, which means that the PO is proved since we have negated it.
-
-% {\tt Periodic}: legal sequence application obligation\\
-% This PO is meant to represent the applicability of statements used in the function {\tt Periodic}, that is, the expressions in {\tt Periodic} are computable to get a result.
-The formula of the PO is in the form of
-% $\forall tr,e,p. ~\forall t \in inds ~tr. ~(~P \rightarrow ~(~Q \rightarrow ~(~R \rightarrow ~(~S \rightarrow ~(~\forall i. ~T \rightarrow U~)~)~)~)$. where 
-% \begin{itemize}
-% \item
-% $tr: Trace$;~$e:Event$;~$p:nat1$;~$t:nat$
-% \item
-% $P = t \in set~(inds~tr)$
-% \item
-% $Q = (tr(t)=e)$
-% \item
-% $R = (t+p) ~<=~ (len~tr)$
-% \item
-% $S = (tr((t+p))=e) ~\land~ \forall i \in \{(t+1, \ldots, (t+p)-1\}. ~tr(i) \neq e )$
-% \item
-% $T = ((t+p) > (len~tr)~)$
-% \item
-% $U = (\forall i \in \{(t+1), \ldots, len~tr \}. ~i \in set (inds~tr)$
-% \end{itemize}
-
-% After taken negation of the formula, we got 
-% $\exists tr,e,p,t,i. ~(P \land Q \land Q \land S \land T \land \neg U~)$. In the negated case, the expected result of the negated PO is unsatisfiable, i.e., unsat. The
+The encoded PO8 (negated) was checked by the solver object in Z3py for satisfiability.
+The result was {\tt unsat}, which means that PO8 was proved.
 
@@
 PO3 says that the postcondition of {\tt FaultHeart} has to be satisfiable so that the specification of {\tt FaultHeart} is implementable.
-Fig.~\ref{fig:po3_case1} shows PO3 along with the specification of state and operation {\tt FaultyHeart}.
+Fig.~\ref{fig:po3_case1} shows PO3 along with the specification of state {\tt Pacemaker} and operation {\tt FaultyHeart}.
 
-% \medskip
 \begin{figure}[t]
@@
 \end{center}
+\vspace{-20pt}
 \end{figure}
-% \medskip
 
@@
 Therefore, PO3 is a simplified postcondition satisfiable obligation saying that there exists a {\tt Trace} {\tt tr} which satisfies the postcondition specified with only {\tt newstate}\footnote{How the {\tt newstate} is calculated from {\tt oldstate} is not specified in {\tt FaltyHeart}.}.
-The postcondition of {\tt FaultyHeart} calls the function {\tt Periodic} with three inputs.
+The postcondition of {\tt FaultyHeart} calls the function {\tt Periodic} twice with three arguments.
 {\tt Periodic} shown below is an explicitly specified function and returns a boolean value.
-Therefore, state {\tt Pacemaker}, postcondition of {\tt FaultyHeart} ({\tt post\_FaultyHeart}, and function {\tt Periodic} are the context information of PO3.
+Therefore, state {\tt Pacemaker}, {\tt post\_FaultyHeart} (postcondition of {\tt FaultyHeart}), and function {\tt Periodic} are the context information of PO3.
 
@@
 Though the specification of {\tt Periodic} is relatively long, the encoding of {\tt Periodic} is similar to the encoding of PO8.
-It is easy to observe similar computations on the trace {\tt tr} such as {\tt in set inds tr} and {\tt len tr}, and we can encode them in the same way as handling PO8.
-Note that the type {\tt Trace} is again needed in proving PO3 so that the encoding in PO8 above can be used as constraints here.
+It is easy to observe similar computations on the trace {\tt tr} such as {\tt in set inds tr} and {\tt len tr}, and we can encode them in the same way as encoding of PO8.
+Note that the type {\tt Trace} is again needed in proving PO3 so that the corresponding encoding in PO8 can be used again as constraints here.
 In the case of PO3, we did not apply negation to the formula since existence quantifier in the prefix of a formula is preferred in Z3.
@@
 \label{tbl:result1}
+\vspace{-20pt}
 \end{table}
@@
 The system state monitors the calling status of users and the connecting status among users.
-Overture tool generated twenty-seven POs, and we selected PO1 and PO14 for report the case study.
+Overture tool generated 27 POs, and we selected PO1 and PO14 to report the case study.
 
@@
 PO1 shown in Fig.~\ref{fig:po1_case2} is a legal map application obligation that checks whether maps in the system state are applicable.
-In PO1, types of the state {\tt Exchange} which is a record type of {\tt status} and {\tt calls} is used.
-Therefore, the context information of PO1 is the state including its invariant.
-Fig.~\ref{fig:types_case2} shows the definitions of state {\tt Exchange} with corresponding quote types.
+In PO1, state {\tt Exchange} which is a record type of {\tt status} and {\tt calls} is mentioned.
+Therefore, the context information of PO1 is the state {\tt Exchange} including its invariant.
+Fig.~\ref{fig:types_case2} shows the specification of state {\tt Exchange} with corresponding quote types.
 The state {\tt Exchange} contains two maps: {\tt status} indicates the discrete calling status of users, and is a map from {\tt Subscriber} to {\tt Status}; {\tt calls} models the connection relation of users, i.e., which users are connected, and is a map from {\tt Subscriber} to {\tt Subscriber}.
@@
 This means if a user is calling another user, both of the users can not have other connections to third party users.
-Also, {\tt Exchange} has state invariant that is specified for the constraints about connections of users and their calling status.
-At this point, the context information of PO1 is determined and we may proceed to encoding.
+Also, {\tt Exchange} has state invariant specifying the relations between connections of users and their calling status.
+At this point, the context information of PO1 is determined.
 
@@
 \end{center}
+\vspace{-10pt}
 \end{figure}
@@
 \end{center}
+\vspace{-20pt}
 \end{figure}
 
-The strategy of encoding the types shown in Fit.~\ref{fig:types_case2} is (1) applying user-defined types in Z3 to encode the quote types; (2) applying uninterpreted functions to encode map type.
+The strategy of encoding the types shown in Fig.~\ref{fig:types_case2} is (1) applying user-defined types in Z3 to encode the quote types; (2) applying uninterpreted functions to encode map type.
 In Fig.~\ref{fig:encode_types_case2}, we demonstrate the encoding of selected types.
+In Fig.~\ref{fig:encode_types_case2}, the type {\tt Recipient} is defined as an user-defined type with values {\tt WR} and {\tt SR}.
+Type {\tt Status} is further defined as a user-defined type including {\tt Recipient} and {\tt Initiator}.
+Though {\tt Status} is actually an union type combines {\tt Initiator} and {\tt Recipient}, and can be considered an enumeration, we encoded {\tt Status} using user-defined type style in Z3py.
+The reason is that we prefer the subtypes {\tt Initiator} and {\tt Status} can be distinguishable in Z3.
 
@@
 \end{center}
+\vspace{-20pt}
 \end{figure}
 
-In Fig.~\ref{fig:encode_types_case2}, the type {\tt Recipient} is defined as an user-defined type with values {\tt WR} and {\tt SR}.
-Type {\tt Status} is further defined as a user-defined type including {\tt Recipient} and {\tt Initiator}.
-Though {\tt Status} is actually an union type combines {\tt Initiator} and {\tt Recipient}, and can be considered an enumeration, we encoded {\tt Status} using user-defined type style in Z3py.
-The reason is that we prefer the subtypes {\tt Initiator} and {\tt Status} can be distinguishable in Z3.
-
-Note that since VDM uses partial functions, for the type to be declared as the range of map, i.e., {\tt Subscriber} and {\tt Status}, we defined their lifted type in Z3py to take into consideration the case that the value to a key of a map is undefined.
-Also, though {\tt Subscriber} is of token type which does not have specific values when declared, we treated token types as quote types with predefined domain of values such as {\tt S1}, {\tt S2}, and so on.
+Because of VDM's loic of partial functions, for the type to be declared as the range of map, i.e., {\tt Subscriber} and {\tt Status}, we defined their lifted types in Z3py to take into consideration the case that the value to a key of a map is undefined.
+Also, though {\tt Subscriber} is of token type which does not have specific values when declared, we treated token types as quote types with predefined domain of values such as {\tt S1}, {\tt S2}.
 The two maps {\tt status} and {\tt calls} in the state {\tt Exchange} are encoded as uninterpreted functions and the state invariant is encoded by the encoded (and lifted) types mentioned above.
@@
 The condition that the range value of a map is defined or not is essential here.
-For example, to encode {\tt status(calls(i)) = <WR>} (line 13 of Fig.~\ref{fig:types_case2}), we have ensure the condition that {\tt calls(i)} has to be defined so that additional constraint {\tt calls(i)!=Subscriber\_lift.NDF} (line 9 of Fig.~\ref{fig:encode_state_inv_case2}).
+For example, to encode {\tt status(calls(i)) = <WR>} (line 13 of Fig.~\ref{fig:types_case2}), we have to ensure the condition that {\tt calls(i)} is defined so that additional constraint {\tt calls(i)!=Subscriber\_lift.NDF} (line 9 of Fig.~\ref{fig:encode_state_inv_case2}) is added.
 
@@
 Thus, we had to add  the conditions about {\tt inmap} as context information: for any two different users, either both users calls no one (undefined) or they are calling different users.
+The latter case also includes the situation that only one of the two users is connecting to a third party user and the other user is calling no one.
 
@@
 
-The encoding of PO1 is done by encoding the negated PO1.
-As the following code shows, outside forall of {\tt Exchange} of PO1 is eliminated by the negation.
+The encoding of PO1 is done by encoding the negated formula of PO1.
+As the following code shows, the outside forall in the formula of PO1 is eliminated by obligation negation.
 The result was {\tt unsat} which means PO1 was proved.
@@
 
-PO14 is an enumeration map injectivity obligation of the operation {\tt Answer} of the system.
+PO14 shown in Fig.~\ref{fig:PO14_case2} is an enumeration map injectivity obligation of the operation {\tt Answer} of the system.
 PO14 has free variables, {\tt r} and {\tt status}, in its formula.
-These free variables are treated as the bounded by forall quantifier.
+These free variables are treated as forall quantifier bounded.
 There are also intermediate variables in the formula such as {\tt m1}, {\tt m2}, {\tt d3}, and {\tt d4}.
-Note that the inverse function of a map, and the range restriction function {\tt :>} are on {\tt calls} and {\tt status} separately.
-(Recall that {\tt calls} is a one-to-one map {\tt inmap}.) The context information of PO14 includes the two maps {\tt status} and {\tt calls} but not the state {\tt Exchange}.
-Therefore, only the two maps have to be encoded.
-The state invariant is excluded in the encoding.
-Note that the constraints for {\tt calls} of type {\tt inmap} is included in the context information of PO14, and the encoding in PO1 can be reused here.
+Note that the map inverse function {\tt inverse} (line 4) and the range restriction operator `{\tt :>}' (line 1) are applied on {\tt calls} and {\tt status} separately.
+(Recall that {\tt calls} is a one-to-one map {\tt inmap}.) The context information of PO14 includes the two maps {\tt status} and {\tt calls} but not the state {\tt Exchange} because {\tt Exchange} is not mentioned in PO14.
+Therefore, only the two maps have to be encoded, and the state invariant is excluded in the encoding.
+The encoding of the constraints for {\tt calls} of type {\tt inmap} described in solving PO1 can be reused here.
 
+\begin{figure}[t]
+\begin{center}
 \begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small]
+\begin{Verbatim}[fontsize=\small,numbers=left]
 (r in set (dom (status :> {<WR>}))) =>
@@
 \end{mdframed}
+\vspace{-10pt}
+\caption{Encoding of state invariant of Telephone Exchange}
+\label{fig:PO14_case2}
+\end{center}
+\vspace{-25pt}
+\end{figure}
 
@@
 
-For the map inverse function, we chose not to define a new map as the inverse of {\tt calls}, but encoded {\tt m1 = \{(inverse calls)(r) |-> <SI>\}} as a case of {\tt m1 in set \{\{r |-> <SR>\}, \{(inverse calls)(r) |-> <SI>\}\}}.
+For the map inverse function, we did not define a new map as the inverse of {\tt calls}, but encoded {\tt m1 = \{(inverse calls)(r) |-> <SI>\}} as a case of {\tt m1 in set \{\{r |-> <SR>\}, \{(inverse calls)(r) |-> <SI>\}\}}.
 The case was encoded as the collection of several conditions.
@@
 
-In above code, the first two conditions (lines 1-4) promise that (1) the part of map, where key {\tt i} is used and inverse is applied, has to be defined; (2) the domain key of the inversed map, i.e., the range value of the original map, is {\tt r}.
+In above code, the first two conditions (lines 1-4) promise that (1) the part of map, where key {\tt i} is used and {\tt inverse} is applied, has to be defined; (2) the domain key of the inversed map, i.e., the range value of the original map, is {\tt r}.
 Then {\tt m1(i)} where key {\tt i} is referred should be {\tt <SI>} (lines 6-7).
-(3) the last condition guarantees that {\tt m1} should be undefined for all indexes except {\tt i} (lines 9-10) because {\tt m1} is a single map that has only one key and one value.
+(3) the last condition guarantees that {\tt m1} should be undefined for all indices except {\tt i} (lines 9-10) because {\tt m1} is a single map that has only one key and one value.
 
@@
 Z3 returned a model that satisfies the negated formula of PO14.
-After examined the counterexample returned by Z3, we found that {\tt calls} in the returned model is a map where users calls themselves.
-This does not violate the type constraint of {\tt inmap} encoded above, but an user is not allowed to call him/her-self in a telephone exchange system.
-Therefore, the checking of PO14 suggested that the VDM-SL model should add a constraint on {\tt calls}.
-As an example of correction, we may add the constraint below as the type invariant of {\tt calls} or add it to the state invariant\footnote{We have added the constraint and checked again, and the PO was proved successfully.}.
+After examined the model returned by Z3, we found that {\tt calls} in the returned model, {\tt calls} is a map where users calls themselves.
+This does not violate the type constraint of {\tt inmap}, but an user is not allowed to call him/her-self in a telephone exchange system.
+Therefore, the checking of PO14 suggested that we should add a constraint on {\tt calls} to the VDM-SL model.
+We may add the constraint below as the type invariant of {\tt calls} or add it to the state invariant\footnote{We have added the constraint and checked again, and PO14 was proved successfully.}.
 
@@
 In this case study, we did not check all the twenty-seven POs since the encoding demonstrated in this paper could not cover all POs.
-The main issue is that the maps are encoded as uninterpreted functions so that {\tt forall} quantifier cannot be applied.
-This results that we could not check the postcondition satisfiable obligations of this model\footnote{Besides postcondition satisfiable obligations, there are two POs with type check function {\tt is\_(name, type)}, for which Z3 encoding is not available and then skipped.}.
+The main issue is that the maps are encoded as uninterpreted functions, which cannot be used with {\tt forall} quantifier.
+This results that we could not check the postcondition satisfiable obligations of this model with the current encoding\footnote{Besides postcondition satisfiable obligations, there are two POs with type check function {\tt is\_(name, type)}, for which Z3 encoding is not available and then skipped.}.
 After all, we successfully checked 16 out of all 27 POs for the telephone exchange VDM-SL model.
+The results of 16 POs checked of the Telephone Exchange VDM-SL model is showed in Table~\ref{tbl:result2} with information of whether negation is applied, and the time used.
 
-\subsection{Discussion}
+% total check result (table)
+%
+\begin{table}[htb]
+\begin{center}
+\begin{tabular}{|c|c|r|r|}
+\hline
+PO\#	&	negated	&	result	&	time (sec.) \\ \hline
+1		&	Y		&	unsat	&	0.032 \\ \hline
+2		&	Y		&	unsat	&	0.031 \\ \hline
+3		&	Y		&	unsat	&	0.031 \\ \hline
+4		&	Y		&	unsat	&	0.032 \\ \hline
+5		&	Y		&	unsat	&	0.03 \\ \hline
+6		&	Y		&	unsat	&	0.047 \\ \hline
+7		&	N		&	sat		&	0.031 \\ \hline
+9		&	Y		&	unsat	&	0.047 \\ \hline
+12		&	Y		&	sat		&	0.015 \\ \hline
+14		&	Y		&	sat		&	0.063 \\ \hline
+17		&	Y		&	sat		&	0.016 \\ \hline
+18		&	Y		&	unsat	&	0.047 \\ \hline
+20		&	Y		&	sat		&	0.016 \\ \hline
+21		&	Y		&	unsat	&	0.046 \\ \hline
+23		&	Y		&	sat		&	0.015 \\ \hline
+25		&	Y		&	sat		&	0.063 \\ \hline
+\end{tabular}
+\end{center}
+\caption{Telephone Exchange Result}
+\label{tbl:result2}
+\vspace{-20pt}
+\end{table}
 
@@
 
-Points of discussion:
-(1) Automation: Is the encoding can be automated? partly or fully? 
-(2) Patterns: What patterns are repeated used in encoding?
+\subsection{Discussion}
+
+As an overall result of our case studies, the objective of our approach that to discharge POs of a VDM-SL model as many as possible is achieved.
+In the first case study (Abstracted Pacemaker), we discharged all nine POs; in the second case study, we discharged 16 POs our of total 27 POs (about 60\%).
+To increase the percentage of discharging POs, we may further apply different encoding, for example, encoding maps using arrays in Z3, to solve remaining POs in the Telephone Exchange case study.
+
+% Usability and Applicability of the Encode Style
+The two case studies demonstrated the application of the encode strategy described in Section~\ref{section:encode-strategy}.
+From our experiences, the cost and complexity of encoding varies from specifications in a VDM-SL model.
+Usually, the complexity of encoded formulas usually depends on the complexity of VDM-SL types.
+For example, natural number based types such as {\tt seq of nat} or {\tt map from nat to nat} have less complexity, and one may encode and solve this kind of types with SMT solvers without difficulty.
+On the contrary, for VDM types such as a map of sequences or a set of maps, the encoding may be difficult and the complexity of the encoded formulas may grow a lot.
+
+The encoding demonstrated in the two case studies also showed some encoding patterns.
+For example, the encoding of union type and quote type with user-defined types or enumeration type, and the encoding of maps with uninterpreted functions, are useful to encoding POs of other VDM-SL models.
+Also, the encoding patterns like the constraints on arrays to represent a sequence, set inclusion of indices of a sequence, or set inclusion of domain/range of a map, are also reusable.
+In our case studies, these encoding patterns are actually applied multiple times.
+
+% Automation of the encoding style and strategy
+Though the encoding patterns are useful, applying the encoding patterns still depends on specification of VDM models because the encoding patterns are not fully systemic or syntax-based.
+Considering automated encoding of POs with related context information, the encoding of types may be fine, but encoding expressions remains difficult and rely on manually applying encoding patterns.
+
+% Scalability and Efficiency}
+Our case studies also showed the efficiency of SMT solvers can do great help in discharging POs.
+Also, the counterexample (model) generation of SMT solvers provides helpful hints for revision and correction of VDM models.
+As reported in Section~\ref{section:case2}, encoding and solving PO14 of the telephone exchange VDM-SL model found an inconsistency and the counterexample provided a guidance to dissolving the inconsistency.
+Finally, the scalability of our approach is not serious because we only encode the formula of a PO and its context information.
+The size of VDM-SL code to be encoded does not escalate with the size of a VDM-SL model.
+
+% \footnote{A previous work~\cite{kenneth:ifm2013} that translates VDM-SL to Alloy~\cite{kenneth:ifm2013} checked the VDM-SL model and found a different problem.
+The model (from the Overture repository) we used for case study is the fixed one which means that the inconsistency we found was not found in~\cite{kenneth:ifm2013}.}
+
+% Points of discussion:
+% (1) Automation: Is the encoding can be automated? partly or fully? 
+% (2) Patterns: What patterns are repeated used in encoding?

### 2017-05-27 paper/vdm.tex
polished (all)

@@
 The VDM Specification Language (VDM-SL) is a higher-order language with formally defined syntax and semantics~\cite{Prehn:1991:LNCS551,Larsen1995585}.
-VDM provides various abstract data types: basic types such as booleans, natural numbers, and tokens; advanced types such as record, product, set, and map.
+VDM provides various abstract data types: basic types such as boolean, natural number, and token; advanced types such as record, product, set, and map.
 A state consists of typed variables.
@@
 The state invariant specifies that adjacent elements of {\tt commands} have to be different.
-It is specified with {\tt forall} quantifier and {\tt len} operation that returns the length of a sequence.
-Finally, the initial value of \texttt{commands} is defined to be the empty sequence.
+It is specified with {\tt forall} quantifier and {\tt len}, a function that returns the length of a sequence.
+Finally, the initial value of \texttt{commands} is defined to be an empty sequence.
 
@@
 \label{fig:module_sample}
-% \vspace{-25pt}
+\vspace{-20pt}
 \end{center}
@@
 The operation {\tt push\_cmd} is specified to accept a command {\tt CMD} or {\tt nil} and pushes the command as the head of {\tt commands}.
-The precondition demands that \texttt{command} is the empty sequence or its head is different from the argument.
-The postcondition is specified using the old state {\tt commands\textasciitilde} with postfixed tilde and says that the old state is the tail of the new state {\tt commands}.
-Observe that {\tt push\_cmd} is an implicit operation that only specifies the signature with pre- and post-conditions without any operation.
+The precondition demands that \texttt{commands} is an empty sequence or its head is different from the argument.
+The postcondition is specified using the old state, the state before the execution of {\tt push\_cmd}, that denoted as {\tt commands\textasciitilde} with postfixed tilde.
+The postcondition says that the old state is the tail of the new state {\tt commands}.
+Observe that {\tt push\_cmd} is an implicit operation that only specifies the signature with pre- and post-conditions.
 The function \texttt{times\_count} that counts the occurrences of commands is specified explicitly.

### 2017-05-27 paper/conclusions.tex
typo, grammar checked

@@
 In this paper, we have proposed an approach to validation of VDM models by discharging proof obligations (POs) with SMT solvers.
-In our approach, Overture tool is used as the proof obligation generator, and the Python API of Z3 SMT sover is used for encoding and solving formulas of generated POs.
+In our approach, Overture tool is used as the proof obligation generator, and the Python API of Z3 SMT solver is used for encoding and solving formulas of generated POs.
 Encoding strategies including steps of encoding as well as encoding patterns are demonstrated, and two case studies are reported.
 From our experiences, it is recognized that discharging POs using SMT solvers is efficient, and the counterexample generation of SMT solvers when the discharging fails can provide useful guidance for correcting inconsistency of VDM models.
-We also discussed other issues such as applicability of encoding strategies considering the automation of our approach.
+We also discussed other issues such as the applicability of encoding strategies considering the automation of our approach.
 As the future work, we are gathering and organizing more encoding patterns for challenging the automation issue.

### 2017-05-27 paper/introduction.tex
typo, grammar checked

@@
 An explicit specification is operational; it describes \emph{how} system components work.
-An implicit specificiation on the other hand is declarative; it describes \emph{what} system components should work.
+An implicit specification, on the other hand, is declarative; it describes \emph{what} system components should work.
 Software specifications in VDM models define functionalities explicitly or implicitly according to software requirements.
-VDM has a strong record of applications for design and specification of software systems in industry~\cite{Larsen:2007:RIA:2227886.2227894,DBLP:conf/fm/KuritaCN08,DBLP:journals/ijsi/KuritaN09}.
-The well-known basic benefits of using VDM to describe software specifications are from the accuracy and unambiguity of VDM, which is common for other formal methods like B.
+VDM has a strong record of applications for design and specification of software systems in the industry~\cite{Larsen:2007:RIA:2227886.2227894,DBLP:conf/fm/KuritaCN08,DBLP:journals/ijsi/KuritaN09}.
+The well-known benefits of using VDM to describe software specifications are from the accuracy and unambiguity of VDM, which is common for other formal methods like B.
 
@@
 To validate a VDM model, existing tools such as Overture~\cite{Larsen:2010:OII:1668862.1668864} and VDMTools provide static consistency check regarding the syntax and type constraints of VDM specifications.
-The semantic validation of VDM specifications are established by discharging proof obligations (POs)~\cite{AL:97:POGV} generated by the above tools.
+The semantic validation of VDM specifications is established by discharging proof obligations (POs)~\cite{AL:97:POGV} generated by the above tools.
 POs are logic formulas characterizing consistency of VDM specifications.
@@
 Discharging POs of VDM models is essential to guarantee the consistency of the formal specifications.
-For a VDM model, the number of POs is usually large compare to the size of the model.
+For a VDM model, the number of POs is usually large compared to the size of the model.
 For example, the sample VDM-SL model shown in Fig.~\ref{fig:module_sample} has less than thirty lines, and Overture generates nine POs for the model.
@@
 All of them have to be discharged to guarantee consistency of software specifications.
-Manually discharging so many POs with theorem provers however can be a daunting task in practice.
-Furthermore, both errorneous VDM specifications and bad proof strategies induce un-dischargeable POs.
+Manually discharging so many POs with theorem provers, however, can be a daunting task in practice.
+Furthermore, both erroneous VDM specifications and bad proof strategies induce un-dischargeable POs.
 Theorem provers provide little information when they fail to discharge POs.
@@
 In this paper, we apply SMT solvers to discharge POs of VDM specifications.
-Here we focus on VDM-SL models and the objective is to discharge POs of an VDM-SL model as many as possible through SMT solvers.
+Here we focus on VDM-SL models and the objective is to discharge POs of a VDM-SL model as many as possible through SMT solvers.
 In our approach, we encode each POs with its context information such as type and state constraints as a first-order formula, and then prove the PO by sending the formula to SMT solvers.
@@
 
-More specifically, in this approach we chose the Overture tool as the PO generator.
+More specifically, in this approach, we chose the Overture tool as the PO generator.
 For the POs generated by the Overture tool, we encode and prove POs of VDM-SL models by solving SMT formulas with Python API of Z3~\cite{MB:08:ZSS}.
 Z3 is a popular SMT solver widely used in software verification and its Python API provides flexibility in constructing SMT formulas~\cite{BarFT-SMTLIB}.
-In order to encode POs, built-in VDM types such as sequences and maps must be characterized by theories in Z3.
+To encode POs, built-in VDM types such as sequences and maps must be characterized by theories in Z3.
 We conducted case studies and the results showed that our approach can efficiently discharge the significant part of POs of a VDM-SL model.
@@
 
-The structure of this paper is as follows: Section~\ref{section:vdm} briefly introduce VDM with a sample VDM-SL example; Section~\ref{section:proof-obligations} describes the POs of VDM and explains the POs generated by the Overture tool; section~\ref{section:encode-strategy} describes our strategy of encoding VDM expressions with Python API of Z3; section~\ref{section:case-studies} demonstrates case studies with discussions; section~\ref{section:related-work} compares our approach to related work; section~\ref{section:conclusions} concludes this paper and points out the future work.
+The structure of this paper is as follows: Section~\ref{section:vdm} briefly introduce VDM with a sample VDM-SL example; Section~\ref{section:proof-obligations} describes the POs of VDM and explains the POs generated by the Overture tool; Section~\ref{section:encode-strategy} describes our strategy of encoding VDM expressions with Python API of Z3; Section~\ref{section:case-studies} demonstrates case studies with discussions; Section~\ref{section:related-work} compares our approach to related work; Section~\ref{section:conclusions} concludes this paper and points out the future work.
 

### 2017-05-27 paper/obligations.tex
typo, grammar checked

@@
 % Generation information of POs
-For a VDM model, a proof obligation (PO) is a statement that must be proved in order to ensure the consistency of the model.
+For a VDM model, a proof obligation (PO) is a statement that must be proved to ensure the consistency of the model.
 A PO contains a predicate with its context information as shown below.
-The context information are code segments of the VDM model that relate to the predicate to be proved.
+The context information is from code segments of the VDM model that relate to the predicate to be proved.
 
@@
 PO1 is called ``legal sequence application obligation'' that relates to the sequence involved expression (line 9-10) of the VDM-SL model {\tt CMDS}).
-PO1 says that for all k in the set from 1 to {\tt len s.commands -1}, k should be always within the bound of defined indices of the sequence {\tt s.commands}.
-Also, it can be recognized that if we use {\tt len s.commands} instead of {\tt len s.commands -1} the proof will fail.
-Note that in VDM, a sequence are indexed from 1 to its length.
+PO1 says that for all k in the set from 1 to {\tt len s.commands -1}, k should always be within the bound of defined indices of the sequence {\tt s.commands}.
+Also, it can be recognized that if we use {\tt len s.commands} instead of {\tt len s.commands -1} the proof will fail since a sequence in VDM is indexed from 1 to its length.
 Also note that in PO1, the state {\tt CMDS'S} is mentioned, and we have to include the state invariant as well though it is not explicitly described in PO1.
@@
 PO7 is an operation postcondition satisfiable obligation for {\tt push\_cmd} (line 15-18).
-PO7 says that for every input {\tt a:CMD} and {\tt oldstate}, if the precondition whch takes inputs {\tt a} and {\tt oldstate} is satisfied, the postcondition which takes inputs {\tt a}, {\tt oldstate}, and {\tt newstate} should be satisfied.
+PO7 says that for every input {\tt a:CMD} and {\tt oldstate}, if the precondition which takes inputs {\tt a} and {\tt oldstate} is satisfied, the postcondition which takes inputs {\tt a}, {\tt oldstate}, and {\tt newstate} should be satisfied.
 Note that instead of the specification of precondition, postcondition, and state in {\tt CMDS} module, special names {\tt oldstate}, {\tt newstate}, and {\tt post\_push\_cmd} are used.

### 2017-05-27 paper/related.tex
typo, grammar checked

@@
 G.
-Larsen worked on proof obligation generation and discharging related to termination of recursive functions~\cite{Ribeiro2010}.
+Larsen worked on proof obligation generation and discharging related to the termination of recursive functions~\cite{Ribeiro2010}.
 
 % Work related to discharging POs using theorem provers. 
-There are several researches focusing on translation from VDM to theorem provers and discharging POs of VDM with theorem provers.
+There are researches focusing on translation from VDM to theorem provers and discharging POs of VDM with theorem provers.
 S.
@@
 D.
-Cuto et al.~\cite{CFP:14:TVCSAP} used the proof obligation generation mechanism of Overture tools for translating CML, a modeling language combing VDM-SL and CSP, to Isabella/HOL for automated proof.
-Recently, work on improvement of translating VDM-SL to Isabella/HOL are proposed~\cite{CT:15:EOCGTIS}.
+Couto et al.~\cite{CFP:14:TVCSAP} used the proof obligation generation mechanism of Overture tools for translating CML, a modeling language combing VDM-SL and CSP, to Isabella/HOL for automated proof.
+Recently, work on improvement of translating VDM-SL to Isabella/HOL was proposed~\cite{CT:15:EOCGTIS}.
 
 % Comparison to our approach
-The theorem prover based approaches has common issues: (1) the translation from VDM-SL to theorem provers may be errorneous since the translation applies on the whole VDM-SL model; (2) The proof process is usually complicated and requires an expert even with the help of proof tactics; (3) Theorem provers provide little information when they fail to discharge POs.
+The theorem prover based approaches has common issues: (1) the translation from VDM-SL to theorem provers may be erroneous since the translation applies to the whole VDM-SL model; (2) The proof process is usually complicated and requires an expert even with the help of proof tactics; (3) Theorem provers provide little information when they fail to discharge POs.
 Compare to our approach that discharges POs of VDM-SL with SMT solvers, we adopt the efficiency of SMT solvers and the encoding does not require to be applied on the whole VDM-SL model.
@@
 Both Lausdahl's and Lin's work require types involved in a VDM-SL model being bounded, and there are issues for Alloy and SPIN to represent the rich and complex VDM-SL types.
-Compare to them, bounds are not necessary for SMT solvers in our approach and our encoding patterns can applied to complex types thanks for build-in theories in SMT sovers.
+Compare to them, bounds are not necessary for SMT solvers in our approach, and our encoding patterns can be applied to complex types thanks for built-in theories in SMT solvers.
 

### 2017-07-20 memocode_cr/strategy.tex
MEMOCODE'17 cr init

--- 2017-05-27 paper/strategy.tex
@@
 \item
-Encode context information: this step encodes the context information determined by the first step, and can be divided into several minor steps.
+Encode context information: this step encodes the context information determined by the first step and can be divided into several minor steps.
 \begin{enumerate}
@@
 \item
-If a PO only considers a part of the specification of a type, function, or an operations, it is only needed to encode the expressions that are directly used in the PO.
+If a PO only considers a part of the specification of a type, a function, or an operations, it is only needed to encode the expressions that are directly used in the PO.
 In other words, if a type, a function, or an operation is not directly mentioned, for example, a variable mentioned with {\tt name:type} style, or a function mentioned with {\tt fun\_name(arguments)} style, there is no need to encode the full specification of the type, the function, or the operation.
@@
 In this work, forall quantifier usually causes difficulties in encoding and solving when complex types are involved.
-Obligation negation described above may achieve the elimination of troublesome foall quantifiers.
+Obligation negation described above may achieve the elimination of troublesome forall quantifiers.
 On the other hand, obligation negation should be applied carefully not to introduce troublesome forall quantifiers from exists quantifiers.
@@
 In this case, {\tt i} is the encoding an instance of type {\tt nat} in VDM.
-Note that we cannot add constraints on types but on instance of types in Z3.
+Note that we cannot add constraints on types but instances of types in Z3.
 This means that if there are several variables of the same VDM type that Z3 does not have, instances of the same number of variables have to be declared in Z3 with constraints for every instance added to the solver object.

### 2017-07-20 memocode_cr/studies.tex
MEMOCODE'17 cr init

--- 2017-05-27 paper/studies.tex
@@
 The first case is a small VDM-SL model specifies an abstract pacemaker.
-The model is not a full specification and focuses on part of definitions and operations related to heartbeat signals detected and recorded in a pacemaker.
+The model is not a full specification and focuses on the part of definitions and operations related to heartbeat signals detected and recorded in a pacemaker.
 For this model of 58 lines, Overture generated nine POs.
@@
 {\tt Trace} is of type sequence of {\tt Event}, where {\tt Event} is of quote type with two values {\tt <A>} and {\tt <V>} that stand for the A-pace and V-pace of heartbeat signals.
-At this point, we have necessary context information needed for discharging PO8 and we may proceed to the next step: encoding context information.
+At this point, we have necessary context information needed for discharging PO8, and we may proceed to the next step: encoding context information.
 For PO3, we need to encode types used in PO8: {\tt nat1} and {\tt Trace}.
@@
 In the above encoding, {\tt Trace} is encoded as an array type in Z3 and this array type is by default indexed through the integer domain.
-An array type can not be a sequence type unless we add constraints regarding to sequences to limit an array to a sequence.
-The idea of adding constraints is the same as the encoding of {\tt nat1} but more complicate.
+An array type can not be a sequence type unless we add constraints regarding sequences to limit an array to a sequence.
+The idea of adding constraints is the same as the encoding of {\tt nat1} but more complicated.
 Fig.~\ref{fig:tr_constraint_case1} shows the encoding of the constraint to {\tt tr}, the instance of {\tt Trace} as an array.
@@
 \item
-The first constraint (line 4) says that all indices of {\tt tr} below 1 should be undefined because in VDM-SL, a sequence's index starts from 1.
+The first constraint (line 4) says that all indices of {\tt tr} less than 1 should be undefined because, in VDM-SL, a sequence's index starts from 1.
 \item
@@
 \item
-(lines 4-7) If {\tt len\_tr} returns 0, then {\tt tr} is an sequence with no element.
+(lines 4-7) If {\tt len\_tr} returns 0, then {\tt tr} is a sequence with no element.
 That is, {\tt tr[1]} has to be undefined so that all indices of {\tt tr} are then undefined based on previous constraints of {\tt tr}.
@@
 Note that {\tt hd} and {\tt tl tr} are operators of the head and tail of a sequence, and {\tt inds} is the operator of gathering indices of defined elements of a sequence and returns a set of natural numbers.
-Also note that in the encoding in Fig.~\ref{fig:po8_encoded_case1}, we did not explicitly encode {\tt hd}, {\tt tl}, and {\tt inds} and simplified the results of these operators as constraints related to ranges (upper and lower bound) of {\tt tr}.
+Also, note that in the encoding in Fig.~\ref{fig:po8_encoded_case1}, we did not explicitly encode {\tt hd}, {\tt tl}, and {\tt inds} and simplified the results of these operators as constraints related to ranges (upper and lower bound) of {\tt tr}.
 The encoded PO8 (negated) was checked by the solver object in Z3py for satisfiability.
@@
 The state {\tt Pacemaker} is specified as a pair of natural numbers {\tt aperiod} and {\tt vdelay}.
-The postcondition of {\tt FaultyHeart} is not specified with {\tt oldstate} which has suffixed tilde, i.e, {\tt aperiod\textasciitilde} and {\tt vdelay\textasciitilde}.
+The postcondition of {\tt FaultyHeart} is not specified with {\tt oldstate} which has a suffixed tilde, i.e, {\tt aperiod\textasciitilde} and {\tt vdelay\textasciitilde}.
 Therefore, PO3 is a simplified postcondition satisfiable obligation saying that there exists a {\tt Trace} {\tt tr} which satisfies the postcondition specified with only {\tt newstate}\footnote{How the {\tt newstate} is calculated from {\tt oldstate} is not specified in {\tt FaltyHeart}.}.
@@
 Though the specification of {\tt Periodic} is relatively long, the encoding of {\tt Periodic} is similar to the encoding of PO8.
-It is easy to observe similar computations on the trace {\tt tr} such as {\tt in set inds tr} and {\tt len tr}, and we can encode them in the same way as encoding of PO8.
+It is easy to observe similar computations on the trace {\tt tr} such as {\tt in set inds tr} and {\tt len tr}, and we can encode them in the same way as the encoding of PO8.
 Note that the type {\tt Trace} is again needed in proving PO3 so that the corresponding encoding in PO8 can be used again as constraints here.
@@
 The state {\tt Exchange} contains two maps: {\tt status} indicates the discrete calling status of users, and is a map from {\tt Subscriber} to {\tt Status}; {\tt calls} models the connection relation of users, i.e., which users are connected, and is a map from {\tt Subscriber} to {\tt Subscriber}.
-Note that {\tt calls} is specified as {\tt inmap} type which means an one-to-one map.
+Note that {\tt calls} is specified as {\tt inmap} type which means a one-to-one map.
 This means if a user is calling another user, both of the users can not have other connections to third party users.
@@
 Type {\tt Status} is further defined as a user-defined type including {\tt Recipient} and {\tt Initiator}.
-Though {\tt Status} is actually an union type combines {\tt Initiator} and {\tt Recipient}, and can be considered an enumeration, we encoded {\tt Status} using user-defined type style in Z3py.
+Though {\tt Status} is a union type combines {\tt Initiator} and {\tt Recipient}, which can be considered an enumeration, we encoded {\tt Status} using user-defined type style in Z3py.
 The reason is that we prefer the subtypes {\tt Initiator} and {\tt Status} can be distinguishable in Z3.
@@
 \end{center}
-\vspace{-20pt}
+\vspace{-15pt}
 \end{figure}
 
-Because of VDM's loic of partial functions, for the type to be declared as the range of map, i.e., {\tt Subscriber} and {\tt Status}, we defined their lifted types in Z3py to take into consideration the case that the value to a key of a map is undefined.
-Also, though {\tt Subscriber} is of token type which does not have specific values when declared, we treated token types as quote types with predefined domain of values such as {\tt S1}, {\tt S2}.
+Because of VDM's logic of partial functions, for the type to be declared as the range of the map, i.e., {\tt Subscriber} and {\tt Status}, we defined their lifted types in Z3py to take into consideration the case that the value of a key of a map is undefined.
+Also, though {\tt Subscriber} is of token type which does not have specific values when declared, we treated token types as quote types with a predefined domain of values such as {\tt S1}, {\tt S2}.
 The two maps {\tt status} and {\tt calls} in the state {\tt Exchange} are encoded as uninterpreted functions and the state invariant is encoded by the encoded (and lifted) types mentioned above.
@@
 Thus, we had to add  the conditions about {\tt inmap} as context information: for any two different users, either both users calls no one (undefined) or they are calling different users.
-The latter case also includes the situation that only one of the two users is connecting to a third party user and the other user is calling no one.
+Note that the latter case also includes the situation that only one of the two users is connecting to a third party user and the other user is calling no one.
 
@@
 There are also intermediate variables in the formula such as {\tt m1}, {\tt m2}, {\tt d3}, and {\tt d4}.
-Note that the map inverse function {\tt inverse} (line 4) and the range restriction operator `{\tt :>}' (line 1) are applied on {\tt calls} and {\tt status} separately.
+Note that the map inverse function {\tt inverse} (line 4) and the range restriction operator `{\tt :>}' (line 1) are applied to {\tt calls} and {\tt status} separately.
 (Recall that {\tt calls} is a one-to-one map {\tt inmap}.) The context information of PO14 includes the two maps {\tt status} and {\tt calls} but not the state {\tt Exchange} because {\tt Exchange} is not mentioned in PO14.
@@
 When encoding the formula of PO14, firstly, we negated PO14 and eliminated the forall quantifiers.
-To encode the negated formula of PO14, similar technique for  encoding {\tt set (dom status)} in PO1 can be applied.
-For the range restriction function, since the range is a singleton set with element value of {\tt <WR>}, we do not have to encode the set notations and only limit the map's range {\tt status(r)} to {\tt <WR>}.
+To encode the negated formula of PO14, a similar technique for  encoding {\tt set (dom status)} in PO1 can be applied.
+For the range restriction function, since the range is a singleton set with an element {\tt <WR>}, we do not have to encode the set notations and only limit the map's range {\tt status(r)} to {\tt <WR>}.
 Thus, {\tt r in set (dom (status :> \{<WR>\}))} is encoded as follows.
@@
 
-The result of checking PO14 (negated) ware expected to be {\tt unsat}, but we got {\tt sat} instead.
+The result of checking PO14 (negated) was expected to be {\tt unsat}, but we got {\tt sat} instead.
 Z3 returned a model that satisfies the negated formula of PO14.
-After examined the model returned by Z3, we found that {\tt calls} in the returned model, {\tt calls} is a map where users calls themselves.
-This does not violate the type constraint of {\tt inmap}, but an user is not allowed to call him/her-self in a telephone exchange system.
+After examined the model returned by Z3, we found that {\tt calls} in the returned model, {\tt calls} is a map where users call themselves.
+This does not violate the type constraint of {\tt inmap}, but a user is not allowed to call him/her-self in a telephone exchange system.
 Therefore, the checking of PO14 suggested that we should add a constraint on {\tt calls} to the VDM-SL model.
@@
 
-As an overall result of our case studies, the objective of our approach that to discharge POs of a VDM-SL model as many as possible is achieved.
+As an overall result of our case studies, the objective of our approach that to discharge POs of a VDM-SL model as many as possible is considered achieved.
 In the first case study (Abstracted Pacemaker), we discharged all nine POs; in the second case study, we discharged 16 POs our of total 27 POs (about 60\%).
@@
 For example, natural number based types such as {\tt seq of nat} or {\tt map from nat to nat} have less complexity, and one may encode and solve this kind of types with SMT solvers without difficulty.
-On the contrary, for VDM types such as a map of sequences or a set of maps, the encoding may be difficult and the complexity of the encoded formulas may grow a lot.
+On the contrary, for VDM types such as a map of sequences or a set of maps, the encoding may be difficult, and the complexity of the encoded formulas may grow a lot.
 
@@
 Also, the encoding patterns like the constraints on arrays to represent a sequence, set inclusion of indices of a sequence, or set inclusion of domain/range of a map, are also reusable.
-In our case studies, these encoding patterns are actually applied multiple times.
+In our case studies, these encoding patterns are applied multiple times.
 
@@
 Also, the counterexample (model) generation of SMT solvers provides helpful hints for revision and correction of VDM models.
-As reported in Section~\ref{section:case2}, encoding and solving PO14 of the telephone exchange VDM-SL model found an inconsistency and the counterexample provided a guidance to dissolving the inconsistency.
+As reported in Section~\ref{section:case2}, encoding and solving PO14 of the telephone exchange VDM-SL model found an inconsistency and the counterexample provided guidance to dissolving the inconsistency.
 Finally, the scalability of our approach is not serious because we only encode the formula of a PO and its context information.

### 2017-07-20 memocode_cr/vdm.tex
MEMOCODE'17 cr init

--- 2017-05-27 paper/vdm.tex
@@
 The VDM Specification Language (VDM-SL) is a higher-order language with formally defined syntax and semantics~\cite{Prehn:1991:LNCS551,Larsen1995585}.
-VDM provides various abstract data types: basic types such as boolean, natural number, and token; advanced types such as record, product, set, and map.
+VDM provides various abstract data types: basic types such as Boolean, natural number, and token; advanced types such as record, product, set, and map.
 A state consists of typed variables.
@@
 The precondition demands that \texttt{commands} is an empty sequence or its head is different from the argument.
-The postcondition is specified using the old state, the state before the execution of {\tt push\_cmd}, that denoted as {\tt commands\textasciitilde} with postfixed tilde.
+The postcondition is specified using the old state, the state before the execution of {\tt push\_cmd}, which is denoted as {\tt commands\textasciitilde} with a postfixed tilde.
 The postcondition says that the old state is the tail of the new state {\tt commands}.
@@
 The function \texttt{times\_count} that counts the occurrences of commands is specified explicitly.
-Instead of using loop, the function uses sequence comprehension to construct the map of type {\tt CMD\_times}.
+Instead of using a loop, the function uses sequence comprehension to construct the map of type {\tt CMD\_times}.
 Specifically, the resultant map has the domain $\{ \texttt{<R>}, \texttt{<L>} \}$.
@@
 That is, the resultant map gives the number of $\texttt{<R>}$ in the sequence $\texttt{a}$ on $\texttt{<R>}$.
-Simiarly, it gives the number of $\texttt{<L>}$ in $\texttt{a}$ on $\texttt{<L>}$.
+Similarly, it gives the number of $\texttt{<L>}$ in $\texttt{a}$ on $\texttt{<L>}$.
 This also shows the expression power of VDM in specification.

### 2017-07-31 memocode_cr/conclusions.tex
MEMOCODE'17 camera ready

--- 2017-05-27 paper/conclusions.tex
@@
 Encoding strategies including steps of encoding as well as encoding patterns are demonstrated, and two case studies are reported.
-From our experiences, it is recognized that discharging POs using SMT solvers is efficient, and the counterexample generation of SMT solvers when the discharging fails can provide useful guidance for correcting inconsistency of VDM models.
-We also discussed other issues such as the applicability of encoding strategies considering the automation of our approach.
+As initial results, it is recognized that discharging POs using SMT solvers is efficient, and the counterexample generation of SMT solvers is useful for correcting inconsistency of VDM models.
+Currently the encoding is done by hand.
+We are working on more concrete encoding patterns and automatic encoding tool for more case studies.
+% We also discussed other issues such as the applicability of encoding strategies considering the automation of our approach.
 As the future work, we are gathering and organizing more encoding patterns for challenging the automation issue.

### 2017-07-31 memocode_cr/discussion.tex
MEMOCODE'17 camera ready

As an overall result of our case studies, the results showed that our approach can efficiently discharge the significant part of POs of a VDM-SL model. In the first case study (Abstracted Pacemaker), we discharged all nine POs; in the second case study, we discharged 16 POs our of total 27 POs (about 59\%). To increase the percentage of discharged POs, we may apply different encoding, even limited to finite types such as BitVector, to solve remaining POs in the Telephone Exchange case study.

% Usability and Applicability of the Encode Style
From our experiences, the cost and complexity of encoding varies from specifications in a VDM-SL model. Usually, the complexity of encoded formulas depends on the complexity of VDM-SL types. For example, natural number based types such as {\tt seq of nat} or {\tt map from nat to nat} have less complexity, and one may encode and solve this kind of types with SMT solvers without difficulty. On the contrary, for VDM types such as a map of sequences or a set of maps, the encoding may be difficult, and the complexity of the encoded formulas may grow a lot so that Z3 may fail to check the satisfiability.

We have found some encoding patterns through the case studies. For example, the encoding of union type and quote type with user-defined types or enumeration type, and the encoding of maps and sequences with uninterpreted functions or arrays, are useful to encoding POs of VDM-SL models. Also, the encoding patterns like the constraints on arrays to represent a sequence, set inclusion of indices of a sequence, or set inclusion of domain/range of a map, are also useful. In our case studies, these encoding patterns are applied multiple times.

% Automation of the encoding style and strategy
Though the encoding patterns are useful, applying the encoding patterns still depends on the specification of VDM models because the encoding patterns are not systemic or syntax-based at this point. Considering automated encoding of POs with related context information, the encoding of types may be fine, but encoding expressions remains difficult and rely on manual assistances may be needed.

% conclusion and future work
As initial results, it is recognized that discharging POs using SMT solvers is efficient, and the counterexample generation of SMT solvers is useful for correcting inconsistency of VDM models. Currently, the encoding is done by hand. As for future work, we are working on systematic encoding patterns and automated encoding for conducting more case studies.

### 2017-07-31 memocode_cr/introduction.tex
MEMOCODE'17 camera ready

--- 2017-05-27 paper/introduction.tex
@@
 % General information about VDM
-The Vienna Development Method (VDM)~\cite{Jones:1990:SSD:94062,Fitzgerald:2005:VDO:1044891,Fitz:PGLarson:2009book} is a formal method which supports modeling and analysis of software systems at various levels of abstraction.
+The Vienna Development Method (VDM)~\cite{Jones:1990:SSD:94062} is a formal method which supports modeling and analysis of software systems at various levels of abstraction.
 A VDM specification can be explicit, implicit, or both.
@@
 Software specifications in VDM models define functionalities explicitly or implicitly according to software requirements.
-VDM has a strong record of applications for design and specification of software systems in the industry~\cite{Larsen:2007:RIA:2227886.2227894,DBLP:conf/fm/KuritaCN08,DBLP:journals/ijsi/KuritaN09}.
+% VDM has a strong record of applications for design and specification of software systems in the industry~\cite{Larsen:2007:RIA:2227886.2227894,DBLP:conf/fm/KuritaCN08,DBLP:journals/ijsi/KuritaN09}.
 The well-known benefits of using VDM to describe software specifications are from the accuracy and unambiguity of VDM, which is common for other formal methods like B.
@@
 % Validation methods of VDM
-To validate a VDM model, existing tools such as Overture~\cite{Larsen:2010:OII:1668862.1668864} and VDMTools provide static consistency check regarding the syntax and type constraints of VDM specifications.
+% To validate a VDM model, existing tools such as Overture~\cite{Larsen:2010:OII:1668862.1668864} and VDMTools~\cite{citeVDMtools} provide static consistency check regarding the syntax and type constraints of VDM specifications.
 The semantic validation of VDM specifications is established by discharging proof obligations (POs)~\cite{AL:97:POGV} generated by the above tools.
@@
 % More about proof obligations and the motivation: solve as many as possible POs
-Discharging POs of VDM models is essential to guarantee the consistency of the formal specifications.
+Discharging proof obligations (POs)~\cite{AL:97:POGV} of VDM models is essential to guarantee the consistency of the formal specifications.
 For a VDM model, the number of POs is usually large compared to the size of the model.
-For example, the sample VDM-SL model shown in Fig.~\ref{fig:module_sample} has less than thirty lines, and Overture generates nine POs for the model.
+% For example, the sample VDM-SL model shown in Fig.~\ref{fig:module_sample} has less than thirty lines, and Overture generates nine POs for the model. 
 Consequently, it is not uncommon to have hundreds of POs for large VDM models.
@@
 In this paper, we apply SMT solvers to discharge POs of VDM specifications.
-Here we focus on VDM-SL models and the objective is to discharge POs of a VDM-SL model as many as possible through SMT solvers.
+We focus on VDM-SL models and the objective is to discharge POs of a VDM-SL model as many as possible through SMT solvers.
 In our approach, we encode each POs with its context information such as type and state constraints as a first-order formula, and then prove the PO by sending the formula to SMT solvers.
@@
 
-More specifically, in this approach, we chose the Overture tool as the PO generator.
+We chose the Overture tool~\cite{Larsen:2010:OII:1668862.1668864} as the PO generator.
 For the POs generated by the Overture tool, we encode and prove POs of VDM-SL models by solving SMT formulas with Python API of Z3~\cite{MB:08:ZSS}.
@@
 To encode POs, built-in VDM types such as sequences and maps must be characterized by theories in Z3.
-We conducted case studies and the results showed that our approach can efficiently discharge the significant part of POs of a VDM-SL model.
+We conducted case studies with manual encoding, and the results showed that our approach can efficiently discharge the significant part of POs of a VDM-SL model.
 %  \footnote{POs generated by Overture and VDMTools are slightly different in the numbers and styles of POs}.
 
-The structure of this paper is as follows: Section~\ref{section:vdm} briefly introduce VDM with a sample VDM-SL example; Section~\ref{section:proof-obligations} describes the POs of VDM and explains the POs generated by the Overture tool; Section~\ref{section:encode-strategy} describes our strategy of encoding VDM expressions with Python API of Z3; Section~\ref{section:case-studies} demonstrates case studies with discussions; Section~\ref{section:related-work} compares our approach to related work; Section~\ref{section:conclusions} concludes this paper and points out the future work.
+% The structure of this paper is as follows: Section~\ref{section:vdm} briefly introduce VDM with a sample VDM-SL example; Section~\ref{section:proof-obligations} describes the POs of VDM and explains the POs generated by the Overture tool; Section~\ref{section:encode-strategy} describes our strategy of encoding VDM expressions with Python API of Z3; Section~\ref{section:case-studies} demonstrates case studies with discussions; Section~\ref{section:related-work} compares our approach to related work; Section~\ref{section:conclusions} concludes this paper and points out the future work.
 

### 2017-07-31 memocode_cr/obligations.tex
MEMOCODE'17 camera ready

% Objective: explain the semantics of Overture generated POs.
% Explain what is proof obligation in VDM. There are kinds of POs: domain, subtype, satisfiability, and termination. Any example?
% Explain POs generated by the Overture tool. Any example?

% Generation information of POs
For a VDM model, a proof obligation (PO) is a statement that must be proved to ensure the consistency of the model. A PO contains a predicate with its context information as shown below. The context information is from code segments of the VDM model that relates to the predicate to be proved.

\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\small]
PO: context information ==> predicate
\end{Verbatim}
\end{mdframed}

% Classifications of POs and sample POs from sample VDM-SL model
POs of VDM-SL can be classified type compatibility (subtype checking), domain checking, and satisfiability~\cite{AL:97:POGV,Vermolen:2010:PCV:1774088.1774608}. Type compatibility relates to types with invariant or subtypes; domain checking relates to the use of partial functions and partial operators; satisfiability relates postconditions of functions and operations.

% Example VDM-SL segment and corresponding POs with explanation
Below is a segment of VDM-SL model and two of the corresponding POs generated by the Overture tool. The VDM-SL specification is a state definition which consists of a sequence of commands {\tt <R>} or {\tt <L>} or {\tt nil}. The state invariant says that for every element of {\tt commands}, its next element must be assigned different value. PO1 is a state invariant satisfiable obligation which is a subtype checking for confirming that there exist values of state {\tt S} satisfying the state invariant. PO2 is a legal sequence application obligation which is a domain checking for confirming that in the state invariant, {\tt (s.commands)(k)}, the k-th element of {\tt commands} has to be defined. Though looked obvious, this kind of POs is important for VDM because of the use of partial functions in VDM.

\begin{mdframed}[roundcorner=5pt]
\begin{Verbatim}[fontsize=\small]
-- //spec. segment
state S of
  commands : seq of [<R> | <L>]
  inv s == forall k in set 
    {1,...,len s.commands - 1} &
    s.commands(k) <> s.commands(k+1)
  init p == p = mk_S([])
end

-- //PO1: state invariant satisfiable obligation
-- //     (subtype check)
(exists commands:seq of [<R> | <L>] &
  (forall k in set {1, ... ,((len (s.commands)) - 1)} &
    ((s.commands)(k) <> (s.commands)((k + 1)))
))

-- //PO2: legal sequence application obligation
-- //    (domain check)
(forall s:CMDS`S & 
  (forall k in set 
    {1, ... ,((len (s.commands)) - 1)} & 
    (k in set (inds (s.commands)))))
\end{Verbatim}
\end{mdframed}

### 2017-07-31 memocode_cr/related.tex
MEMOCODE'17 camera ready

% Development history of VDM POG.
% Proofs are essential to guarantee the consistency of VDM models~\cite{978-3-540-19813-0},
Proof obligation generation of VDM is firstly proposed by B. K. Aichernig and P. G. Larsen~\cite{AL:97:POGV}. Later, A. Ribeiro and P. G. Larsen worked on proof obligation generation and discharging related to the termination of recursive functions~\cite{Ribeiro2010}.
% Work related to discharging POs using theorem provers. 
The mainstream of discharging POs of VDM models rely on translation from VDM to theorem provers. The generation and discharging of POs are conducted on theorem provers such as PVS and Isabella/HOL. S. Agerholm~\cite{Agerholm1996} proposed a translation from a subset of VDM-SL to PVS for type checking. S. Maharaj and J. Bicarregui~\cite{632849} used Agerholm's translation to assist the verification of VDM-SL models and their refinements. Later, S. D. Vermolen et al.~\cite{Verm:2007:master,Vermolen:2010:PCV:1774088.1774608} utilized the parsing mechanism of the Overture tool and developed automated translation from a subset of VDM-SL to HOL theorem prover. 
% Vermolen's work also provided some tactics to support automated discharging of POs. L. D. Couto et al.~\cite{CFP:14:TVCSAP} used the proof obligation generation mechanism of the Overture tool for translating CML, a modeling language combing VDM-SL and CSP, to Isabella/HOL for automated proof. 
Recently, work on improvement of translating VDM-SL to Isabella/HOL was proposed~\cite{CT:15:EOCGTIS}.

% Comparison to our approach
In general, the theorem prover approach has two major issues: 
% (1) the translation from VDM-SL to theorem provers may be erroneous since the translation applies to the whole VDM-SL model; 
(1) The proof process is usually complicated and requires an expert even with the help of proof tactics; (2) Theorem provers provide little information when they fail to discharge POs. Compared to our approach, we adopt the efficiency of SMT solvers, and the counterexample generation of SMT solvers is helpful for locating the problem. However, encoding in our approach is manual at this point and still need improvements.

Theorem proving with SMT solvers is not a new idea. S. Merz and H. Vanzetto~\cite{Merz2012TLASMT} has proposed a backend using SMT solvers to assist discharging proof obligations of TLA{\thinspace}+{\thinspace}. D. Kr\"oning et al.~\cite{smtLSM2009} has proposed a theory of finit sets, lists, and maps for SMT-LIB. The proposed theory is general but limited in basic types of {\tt Int} and {\tt Real}. The authors only discussed how to realize the theory by reduction to existing SMT theories with a report of initial benchmarks\footnote{http://www.cprover.org/SMT-LIB-LSM/} on a Event-B case study. The idea of adopting existing theories of SMT is the same as our approach. However, we argue that a strategic encoding like our approach might is necessary for discharging VDM POs than introducing general theories because the infinite and undecidable nature of sets, lists, and maps in VDM.

% Work related to validation/verfication of VDM models with model checking.
% Verification and validation of VDM-SL models with model checking techniques are recently proposed. K. Lausdahl~ \cite{kenneth:ifm2013} proposed a semantics-preserving translation that constructs an Alloy model from a subset of VDM-SL. This work aims to support the validation of implicitly specified VDM-SL models with Alloy. H-H. Lin et al.~\cite{DBLP:conf/ftscs/LinOKA15} proposed an approach that combines SPIN model checker and the VDMJ interpreter for verifying explicitly specified VDM-SL models. Lausdahl's work is similar to our approach since Alloy uses SAT solvers, but SMT solvers in our approach can solve wider formulas with more efficiency. Both Lausdahl's and Lin's work require types involved in a VDM-SL model being bounded, and there are issues for Alloy and SPIN to represent the rich and complex VDM-SL types. Compare to them, bounds are not necessary for SMT solvers in our approach, and our encoding patterns can be applied to complex types thanks for built-in theories in SMT solvers.

### 2017-07-31 memocode_cr/strategy.tex
MEMOCODE'17 camera ready

@@
 % Z3 introduction, API introduction
-As mentioned in Section~\ref{section:introduction}, our approach encodes and proves VDM POs with the Z3 SMT solver.
+In this work, we encode and prove VDM POs with the Z3 SMT solver.
 Z3 is one of the popular SMT solver wildly used in verification of programs and software systems.
@@
 \item
-Determine the context information of the PO: The context information is the definitions specified in the VDM model that are involved in the obligation.
+Determine the context information of the PO.
+% : The context information is the definitions specified in the VDM model that are involved in the obligation.
 The context information may include types, functions, and operations.
@@
 \item
-Encode context information: this step encodes the context information determined by the first step and can be divided into several minor steps.
+Encode context information.
+% : this step encodes the context information determined by the first step and can be divided into several minor steps.
 \begin{enumerate}
@@
 \item
-For pre- and post-conditions of a function/operation specified in the VDM model, the encoding is needed if {\tt pre\_<name>} or {\tt post\_<name>} are used in the PO.
+If pre/post-conditions of functions/operations are used in a PO in the form of {\tt pre\_<name>} or {\tt post\_<name>}, encoding of pre/post-conditions is needed.
+%  are used in the PO.
 For example, PO7 of the {\tt CMDS} module described in Section~\ref{section:proof-obligations} mentioned both precondition and postcondition, so both conditions shall be encoded.
 \item
-If a PO only considers a part of the specification of a type, a function, or an operations, it is only needed to encode the expressions that are directly used in the PO.
-In other words, if a type, a function, or an operation is not directly mentioned, for example, a variable mentioned with {\tt name:type} style, or a function mentioned with {\tt fun\_name(arguments)} style, there is no need to encode the full specification of the type, the function, or the operation.
+If a PO only considers a part of the specification of a type, a function, or an operation, it is only needed to encode the expressions that are directly used in the PO.
+%  In other words, if a type, a function, or an operation is not directly mentioned, for example, a variable mentioned with {\tt name:type} style, or a function mentioned with {\tt fun\_name(arguments)} style, there is no need to encode the full specification of the type, the function, or the operation.
 \end{enumerate}
 \item
-Obligation negation and quantifier elimination
+Obligation negation and quantifier elimination.
 \begin{enumerate}
 \item
-Obligation negation: In model checking with SMT solvers, the assertion to be checked is usually negated, and the result of unsatmeans the proof of the assertion succeeds.
+Obligation negation: In model checking with SMT solvers, the assertion to be checked is usually negated, and the result of unsat means the proof of the assertion succeeds.
 For the result of sat which means the assertion can be violated, SMT solvers return a model as the counterexample that shows an example of how the assertion is violated.
-In this work, we also take negation on the PO to be discharged.
-However, the elimination of forall quantifier should be considered together when applying the negation to the PO.
+In this work, we also negate the PO to be discharged.
+However, the elimination of universal quantifier should be considered together when applying the negation to the PO.
 \item
-Elimination of forall quantifiers: POs of VDM usually have forall quantifiers.
-To solve the formulas with forall quantifiers in Z3, though Z3 accepts formulas with forall quantifiers, there are limitations applying forall quantifiers in Z3 so that by default we have to eliminate forall quantifiers before encoding.
-In this work, forall quantifier usually causes difficulties in encoding and solving when complex types are involved.
-Obligation negation described above may achieve the elimination of troublesome forall quantifiers.
-On the other hand, obligation negation should be applied carefully not to introduce troublesome forall quantifiers from exists quantifiers.
+Elimination of universal quantifiers: POs of VDM are usually with universal quantifiers.
+Though Z3 accepts formulas with universal quantifiers, there are limitations applying universal quantifier in Z3.
+Therefore, we eliminate universal quantifiers in the predicate of a PO before encoding.
+Obligation negation described above also achieves the elimination of universal quantifiers.
+% On the other hand, for POs with existential quantifiers, obligation negation is not needed for not introducing universal quantifiers. 
+However, for POs with both universal and existential quantifiers, or multiple universal quantifiers, it should be avoid introducing troublesome universal quantifiers, i.e., universal quantifiers on sequences, sets, and maps, from exists quantifiers.
 \end{enumerate}
@@
 % More details of encoding types, operators, and expressions
-VDM has rich types from natural number and quote to sequence and maps.
+VDM has rich types from natural number and quotes to sequences and maps.
 There are no direct matching for all VDM types to the types in Z3.
@@
 Based on our case studies, there are patterns of encoding that are frequently used for corresponding expression styles in VDM-SL models.
-
-% Demonstrate Z3Py encoding
-For more details, the Z3py encoding firstly needs to include the Z3 module and declare a solver object. We can add constraints of a PO to the solver object and solve the formulas for proof.
-
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
-  from z3 import *
-  s = Solver()
-  i = Int('i')
-  s.add(i>=0)
-\end{Verbatim}
-\end{mdframed}
-
-The above Python code imports the z3 module and create a solver object, then declare an integer {\tt i} with a constraint that {\tt i} is greater and equal to zero.
-In this case, {\tt i} is the encoding an instance of type {\tt nat} in VDM.
-Note that we cannot add constraints on types but instances of types in Z3.
-This means that if there are several variables of the same VDM type that Z3 does not have, instances of the same number of variables have to be declared in Z3 with constraints for every instance added to the solver object.
-The steps and strategies of encoding discussed above will be demonstrated with details in Section~\ref{section:case-studies} through the report of case studies.
-
-% Z3py encoding: taking negation and forall quantifier elimination
-% Besides the encoding of VDM types, 
-%  For example, if a VDM type is encoded as an uninterpreted function, the type cannot be used within forall quantifier.
-A simple way of quantifier elimination is negation the formula so the forall quantifier turns into exists quantifier which can be omitted in SMT formulas usually.

### 2017-07-31 memocode_cr/studies.tex
MEMOCODE'17 camera ready

@@
-This section gives studies that apply our approach to selected VDM-SL models and discusses the results\footnote{All the selected VDM-SL models can be found in Overture repository: {\tt http://overturetool.org/download/examples/VDMSL/}, and the encoding and results with Z3py can be founded in our repository: {\tt https://github.com/fmlab-iis/VDM-SMT}}.
+We have conduct two case studies\footnote{The VDM-SL models are from the Overture repository: {\tt http://overturetool.org/download/examples/VDMSL/}.
+The encoding of POs done by hand and their checking results can be founded in our repository: {\tt https://github.com/fmlab-iis/VDM-SMT}}: an abstract pacemaker and a telephone exchange system.
 
-\subsection{Abstract Pacemaker}
-\label{section:case1}
+% \subsection{Abstract Pacemaker}
+% \label{section:case1}
 
-The first case is a small VDM-SL model specifies an abstract pacemaker.
+\noindent{\bf Abstract Pacemaker.} ~~The first case is a small VDM-SL model specifies an abstract pacemaker.
 The model is not a full specification and focuses on the part of definitions and operations related to heartbeat signals detected and recorded in a pacemaker.
-For this model of 58 lines, Overture generated nine POs.
-Here we report the encoding and solving of obligations PO8 and PO3 below.
+For this model of 58 lines, the Overture tool generated nine POs.
+The encoding and checking of POs follow the encoding steps described in Section~\ref{section:encode-strategy}.
+In this section, we demonstrate some of the encoding patterns for encoding lists with arrays.
 
-\subsubsection{Proof Obligation 8}
+As the code shown below, in the abstract pacemaker model, the core types are {\tt Trace} and {\tt Event} since they show in almost all POs of the model.
+{\tt Trace} is of type sequence of {\tt Event}, where {\tt Event} is of quote type with two values {\tt <A>} and {\tt <V>} that stand for the A-pace and V-pace of heartbeat signals.
 
-PO8 shown in Fig.~\ref{fig:po8_case1} is a legal sequence application obligation of operation {\tt Pace}.
-PO8 requires that the formula specified with trace {\tt tr} of type {\tt Trace} be legal, i.e., computable to get its result.
-For more understanding of PO8, we may take a look on the definition of the operation {\tt Pace} as shown in Fig.~\ref{fig:pace_case1}.
-{\tt Pace} is an explicitly specified operation and takes three arguments of types {\tt Trace}, {\tt nat1}, and {\tt nat1}.
-It is easy to recognize that PO8 is checking about the sequence used in line 4 of PO8 in Fig.~\ref{fig:pace_case1}.
-Note that to discharge PO8, we do not need the full definition of {\tt Pace}.
-
-\begin{figure}[t]
-\begin{center}
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small]
-(forall tr:Trace, aperi:nat1, vdel:nat1,
-  oldstate & 
-    (forall i in set (inds (tl tr)) &
-      (((i mod aperi) = (vdel + 1)) =>
-        (i in set (inds tr)))
-))
-\end{Verbatim}
-\end{mdframed}
-\vspace{-10pt}
-\caption{Proof Obligation 8 of Abstract Pacemaker}
-\label{fig:po8_case1}
-\end{center}
-% \vspace{-20pt}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small]
-Pace: Trace * nat1 * nat1 ==> Trace
-Pace(tr,aperi,vdel) ==
-  return [nil] ^
-     [ if (i mod aperi = vdel + 1) and
-          tr(i) <> <V> 
-       then <V> else nil
-     | i in set inds tl tr];
-\end{Verbatim}
-\end{mdframed}
-\vspace{-10pt}
-\caption{Pace of Abstract Pacemaker}
-\label{fig:pace_case1}
-\end{center}
-\vspace{-20pt}
-\end{figure}
-
-As described in Section~\ref{section:encode-strategy}, the first step of encoding is to determine the context information of PO8.
-To decide what to include as the context information, we firstly examine the types used in PO8: {\tt nat1} and {\tt Pace}.
-{\tt nat1} is a type of natural numbers that excludes zero and starts from one.
-{\tt nat1} is a primitive type of VDM so we do not need more information about it.
-{\tt Pace} is a user-defined VDM type so we have check its specification in the VDM-SL model.
-Below shows the type definition of {\tt Trace}.
-Note that {\tt oldstate} bounded by the forall quantifier is never used in the predicate of PO8 and therefore will be ignored in the encoding.
-
+\medskip
 \begin{mdframed}[roundcorner=5pt]
@@
 \end{mdframed}
+\medskip
 
-{\tt Trace} is of type sequence of {\tt Event}, where {\tt Event} is of quote type with two values {\tt <A>} and {\tt <V>} that stand for the A-pace and V-pace of heartbeat signals.
-At this point, we have necessary context information needed for discharging PO8, and we may proceed to the next step: encoding context information.
-For PO3, we need to encode types used in PO8: {\tt nat1} and {\tt Trace}.
-{\tt nat1} is the natural number that excludes zero, so the encoding of {\tt nat1} is as follows.
-
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
-  aperi = Int('aperi')
-  s.add(aperi>=1)
-  vdel = Int('vdel')
-  s.add(vdel>=1)
-\end{Verbatim}
-\end{mdframed}
-
-As the above code shows, we defined two instances of integer in Z3, {\tt aperi} and {\tt vdel}.
-Recall that in Z3, we cannot add constraints to types but instances, so the constraint of greater and equal to one is added twice to the solver object for both {\tt aperi} and {\tt vdel}.
-The remaining type not encoded yet is {\tt Trace}, the sequence type.
-We defined {\tt Event} firstly since {\tt Trace} is a trace of {\tt Event}.
 The code below shows the encoding of {\tt Event} and {\tt Trace}.
@@
 
+\medskip
 \begin{mdframed}[roundcorner=5pt,shadow=true]
@@
 An array type can not be a sequence type unless we add constraints regarding sequences to limit an array to a sequence.
-The idea of adding constraints is the same as the encoding of {\tt nat1} but more complicated.
-Fig.~\ref{fig:tr_constraint_case1} shows the encoding of the constraint to {\tt tr}, the instance of {\tt Trace} as an array.
-Note again that the constraints are applied on {\tt tr}, the instance of {\tt Trace}, not the type {\tt Trace}.
-As shown in Fig.~\ref{fig:tr_constraint_case1}, we added three constraints:
+Below shows the encoding of the constraint to {\tt tr}, the instance of {\tt Trace} as an array.
+Note that in Z3, we cannot add constraints to types but instances.
+The code describes three constraints to reduce an array ({\tt tr}) to a sequence that starts from index 1 to its last element.
 
-\begin{figure}[t]
-\begin{center}
+\medskip
 \begin{mdframed}[roundcorner=5pt,shadow=true]
@@
 \end{mdframed}
-\vspace{-10pt}
-\caption{Encode constraints for {\tt tr:Trace} in Z3py}
-\label{fig:tr_constraint_case1}
-\end{center}
-\vspace{-20pt}
-\end{figure}
+\medskip
 
@@
 \item
-The first constraint (line 4) says that all indices of {\tt tr} less than 1 should be undefined because, in VDM-SL, a sequence's index starts from 1.
+The first constraint (line 4) says that all indices of {\tt tr} less than 1 should be undefined because a sequence's index starts from 1 in VDM.
 \item
-The second constraint (lines 6-13) says that if an index is defined, i.e., not undefined, all indices lower than it should be all defined.
+The second constraint (lines 6-13) says that if an index is defined, i.e., not undefined, all indices lower than it should be defined.
 \item
-The third constraint (line 15-22) says tat if an index is undefined, all indices greater than it should be all undefined.
+The third constraint (line 15-22) says tat if an index is undefined, all indices greater than it should be undefined.
 \end{itemize}
 
-From the above three constraints, we limited an array ({\tt tr}) to a sequence that starts from index 1 to its last element.
-Now we have encoded both types {\tt nat1} and {\tt Trace} used in PO8.
-Before proceed to encode the obligation, it is necessary to encode the function {\tt len} that computes the length of a sequence.
-As shown in Fig.~\ref{fig:tr_len_case1}, we defined an uninterpreted function {\tt len\_tr} for {\tt tr}, the instance of {\tt Trace}.
-Line 1 is the type declaration of {\tt len\_tr} which is an uninterpreted function taking an argument of a {\tt Trace} type and returning an integer type.
-Similar to defining constraints for {\tt tr:Trace}, we gave {\tt len\_tr} the constraint considering two cases: the length of a sequence can be equal or greater than 0.
+With another encoding for {\tt len} operator of sequences.
+We were able to prove all nine POs of the abstracted pacemaker model.
+We also applied some other encoding patterns to encode operators {\tt hd}, {\tt tl}, {\tt in set inds}.
+These operators can be encoded as constraints on the range of a sequence so that we do not have to declare corresponding instances of sequences or sets.
 
-\begin{figure}[t]
-\begin{center}
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small,numbers=left]
-len_tr=Function('len_tr',Trace,IntSort())
+\noindent{\bf Telephone Exchange.} ~~The second case study is a VDM-SL model specifies an abstracted telephone exchange system.
+In this model, the operations specify the events which can be initiated either by the system or by a subscriber (user) with implicit style.
+The system state monitors the calling status of users and the connecting status among users.
+The Overture tool generated 27 POs, and we managed to encode and prove 16 POs.
+The main reason of leaving some POs unproved is because of nested universal quantifiers.
 
-Or(
-  And(
-    len_tr(tr)==0,
-    tr[1]==NDF
-  ),
-  And(
-    len_tr(tr)>0,
-    tr[len_tr(tr)]!=NDF,
-    tr[len_tr(tr)+1]==NDF
-  )
-)
-\end{Verbatim}
-\end{mdframed}
-\vspace{-10pt}
-\caption{Encode {\tt len\_tr} for {\tt tr:Trace} in Z3py}
-\label{fig:tr_len_case1}
-\end{center}
-\end{figure}
+Instead of using arrays, we tried using uninterpreted functions to encode maps.
+We also applied encoding patterns mentioned in the abstract pacemaker case.
+In the checking, we found an inconsistency about the constraint regarding {\tt calls}, the map specifying relations between subscribers.
+The counterexamples generated by Z3 were helpful for suggesting how to correct the model.
 
-\begin{enumerate}
-\item
-(lines 4-7) If {\tt len\_tr} returns 0, then {\tt tr} is a sequence with no element.
-That is, {\tt tr[1]} has to be undefined so that all indices of {\tt tr} are then undefined based on previous constraints of {\tt tr}.
-\item
-(line 8-12) If {\tt len\_tr} returns an integer greater than 0, then {\tt tr} has defined elements till its length index {\tt len\_tr(tr)} such that the element at and after index {\tt len\_tr(tr)+1} should be undefined.
-\end{enumerate}
-
-At this point, we have defined types and constraints needed for encoding PO8.
-According to the steps described in Section~\ref{section:encode-strategy}, we now proceed to the next step: obligation negation.
-The negated PO8 is shown in Fig.~\ref{fig:po8_negation_case1}.
-Compare to the original PO8 in Fig.~\ref{fig:po8_case1}, the obligation negation also eliminated the forall quantifier.
-The negated PO8 can be recognized as the formula $\exists~[tr:Trace,aperi:nat1,vdel:nat1,i:nat] P \land Q \land \neg R$ where $P$, $Q$, and $\neg R$ can be encoded separately.
-
-\begin{figure}[t]
-\begin{center}
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small]
-Exists tr:Trace, aperi:nat1, vdel:nat1 &
-  (Exists i in set (inds (tl tr)) &
-     Not(
-      ((i mod aperi) = (vdel + 1)) =>
-       (i in set (inds tr))
-     )
-  )
-\end{Verbatim}
-\end{mdframed}
-\vspace{-10pt}
-\caption{Obligation negation of PO8}
-\label{fig:po8_negation_case1}
-\end{center}
-\vspace{-20pt}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small,numbers=left]
-  And(aperi>=1,vdel>=1)        #nat1
-  And(i>=1, i<=len_tr(tr)-1)   #P
-  i%aperi==vdel+1              #Q
-  Not(And(i>=1,i<=len_tr(tr))) #not R
-\end{Verbatim}
-\end{mdframed}
-\vspace{-10pt}
-\caption{Encoded formulas of PO8 in Z3py}
-\label{fig:po8_encoded_case1}
-\end{center}
-% \vspace{-20pt}
-\end{figure}
-
-Finally, we encoded the negated PO8 as the Z3py code shown in Fig.~\ref{fig:po8_encoded_case1}.
-The encoding has four constraints: two {\tt nat1} variables (line 1); encoding of $P$ ({\tt i in set (inds (tl tr)}) that is actually indices from 1 to {\tt len tr -1} (line 2); encoding of $Q$ ({\tt (i mod aperi) = (vdel + 1)}) (line 3); encoding of $\neg R$ ({\tt not(i in set (inds tr))}) (line 4).
-Note that {\tt hd} and {\tt tl tr} are operators of the head and tail of a sequence, and {\tt inds} is the operator of gathering indices of defined elements of a sequence and returns a set of natural numbers.
-Also, note that in the encoding in Fig.~\ref{fig:po8_encoded_case1}, we did not explicitly encode {\tt hd}, {\tt tl}, and {\tt inds} and simplified the results of these operators as constraints related to ranges (upper and lower bound) of {\tt tr}.
-The encoded PO8 (negated) was checked by the solver object in Z3py for satisfiability.
-The result was {\tt unsat}, which means that PO8 was proved.
-
-\subsubsection{Proof Obligation 3}
-
-PO3 of the abstracted pacemaker model is the postcondition satisfiable obligation of the implicitly specified operation {\tt FaultHeart} that generates a trace of heart pulse signals.
-PO3 says that the postcondition of {\tt FaultHeart} has to be satisfiable so that the specification of {\tt FaultHeart} is implementable.
-Fig.~\ref{fig:po3_case1} shows PO3 along with the specification of state {\tt Pacemaker} and operation {\tt FaultyHeart}.
-
-\begin{figure}[t]
-\begin{center}
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small,numbers=left]
-exists tr:Trace &
-  post_FaultyHeart(oldstate, tr, newstate)
-
-state Pacemaker of
-    aperiod : nat 
-    vdelay  : nat
-  init p == p = mk_Pacemaker(15,10)
-  end
-
-FaultyHeart() tr : Trace
-post len tr = 100 and
-  Periodic(tr,<A>,aperiod) and 
-  not Periodic(tr,<V>,aperiod);
-\end{Verbatim}
-\end{mdframed}
-\vspace{-10pt}
-\caption{PO3, State {\tt Pacemaker}, and operation {\tt FaultyHeart} of Abstraced Pacemaker}
-\label{fig:po3_case1}
-\end{center}
-\vspace{-20pt}
-\end{figure}
-
-The state {\tt Pacemaker} is specified as a pair of natural numbers {\tt aperiod} and {\tt vdelay}.
-The postcondition of {\tt FaultyHeart} is not specified with {\tt oldstate} which has a suffixed tilde, i.e, {\tt aperiod\textasciitilde} and {\tt vdelay\textasciitilde}.
-Therefore, PO3 is a simplified postcondition satisfiable obligation saying that there exists a {\tt Trace} {\tt tr} which satisfies the postcondition specified with only {\tt newstate}\footnote{How the {\tt newstate} is calculated from {\tt oldstate} is not specified in {\tt FaltyHeart}.}.
-The postcondition of {\tt FaultyHeart} calls the function {\tt Periodic} twice with three arguments.
-{\tt Periodic} shown below is an explicitly specified function and returns a boolean value.
-Therefore, state {\tt Pacemaker}, {\tt post\_FaultyHeart} (postcondition of {\tt FaultyHeart}), and function {\tt Periodic} are the context information of PO3.
-
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small]
-Periodic: Trace * Event * nat1 -> bool
-Periodic(tr,e,p) ==
-  forall t in set inds tr &
-   (tr(t) = e) =>
-   (t + p <= len tr =>
-   ((tr(t+p) = e and
-     forall i in set {t+1, ..., t+p-1} &
-       tr(i) <> e)) and
-    (t + p > len tr =>
-     forall i in set {t+1, ..., len tr} &
-       tr(i) <> e));
-\end{Verbatim}
-\end{mdframed}
-
-Though the specification of {\tt Periodic} is relatively long, the encoding of {\tt Periodic} is similar to the encoding of PO8.
-It is easy to observe similar computations on the trace {\tt tr} such as {\tt in set inds tr} and {\tt len tr}, and we can encode them in the same way as the encoding of PO8.
-Note that the type {\tt Trace} is again needed in proving PO3 so that the corresponding encoding in PO8 can be used again as constraints here.
-In the case of PO3, we did not apply negation to the formula since existence quantifier in the prefix of a formula is preferred in Z3.
-As a result, we got a result of {\tt sat} along with a model that satisfies PO3.
-Therefore, PO3 was proved.
-
-The results of all nine POs of the Abstracted Pacemaker VDM-SL model is showed in Table~\ref{tbl:result1} with information of whether negation is applied, and the time used.
+Table~\ref{tbl:results} shows the results of the two case studies. The checks are conducted on a machine with Intel Core i7 2.4GHz CPU and 8GB RAM.
 
@@
 \begin{center}
-\begin{tabular}{|c|c|r|r|}
+\begin{tabular}{|c|c|r|r||c|c|r|r|}
 \hline
-PO\#	&	negated	&	result	&	time (sec.) \\ \hline
-1		&	Y		&	sat		&	0.031 \\ \hline
-2		&	Y		&	unsat	&	0.031 \\ \hline
-3		&	N		&	sat		&	15.109 \\ \hline
-4		&	Y		&	unsat	&	0.032 \\ \hline
-5		&	Y		&	unsat	&	0.046 \\ \hline
-6		&	Y		&	unsat	&	0.048 \\ \hline
-7		&	Y		&	unsat	&	0.062 \\ \hline
-8		&	Y		&	unsat	&	0.031 \\ \hline
-9		&	Y		&	sat		&	0.047 \\ \hline
+\multicolumn{4}{|l||}{Abstract Pacemaker} & \multicolumn{4}{|l|}{Telephone Exchange} \\
+\hline \hline
+PO\#	&	neg.	&	res.	&	time(s)	&	PO\#	&	neg.	&	res.	&	time(s) \\ \hline
+1		&	+		&	sat		&	0.031	&	1		&	+		&	unsat	&	0.032 \\ \hline
+2		&	+		&	unsat	&	0.031	&	2		&	+		&	unsat	&	0.031 \\ \hline
+3		&	-		&	sat		&	15.109	&	3		&	+		&	unsat	&	0.031 \\ \hline
+4		&	+		&	unsat	&	0.032	&	4		&	+		&	unsat	&	0.032 \\ \hline
+5		&	+		&	unsat	&	0.046	&	5		&	+		&	unsat	&	0.03 \\ \hline
+6		&	+		&	unsat	&	0.048	&	6		&	+		&	unsat	&	0.047 \\ \hline
+7		&	+		&	unsat	&	0.062	&	7		&	-		&	sat		&	0.031 \\ \hline
+8		&	+		&	unsat	&	0.031	&	9		&	+		&	unsat	&	0.047 \\ \hline
+9		&	+		&	sat		&	0.047	&	12		&	+		&	sat		&	0.015 \\ \hline
+		&			&			&			&	14		&	+		&	sat		&	0.063 \\ \hline
+		&			&			&			&	17		&	+				&	sat		&	0.016 \\ \hline
+		&			&			&			&	18		&	+		&	unsat	&	0.047 \\ \hline
+		&			&			&			&	20		&	+		&	sat		&	0.016 \\ \hline
+		&			&			&			&	21		&	+		&	unsat	&	0.046 \\ \hline
+		&			&			&			&	23		&	+		&	sat		&	0.015 \\ \hline
+		&			&			&			&	25		&	+		&	sat		&	0.063 \\ \hline \hline
+\multicolumn{4}{|l||}{POs checked: 9 / 9 (100\%)} & \multicolumn{4}{|l|}{POs checked: 16 / 27 (59\%)} \\ \hline		
 \end{tabular}
 \end{center}
-\caption{Absract Pacemaker Result}
-\label{tbl:result1}
+\caption{Results of Case Studies}
+\label{tbl:results}
 \vspace{-20pt}
 \end{table}
-
-\subsection{Telephone Exchange}
-\label{section:case2}
-
-This model specifies an abstracted telephone exchange system.
-In this model, the operations specify the events which can be initiated either by the system or by a subscriber (user) with implicit style.
-The system state monitors the calling status of users and the connecting status among users.
-Overture tool generated 27 POs, and we selected PO1 and PO14 to report the case study.
-
-\subsubsection{Proof Obligation 1}
-
-PO1 shown in Fig.~\ref{fig:po1_case2} is a legal map application obligation that checks whether maps in the system state are applicable.
-In PO1, state {\tt Exchange} which is a record type of {\tt status} and {\tt calls} is mentioned.
-Therefore, the context information of PO1 is the state {\tt Exchange} including its invariant.
-Fig.~\ref{fig:types_case2} shows the specification of state {\tt Exchange} with corresponding quote types.
-The state {\tt Exchange} contains two maps: {\tt status} indicates the discrete calling status of users, and is a map from {\tt Subscriber} to {\tt Status}; {\tt calls} models the connection relation of users, i.e., which users are connected, and is a map from {\tt Subscriber} to {\tt Subscriber}.
-Note that {\tt calls} is specified as {\tt inmap} type which means a one-to-one map.
-This means if a user is calling another user, both of the users can not have other connections to third party users.
-Also, {\tt Exchange} has state invariant specifying the relations between connections of users and their calling status.
-At this point, the context information of PO1 is determined.
-
-\begin{figure}[t]
-\begin{center}
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small]
-forall
-  mk_Exchange(status, calls):EXCH`Exchange &
-    (forall i in set (dom calls) &
-      (i in set (dom status))
-    )
-)
-\end{Verbatim}
-\end{mdframed}
-\vspace{-10pt}
-\caption{PO1 of Telephone Exchange}
-\label{fig:po1_case2}
-\end{center}
-\vspace{-10pt}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small,numbers=left]
-Subscriber = token;
-Initiator =  <AI> | <WI> | <SI>;
-Recipient = <WR> | <SR>;
-Status = <fr> | <un> |
-         Initiator | Recipient;
-
-state Exchange of
-  status: map Subscriber to Status
-  calls:  inmap Subscriber to Subscriber
-inv mk_Exchange(status, calls) == 
-  forall i in set dom calls & 
-    (status(i) = <WI> and
-     status(calls(i)) = <WR>) or
-    (status(i) = <SI> and
-     status(calls(i)) = <SR>)
-init s == s = mk_Exchange({|->},{|->})
-end
-\end{Verbatim}
-\end{mdframed}
-\vspace{-10pt}
-\caption{Types and state of Telephone Exchange}
-\label{fig:types_case2}
-\end{center}
-\vspace{-20pt}
-\end{figure}
-
-The strategy of encoding the types shown in Fig.~\ref{fig:types_case2} is (1) applying user-defined types in Z3 to encode the quote types; (2) applying uninterpreted functions to encode map type.
-In Fig.~\ref{fig:encode_types_case2}, we demonstrate the encoding of selected types.
-In Fig.~\ref{fig:encode_types_case2}, the type {\tt Recipient} is defined as an user-defined type with values {\tt WR} and {\tt SR}.
-Type {\tt Status} is further defined as a user-defined type including {\tt Recipient} and {\tt Initiator}.
-Though {\tt Status} is a union type combines {\tt Initiator} and {\tt Recipient}, which can be considered an enumeration, we encoded {\tt Status} using user-defined type style in Z3py.
-The reason is that we prefer the subtypes {\tt Initiator} and {\tt Status} can be distinguishable in Z3.
-
-\begin{figure}[t]
-\begin{center}
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
-Recipient = Datatype('Recipient')
-Recipient.declare('WR')
-Recipient.declare('SR')
-Recipient = Recipient.create()
-
-Status = Datatype('Status')
-Status.declare('fr')
-Status.declare('un')
-Status.declare('INITIATOR',
-               ('get_initiator', Initiator))
-Status.declare('RECIPIENT',
-               ('get_recipient', Recipient))
-Status = Status.create()
-
-Status_lift = Datatype('Status_lift')
-Status_lift.declare('STATUS',
-                    ('get_status', Status))
-Status_lift.declare('NDF')
-Status_lift = Status_lift.create()
-
-status = Function('status',
-                  Subscriber,Status_lift)
-calls = Function('calls',
-                 Subscriber,Subscriber_lift)
-\end{Verbatim}
-\end{mdframed}
-\vspace{-10pt}
-\caption{Encoding of quote and map types in Fig.\ref{fig:types_case2} (partial)}
-\label{fig:encode_types_case2}
-\end{center}
-\vspace{-15pt}
-\end{figure}
-
-Because of VDM's logic of partial functions, for the type to be declared as the range of the map, i.e., {\tt Subscriber} and {\tt Status}, we defined their lifted types in Z3py to take into consideration the case that the value of a key of a map is undefined.
-Also, though {\tt Subscriber} is of token type which does not have specific values when declared, we treated token types as quote types with a predefined domain of values such as {\tt S1}, {\tt S2}.
-The two maps {\tt status} and {\tt calls} in the state {\tt Exchange} are encoded as uninterpreted functions and the state invariant is encoded by the encoded (and lifted) types mentioned above.
-
-Fig.~\ref{fig:encode_state_inv_case2} shows the encoding of the state invariant.
-Note that the encoding of {\tt i in set (dom calls)} (line 11 of Fig.~\ref{fig:types_case2}) is encoded as {\tt calls(i)!=Subscriber\_lift.NDF} (line 3 of Fig.~\ref{fig:encode_state_inv_case2}) because the set inclusion only means that the domain (key) of {\tt calls} has defined value in the range of {\tt calls}.
-Therefore, we do not have to encode the set of domain neither the set inclusion operator.
-The condition that the range value of a map is defined or not is essential here.
-For example, to encode {\tt status(calls(i)) = <WR>} (line 13 of Fig.~\ref{fig:types_case2}), we have to ensure the condition that {\tt calls(i)} is defined so that additional constraint {\tt calls(i)!=Subscriber\_lift.NDF} (line 9 of Fig.~\ref{fig:encode_state_inv_case2}) is added.
-
-% Instead of demonstrate the full encoding of the invariant, we only demonstrate the key encoding on {\tt i in set (dom calls)} and {\tt i in set (dom status)}.
-Though there is set inclusion used, instead of defining a set, we can encode the set inclusion as whether {\tt calls(i)} is defined or not.
-
-\begin{figure}[t]
-\begin{center}
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small,numbers=left]
-ForAll(i,
-  Implies(
-  calls(i)!=Subscriber_lift.NDF,
-  Or(
-    And(
-      status(i)==
-        Status_lift.STATUS(
-        Status.INITIATOR(Initiator.WI)),
-	  calls(i)!=Subscriber_lift.NDF,
-      status(Subscriber_lift.
-        get_subscriber(calls(i)))==
-        Status_lift.STATUS(
-          Status.RECIPIENT(Recipient.WR))),
-    And(
-      status(i)==
-        Status_lift.STATUS(
-        Status.INITIATOR(Initiator.SI)),
-      calls(i)!=Subscriber_lift.NDF,
-      status(Subscriber_lift.
-        get_subscriber(calls(i)))==
-        Status_lift.STATUS(
-          Status.RECIPIENT(Recipient.SR)))
-)))
-\end{Verbatim}
-\end{mdframed}
-\vspace{-10pt}
-\caption{Encoding of state invariant of Telephone Exchange}
-\label{fig:encode_state_inv_case2}
-\end{center}
 % \vspace{-20pt}
-\end{figure}
-
-Recall that {\tt calls} is defined as {\tt inmap}, a one-to-one map.
-Thus, we had to add  the conditions about {\tt inmap} as context information: for any two different users, either both users calls no one (undefined) or they are calling different users.
-Note that the latter case also includes the situation that only one of the two users is connecting to a third party user and the other user is calling no one.
-
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
-k  = Const('k',Subscriber)
-l  = Const('l',Subscriber)
-ForAll([k,l],
-  If(
-    k==l,
-    calls(k)==calls(l),
-    Or(
-      And(
-        calls(k)==Subscriber_lift.NDF,
-        calls(l)==Subscriber_lift.NDF
-      ),
-      calls(k)!=calls(l)
-    )
-  )
-)
-\end{Verbatim}
-\end{mdframed}
-
-The encoding of PO1 is done by encoding the negated formula of PO1.
-As the following code shows, the outside forall in the formula of PO1 is eliminated by obligation negation.
-The result was {\tt unsat} which means PO1 was proved.
-
-% \begin{mdframed}[roundcorner=5pt,shadow=true]
-% \begin{Verbatim}[fontsize=\small]
-%   calls(i) != Subscriber_lift.NDF
-%   status(i)!= Status_lift.NDF
-% \end{Verbatim}
-% \end{mdframed}
-
-% For PO1, we applied negation to the formula so that the {\tt forall} quantifier for maps {\tt calls} and {\tt status} is removed. The result was {\tt unsat} which means PO1 was proved.
-
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
-Not(
-  ForAll(i,
-    Implies(
-      calls(i)!=Subscriber_lift.NDF,
-      status(i)!=Status_lift.NDF
-    )
-  )
-)
-\end{Verbatim}
-\end{mdframed}
-
-\subsubsection{Proof Obligation 14}
-
-PO14 shown in Fig.~\ref{fig:PO14_case2} is an enumeration map injectivity obligation of the operation {\tt Answer} of the system.
-PO14 has free variables, {\tt r} and {\tt status}, in its formula.
-These free variables are treated as forall quantifier bounded.
-There are also intermediate variables in the formula such as {\tt m1}, {\tt m2}, {\tt d3}, and {\tt d4}.
-Note that the map inverse function {\tt inverse} (line 4) and the range restriction operator `{\tt :>}' (line 1) are applied to {\tt calls} and {\tt status} separately.
-(Recall that {\tt calls} is a one-to-one map {\tt inmap}.) The context information of PO14 includes the two maps {\tt status} and {\tt calls} but not the state {\tt Exchange} because {\tt Exchange} is not mentioned in PO14.
-Therefore, only the two maps have to be encoded, and the state invariant is excluded in the encoding.
-The encoding of the constraints for {\tt calls} of type {\tt inmap} described in solving PO1 can be reused here.
-
-\begin{figure}[t]
-\begin{center}
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small,numbers=left]
-(r in set (dom (status :> {<WR>}))) =>
-  (forall m1, m2 in 
-    set {{r |-> <SR>}, 
-         {(inverse calls)(r) |-> <SI>}} &
-    (forall d3 in set (dom m1),
-            d4 in set (dom m2) &
-      ((d3 = d4) => (m1(d3) = m2(d4)))
-))
-\end{Verbatim}
-\end{mdframed}
-\vspace{-10pt}
-\caption{Encoding of state invariant of Telephone Exchange}
-\label{fig:PO14_case2}
-\end{center}
-\vspace{-25pt}
-\end{figure}
-
-When encoding the formula of PO14, firstly, we negated PO14 and eliminated the forall quantifiers.
-To encode the negated formula of PO14, a similar technique for  encoding {\tt set (dom status)} in PO1 can be applied.
-For the range restriction function, since the range is a singleton set with an element {\tt <WR>}, we do not have to encode the set notations and only limit the map's range {\tt status(r)} to {\tt <WR>}.
-Thus, {\tt r in set (dom (status :> \{<WR>\}))} is encoded as follows.
-
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small]
-status(r)==
-  Status_lift.STATUS(
-    Status.RECIPIENT(Recipient.WR) )
-\end{Verbatim}
-\end{mdframed}
-
-For the map inverse function, we did not define a new map as the inverse of {\tt calls}, but encoded {\tt m1 = \{(inverse calls)(r) |-> <SI>\}} as a case of {\tt m1 in set \{\{r |-> <SR>\}, \{(inverse calls)(r) |-> <SI>\}\}}.
-The case was encoded as the collection of several conditions.
-
-\begin{mdframed}[roundcorner=5pt,shadow=true]
-\begin{Verbatim}[fontsize=\small,numbers=left]
-# ensure the inverse map is not empty
-calls(i)!=Subscriber_lift.NDF 
-# inverse_calls(r) == i
-Subscriber_lift.get_subscriber(calls(i))==r
-
-m1(i)==Status_lift.STATUS(
-         Status.INITIATOR(Initiator.SI)),
-
-ForAll(k, 
-  Implies(k!=i, m1(k)==Status_lift.NDF) )
-\end{Verbatim}
-\end{mdframed}
-
-In above code, the first two conditions (lines 1-4) promise that (1) the part of map, where key {\tt i} is used and {\tt inverse} is applied, has to be defined; (2) the domain key of the inversed map, i.e., the range value of the original map, is {\tt r}.
-Then {\tt m1(i)} where key {\tt i} is referred should be {\tt <SI>} (lines 6-7).
-(3) the last condition guarantees that {\tt m1} should be undefined for all indices except {\tt i} (lines 9-10) because {\tt m1} is a single map that has only one key and one value.
-
-The result of checking PO14 (negated) was expected to be {\tt unsat}, but we got {\tt sat} instead.
-Z3 returned a model that satisfies the negated formula of PO14.
-After examined the model returned by Z3, we found that {\tt calls} in the returned model, {\tt calls} is a map where users call themselves.
-This does not violate the type constraint of {\tt inmap}, but a user is not allowed to call him/her-self in a telephone exchange system.
-Therefore, the checking of PO14 suggested that we should add a constraint on {\tt calls} to the VDM-SL model.
-We may add the constraint below as the type invariant of {\tt calls} or add it to the state invariant\footnote{We have added the constraint and checked again, and PO14 was proved successfully.}.
-
-\begin{mdframed}[roundcorner=5pt]
-\begin{Verbatim}[fontsize=\small]
-forall i in set dom calls & calls(i) != i
-\end{Verbatim}
-\end{mdframed}
-
-In this case study, we did not check all the twenty-seven POs since the encoding demonstrated in this paper could not cover all POs.
-The main issue is that the maps are encoded as uninterpreted functions, which cannot be used with {\tt forall} quantifier.
-This results that we could not check the postcondition satisfiable obligations of this model with the current encoding\footnote{Besides postcondition satisfiable obligations, there are two POs with type check function {\tt is\_(name, type)}, for which Z3 encoding is not available and then skipped.}.
-After all, we successfully checked 16 out of all 27 POs for the telephone exchange VDM-SL model.
-The results of 16 POs checked of the Telephone Exchange VDM-SL model is showed in Table~\ref{tbl:result2} with information of whether negation is applied, and the time used.
-
-% total check result (table)
-%
-\begin{table}[htb]
-\begin{center}
-\begin{tabular}{|c|c|r|r|}
-\hline
-PO\#	&	negated	&	result	&	time (sec.) \\ \hline
-1		&	Y		&	unsat	&	0.032 \\ \hline
-2		&	Y		&	unsat	&	0.031 \\ \hline
-3		&	Y		&	unsat	&	0.031 \\ \hline
-4		&	Y		&	unsat	&	0.032 \\ \hline
-5		&	Y		&	unsat	&	0.03 \\ \hline
-6		&	Y		&	unsat	&	0.047 \\ \hline
-7		&	N		&	sat		&	0.031 \\ \hline
-9		&	Y		&	unsat	&	0.047 \\ \hline
-12		&	Y		&	sat		&	0.015 \\ \hline
-14		&	Y		&	sat		&	0.063 \\ \hline
-17		&	Y		&	sat		&	0.016 \\ \hline
-18		&	Y		&	unsat	&	0.047 \\ \hline
-20		&	Y		&	sat		&	0.016 \\ \hline
-21		&	Y		&	unsat	&	0.046 \\ \hline
-23		&	Y		&	sat		&	0.015 \\ \hline
-25		&	Y		&	sat		&	0.063 \\ \hline
-\end{tabular}
-\end{center}
-\caption{Telephone Exchange Result}
-\label{tbl:result2}
-\vspace{-20pt}
-\end{table}
-
-% Overture generated 27 POs for this model.
-These POs can be divided into several groups as showed below.
-In this case study we encoded and checked the first three groups using the above encoding of types.
-The reason of not handling the postcondition obligations is that the maps defined using uninterpreted functions cannot be used together with quantifiers.
-Also, map inverse obligations have {\tt is\_(name, type)}, the function of type check provided by VDM, in their formulas.
-There is no appropriate encoding for such type check functions in Z3.
-
-% \begin{enumerate}
-% \item
-% Legal map application obligations, invariant satisfiable obligations: 1 to 7.
-% \item
-% Enumeration map injectivity obligations: 9, 14, 18, 21, 25.
-% \item
-% Legal map application obligations with inverse map: 12, 17, 20, 23.
-% \item
-% Operation postcondition satisfiable obligations: 8, 10, 11, 15, 16, 19, 22, 26, 27
-% \item
-% Map inverse obligations: 13, 24.
-% \end{enumerate}
-
-\subsection{Discussion}
-
-As an overall result of our case studies, the objective of our approach that to discharge POs of a VDM-SL model as many as possible is considered achieved.
-In the first case study (Abstracted Pacemaker), we discharged all nine POs; in the second case study, we discharged 16 POs our of total 27 POs (about 60\%).
-To increase the percentage of discharging POs, we may further apply different encoding, for example, encoding maps using arrays in Z3, to solve remaining POs in the Telephone Exchange case study.
-
-% Usability and Applicability of the Encode Style
-The two case studies demonstrated the application of the encode strategy described in Section~\ref{section:encode-strategy}.
-From our experiences, the cost and complexity of encoding varies from specifications in a VDM-SL model.
-Usually, the complexity of encoded formulas usually depends on the complexity of VDM-SL types.
-For example, natural number based types such as {\tt seq of nat} or {\tt map from nat to nat} have less complexity, and one may encode and solve this kind of types with SMT solvers without difficulty.
-On the contrary, for VDM types such as a map of sequences or a set of maps, the encoding may be difficult, and the complexity of the encoded formulas may grow a lot.
-
-The encoding demonstrated in the two case studies also showed some encoding patterns.
-For example, the encoding of union type and quote type with user-defined types or enumeration type, and the encoding of maps with uninterpreted functions, are useful to encoding POs of other VDM-SL models.
-Also, the encoding patterns like the constraints on arrays to represent a sequence, set inclusion of indices of a sequence, or set inclusion of domain/range of a map, are also reusable.
-In our case studies, these encoding patterns are applied multiple times.
-
-% Automation of the encoding style and strategy
-Though the encoding patterns are useful, applying the encoding patterns still depends on specification of VDM models because the encoding patterns are not fully systemic or syntax-based.
-Considering automated encoding of POs with related context information, the encoding of types may be fine, but encoding expressions remains difficult and rely on manually applying encoding patterns.
-
-% Scalability and Efficiency}
-Our case studies also showed the efficiency of SMT solvers can do great help in discharging POs.
-Also, the counterexample (model) generation of SMT solvers provides helpful hints for revision and correction of VDM models.
-As reported in Section~\ref{section:case2}, encoding and solving PO14 of the telephone exchange VDM-SL model found an inconsistency and the counterexample provided guidance to dissolving the inconsistency.
-Finally, the scalability of our approach is not serious because we only encode the formula of a PO and its context information.
-The size of VDM-SL code to be encoded does not escalate with the size of a VDM-SL model.
-
-% \footnote{A previous work~\cite{kenneth:ifm2013} that translates VDM-SL to Alloy~\cite{kenneth:ifm2013} checked the VDM-SL model and found a different problem.
-The model (from the Overture repository) we used for case study is the fixed one which means that the inconsistency we found was not found in~\cite{kenneth:ifm2013}.}
-
-% Points of discussion:
-% (1) Automation: Is the encoding can be automated? partly or fully? 
-% (2) Patterns: What patterns are repeated used in encoding?

### 2018-08-15 memocode_cr/related.tex
Add MEMOCODE'17 camera ready and presentation files

@@
 The idea of adopting existing theories of SMT is the same as our approach.
-However, we argue that a strategic encoding like our approach might is necessary for discharging VDM POs than introducing general theories because the infinite and undecidable nature of sets, lists, and maps in VDM.
+However, we argue that a strategic encoding like our approach is necessary for discharging VDM POs than introducing general theories because the infinite and undecidable nature of sets, lists, and maps in VDM.
 
